### 2024-03-11 R/paper.Rmd
Added datasets and R scripts. Updated until fig. 1

---
title: "Systemix Profiling Article: Combined MAG-SMAG Analysis"
author: "Moises A. Rojas"
date: "2023-12-01"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = '/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/')

suppressMessages(library(ape)) #tree basics
suppressMessages(library(treeio))
suppressMessages(library(ggtree))
suppressMessages(library(ggplot2))
suppressMessages(library(ggtreeExtra)) #geom_fruit
suppressMessages(library(ggnewscale)) #scales in ggplot2
suppressMessages(library(ggstar)) #geom_star for ggplot2
suppressMessages(library(phytools)) #MRCA
suppressMessages(library(tidyverse)) #strings, dataframes processing
suppressMessages(library(magrittr)) #pipes %>%
suppressMessages(library(tidytree)) #trees as dataframes
suppressMessages(library(RColorBrewer)) #colors palette
suppressWarnings(library(colorspace))
suppressWarnings(library(patchwork))
```

## Prepare and merge MAG-SMAG metadata

```{r magsdata}
library(openxlsx)

# Read Systemix MAGs taxonomy and QC files
gtdb_taxonomy <- read.csv("1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
mags_qc <- read.csv("2-statistics/checkm_parsed_bins.tsv", sep = "\t")

# Inspect completeness and contamination per sample, then select one
filter(mags_qc, grepl("\\-S15\\.", bin)) %>%
  dplyr::select(2:3) %>%
  summary()
mag_samples <- filter(gtdb_taxonomy, grepl("\\-S15\\.", genome))

# Add QC attributes
colnames(mag_samples)[1] <- "bin"
mag_samples <- left_join(mag_samples, mags_qc, by = c("bin"))

# Parse and reorder columns
mag_samples <- dplyr::select(mag_samples, -c("taxonomy", "gc", "contigs"))
mag_samples <- mag_samples %>%
  mutate_if(is.character, function(x) gsub("d__|p__|c__|o__|f__|g__|s__", "", x))
mag_samples <- mag_samples %>% separate(bin, "\\.", into = c("mag_name", "mag_id"))
mag_samples <- mag_samples[, c(1, 2, 10, 11, 12, 13, 3, 4, 5, 6, 7, 8, 9)]

# Define final attributes
mag_samples['ecosystem'] <- "Biomining"
mag_samples['continent'] <- "South America"

# Load SMAG metadata (tables 1 and 2)
smag_md1 <- read.xlsx("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/SMAG/Extended_Data_Table_1.xlsx")
smag_md2 <- read.xlsx("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/SMAG/Extended_Data_Table_2.xlsx")

# Save attributes
smag_md1_sel <- smag_md1 %>% dplyr::select("mag_name", "Ecosystem", "Continent")
smag_md2_sel <- smag_md2 %>%
  dplyr::select("user_genome", "completeness", "contamination",
                "strain_heterogeneity", "N50", "classification",
                "phylum", "class", "order", "family", "genus", "species")

# Merge selections and parse taxa
smag_data <- smag_md2_sel %>% separate(user_genome, "\\.", into = c("mag_name", "mag_id"))
smag_data <- left_join(smag_data, smag_md1_sel, by = c("mag_name"))
smag_data <- smag_data %>%
  mutate_if(is.character, function(x) gsub("d__|p__|c__|o__|f__|g__|s__", "", x))

# Rename columns
colnames(smag_data)[7] <- "kingdom"
colnames(smag_data)[14] <- "ecosystem"
colnames(smag_data)[15] <- "continent"

# Clean memory
rm(smag_md1, smag_md2, smag_md1_sel, smag_md2_sel, mags_qc)

# Merge metadata and save to file
mags_metadata <- merge(mag_samples, smag_data, all = TRUE)
if (file.exists("4-mags-smag/mags_metadata.tsv") == FALSE) {
  write.table(mags_metadata, "4-mags-smag/mags_metadata.tsv",
              sep = "\t", quote = FALSE, row.names = FALSE)
}

# Save parsed SMAG metadata
if (file.exists("4-mags-smag/smag_data.tsv") == FALSE) {
  write.table(smag_data, "4-mags-smag/smag_data.tsv",
              sep = "\t", quote = FALSE, row.names = FALSE)
}
```

## In-depth taxonomic analysis of MAG-SMAG

```{r magstaxa}
# Biomining provides a new phylum -> Nitrospirota_A
mag_samples_new_phy <- mag_samples[!(mag_samples$phylum %in% smag_data$phylum), ]

# Explore metadata: some bins are not assigned to ecosystems
mags_metadata %>%
  filter(is.na(ecosystem)) %>%
  group_by(mag_name) %>%
  dplyr::summarise() #n=401

# Keep only bacteria and defined ecosystems (discard archaea kingdom and non-NA eco.)
mags_metadata_bac <- mags_metadata %>%
  filter(kingdom == "Bacteria" & !is.na(ecosystem)) #n=32975

# Save to file
if (file.exists("4-mags-smag/mags_metadata_bac.tsv") == FALSE) {
  write.table(mags_metadata_bac, "4-mags-smag/mags_metadata_bac.tsv",
              sep = "\t", quote = FALSE, row.names = FALSE)
}

# Select top phyla and plot a histogram
mags_md_top_phy <- table(mags_metadata_bac$phylum) %>%
  sort(TRUE) %>%
  head(n = 30) %>%
  as.data.frame() %>%
  dplyr::rename(phy_top = Var1, freq = Freq)

plot_md_top_phyla <- ggplot(mags_md_top_phy, aes(x = reorder(phy_top, freq), y = freq)) +
  geom_bar(stat = "identity", fill = "#459682") +
  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.5, size = 3) +
  labs(x = "Phylum", y = "Count") +
  scale_y_continuous(limits = c(0, 10000)) +
  theme_classic() +
  coord_flip()
ggsave("plot_md_top_phyla.pdf", plot_md_top_phyla, height = 7, width = 6)

## Filtering metadata

# Set colors for ecosystem
#mags_colors_eco <- vector()
#for (i in 1:length(kegg_colors_eco)) {
#  mags_colors_eco <- append(mags_colors_eco, kegg_colors_eco[[i]])
#}
#mags_colors_eco <- colorRampPalette(brewer.pal(12, "Paired"))(length(unique(mags_metadata_bac$ecosystem)))
#mags_colors_eco <- sample(mags_colors_eco)
mags_colors_eco <- c("#1F78B4", "#FF7F00", "#B2DF8A", "#E31A1C", "#A6CEE3",
                     "#6A3D9A", "#33A02C", "#CAB2D6", "#FDBF6F", "#FB9A99")

# Which bacteria from biomining sites are annotated as known?
mags_depth_spec_biom <- mags_metadata_bac %>%
  dplyr::select(mag_name, family, genus, species, ecosystem) %>%
  filter(species != "" & ecosystem == "Biomining")

#are they present in other SMAG?
mags_depth_spec <- mags_metadata_bac[mags_metadata_bac$species %in%
                                     mags_depth_spec_biom$species, ] %>%
  dplyr::select(mag_name, mag_id, phylum, family, genus, species, ecosystem, continent) %>%
  group_by(ecosystem)

plot_depth_spec <- ggplot(mags_depth_spec) +
  geom_bar(aes(x = species, fill = phylum)) +
  labs(x = "Known biomining species", y = "Count") +
  scale_fill_manual(values = "#DD7E8B") +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = c(0.85, 0.05),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  facet_wrap(.~ecosystem)
ggsave("plot_md_biom_species.pdf", plot_depth_spec, height = 6, width = 8)

# Which ecosystem has more known species (# and ratio)?
mags_depth_spec_eco <- mags_metadata_bac %>%
  dplyr::filter(species != "") %>%
  group_by(species, ecosystem) %>%
  dplyr::summarise(count_sp = table(species)) %>% #.groups = "drop"
  ungroup() %>%
  group_by(ecosystem) %>%
  dplyr::summarise(count_eco = sum(count_sp)) %>%
  dplyr::arrange(desc(count_eco)) #max 1821

#count number of MAGs per ecosystem
mags_depth_spec_eco_n <- mags_metadata_bac %>%
  group_by(ecosystem) %>%
  dplyr::summarise(mags = n_distinct(mag_name))
mags_depth_spec_eco <- left_join(mags_depth_spec_eco, mags_depth_spec_eco_n, by = c("ecosystem"))
rm(mags_depth_spec_eco_n)

#add column
mags_depth_spec_eco <- mags_depth_spec_eco %>%
  dplyr::mutate(mags = replace(mags, ecosystem == "Biomining", 1)) %>%
  dplyr::mutate(ratio = count_eco / mags) #max 6.52

plot_depth_spec_eco <- ggplot(mags_depth_spec_eco, aes(x = reorder(ecosystem, ratio),
                                                       y = ratio, fill = ecosystem)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.5, size = 3) +
  labs(x = "Ecosystem", y = "Number of kSGBs / Number of MAGs") +
  scale_y_continuous(limits = c(0, 10)) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none") +
  coord_flip()
ggsave("plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)

# What about unknown/known species-level genome bins (SGBs) at phylum level?
mags_depth_sgbs <- mags_metadata_bac %>%
  group_by(phylum, species) %>%
  dplyr::summarise(n_spe = n()) %>%
  ungroup() %>%
  group_by(phylum) %>%
  dplyr::mutate(t_phy = sum(n_spe)) %>% ###
  dplyr::filter(species == "") %>%
  dplyr::mutate(diff_k = t_phy - n_spe) %>%
  dplyr::mutate(perc_u = n_spe * 100 / t_phy) %>%
  dplyr::arrange(desc(t_phy)) %>%
  pivot_longer(cols = c(n_spe, diff_k), names_to = "cond", values_to = "tot")

plot_depth_sgbs <- ggplot(mags_depth_sgbs[1:60, ],
                          aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"),
                    values = c("#DD7E8B", "#459682")) +
  scale_y_continuous(limits = c(0, 8000)) +
  labs(x = "Phylum", y = "Number of species") +
  theme_classic() +
  theme(legend.position = c(0.85, 0.15), strip.placement = "outside") +
  coord_flip()
#ggsave("plot_md_SGBs_phylum.pdf", plot_depth_sgbs, height = 7, width = 6)

#add % column
mags_depth_sgbs_p <- mags_depth_sgbs %>%
  dplyr::select(-c("species")) %>%
  dplyr::mutate(perc_k = 100 - perc_u) %>%
  pivot_longer(cols = c(perc_u, perc_k), names_to = "cond_p", values_to = "tot_p") %>%
  dplyr::filter(cond == "n_spe") #just to remove 1 row per phylum/cond

plot_depth_sgbs_p <- ggplot(mags_depth_sgbs_p[1:60, ],
                            aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"),
                    values = c("#DD7E8B", "#459682")) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "", y = "% of species") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank()) +
  coord_flip()
#ggsave("plot_md_SGBs_phylum_p.pdf", plot_depth_sgbs_p, height = 7, width = 5)

plot_depth_sgbs_both <- (plot_depth_sgbs | plot_depth_sgbs_p) +
  plot_layout(nrow = 1, ncol = 2, widths = c(0.55, 0.5))
ggsave("plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)

# And unknown/known SGBs for biomining?
mags_depth_sgbs_biom <- mags_metadata_bac %>%
  dplyr::filter(ecosystem == "Biomining") %>%
  group_by(phylum, species) %>%
  dplyr::summarise(n_spe = n()) %>%
  ungroup() %>%
  group_by(phylum) %>%
  dplyr::mutate(t_phy = sum(n_spe)) %>%
  dplyr::filter(species == "") %>%
  dplyr::mutate(diff_k = t_phy - n_spe) %>%
  dplyr::mutate(perc_u = n_spe * 100 / t_phy) %>%
  dplyr::arrange(desc(t_phy)) %>%
  pivot_longer(cols = c(n_spe, diff_k), names_to = "cond", values_to = "tot")

plot_depth_sgbs_biom <- ggplot(mags_depth_sgbs_biom,
                               aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"),
                    values = c("#DD7E8B", "#459682")) +
  scale_y_continuous(limits = c(0, 25)) +
  labs(x = "Phylum", y = "Number of biomining species") +
  theme_classic() +
  theme(legend.position = c(0.75, 0.2), strip.placement = "outside") +
  coord_flip()
ggsave("plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 5, width = 4)

# What is the distribution of uSGBs by ecosystem?
mags_depth_usgbs_eco <- mags_metadata_bac %>%
  dplyr::filter(species == "") %>%
  group_by(phylum, ecosystem) %>%
  dplyr::summarise(num = table(ecosystem)) %>%
  dplyr::mutate(phylum = replace(phylum, num < 5, "Other")) #group lows as "other"

plot_depth_usgbs <- ggplot(mags_depth_usgbs_eco,
                           aes(x = reorder(phylum, num), y = num, fill = ecosystem)) +
  geom_col() +
  labs(x = "Phylum", y = "Number of uSGBs") +
  scale_y_continuous(n.breaks = 3) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none",
        strip.placement = "outside",
        panel.border = element_blank()) +
  facet_grid(.~ecosystem, scales = "free") +
  coord_flip()
ggsave("plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 8, width = 12)

# What about the annotated bins at genus level (kGGBs) from biomining?
mags_depth_kggbs_biom <- mags_metadata_bac %>%
  dplyr::select(mag_name, mag_id, phylum, family, genus, ecosystem) %>%
  dplyr::filter(genus != "" & ecosystem == "Biomining") #nrow=21

#match with the other ecosystems
mags_depth_kggbs_biom <- mags_metadata_bac[mags_metadata_bac$genus %in%
                                           mags_depth_kggbs_biom$genus, ] %>%
  dplyr::select(mag_name, mag_id, phylum, family, genus, ecosystem) %>%
  group_by(genus) %>%
  dplyr::mutate(tot_gen = n()) %>%
  dplyr::arrange(desc(tot_gen)) #nrow=535

#scale to 0-100%
mags_depth_kggbs_biom <- mags_depth_kggbs_biom %>%
  group_by(genus, ecosystem) %>%
  dplyr::mutate(sum = n()) %>% #sum genus per ecosystem
  dplyr::mutate(norm = sum / tot_gen * 100) %>%
  distinct(genus, ecosystem, .keep_all = TRUE) #reduce to uniques (plot scale!)

plot_depth_kggbs_biom <- ggplot(mags_depth_kggbs_biom,
                                aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_x_continuous(limits = c(0, 100), n.breaks = 5) +
  scale_fill_manual(values = mags_colors_eco) +
  labs(x = "% kGGBs", y = "Known genus from biomining") +
  theme_classic() +
  theme(legend.title = element_blank())
ggsave("plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)

### --OTHER/PENDING TAXA-- ###
# Inspect unannotated family-level bins (uFGBs)? #1166
#mags_depth_ufgbs

# What about novel bacteria at genus level?
mags_depth_novel_gen <- mags_metadata_bac %>%
  dplyr::filter(genus == "") %>%
  group_by(phylum, genus, ecosystem) %>%
  dplyr::summarise(gen_unk = n()) %>%
  #arrange(desc(gen)) %>%
  ungroup() %>%
  group_by(genus) %>%
  dplyr::mutate(tot_gen = sum(gen_unk)) %>%
  dplyr::mutate(perc = gen_unk * 100 / tot_gen)

# Get unique classes of unknown genus by ecosystem
#Leptospirillia
#Sulfobacillia
mags_md_unk_gen2 <- mags_metadata_bac %>%
  dplyr::filter(genus == "") %>%
  group_by(class, ecosystem) %>%
  tally() %>%
  #slice(n()) %>%
  ungroup()
unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Biomining", "class"])

# Unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
unique(mag_samples[mag_samples[, "family"] != "", "genus"])
mags_metadata_bac %>% filter(grepl("Acidithrix", genus, ignore.case = TRUE) == TRUE)

##plot
plot_md_unk_cla <- ggplot(mags_md_unk_gen2[mags_md_unk_gen2$n > 10, ],
                          aes(x = reorder(class, n), y = n, fill = ecosystem)) +
  geom_bar(stat = "identity") +
  labs(x = "Class", y = "Count") +
  scale_y_continuous(limits = c(0, 250)) +
  theme_minimal() +
  coord_flip() +
  facet_wrap(~ecosystem)
##
ggsave("plot_md_unk_class_eco.pdf", plot_md_unk_cla, height = 7, width = 10)

## Class of interest of acidic soils:
# Alphaproteobacteria, Gammaproteobacteria
# Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
table(mags_metadata_bac[mags_metadata_bac[, "phylum"] == "Proteobacteria", "class"])
table(mags_metadata_bac[mags_metadata_bac[, "class"] == "Acidimicrobiia", "order"])

## Family of interest reported for Systemix [ref 2022]:
# Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
mags_md_fams_acid <- mags_metadata_bac %>%
  filter(family == "Pseudomonadaceae") %>%
  group_by(ecosystem) %>%
  dplyr::summarise(count = n())
```

### QC (biomining)

```{r magsqc}
# Read QC file from CheckM
mags_qc <- read.csv("2-statistics/checkm_parsed_bins.tsv", sep = "\t")

# Select current sample
mags_qc <- mags_qc %>% filter(grepl("Refined-S15", bin))

# Parse QCs
mags_qc_hq <- mags_qc %>%
  dplyr::filter(completeness >= 90 & contamination <= 5) %>%
  dplyr::summarise(mags = n(), contigs = sum(contigs))
mags_qc_hq$label = "High Quality"

mags_qc_mq <- mags_qc %>%
  dplyr::filter(completeness < 90 & completeness >= 50 & contamination <= 10) %>%
  dplyr::summarise(mags = n(), contigs = sum(contigs))
mags_qc_mq$label = "Medium Quality"

mags_qc_lq <- mags_qc %>%
  dplyr::filter(completeness < 50 & contamination <= 10) %>%
  dplyr::summarise(mags = n(), contigs = sum(contigs))
mags_qc_lq$label = "Low Quality"

mags_qcs <- rbind(mags_qc_hq, mags_qc_mq, mags_qc_lq)
rm(mags_qc_hq, mags_qc_mq, mags_qc_lq)

# Plot
plot_mags_qc <- ggplot(mags_qcs,
                       aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
  geom_col(position = "dodge2") +
  scale_fill_manual(values = c("#DB7971", "#5CA650", "#859AC7")) +
  labs(x = "Number of MAGs", y = "") +
  theme_classic() +
  theme(legend.position = "none")
ggsave("plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 3, width = 5)
```

### Phylogenetic tree (biomining)

```{r magstree}
# Load GTDB taxonomy
gtdb_taxonomy <- read.csv("1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)

# Subset taxonomy to selected S15 sample and remove underscores
gtdb_taxonomy <- gtdb_taxonomy %>%
  dplyr::filter(grepl("Refined-S15", genome) == TRUE) %>%
  dplyr::select(-c("taxonomy")) %>%
  mutate_all(function(x) gsub("d__|p__|c__|o__|f__|g__|s__", "", x))

# Read full tree (S15)
tree <- ape::read.tree("1-taxonomy/trees-gtdbtk/gtdbtk.DASTool-S15.tree")

# Subset to biomining placements
tree_s15 <- keep.tip(tree, as.character(gtdb_taxonomy$genome))
tree_s15 #44 tips

# Export tree to file
if (file.exists("1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree") == FALSE) {
  ape::write.tree(tree_s15, "1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree")
}
rm(tree)

# Rename column {genome -> label} (to match with tree) from GTDB data
colnames(gtdb_taxonomy)[1] <- "label"

# Convert subset tree to dataframe and join some taxa
tree_s15_df <- tidytree::as_tibble(tree_s15)
tree_s15_df <- treeio::full_join(tree_s15_df, gtdb_taxonomy[c("label", "phylum", "family", "genus", "species")], by = "label")

# Back tree to phylogenetic object
tree_s15_md <- tidytree::as.treedata(tree_s15_df)

# Assign colors to each phylum
tree_colors <- colorRampPalette(brewer.pal(12, "Paired"))(length(unique(gtdb_taxonomy$phylum)))

# Plot circular tree
plot_tree_s15 <- ggtree(tree_s15_md, layout = "circular", size = 0.2)

# Plot tree adding phylum
plot_tree_s15_phyla <- plot_tree_s15 +
  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1, show.legend = FALSE) +
  geom_tiplab(aes(label = phylum, color = phylum),
              size = 1.5, hjust = -0.1, show.legend = FALSE) +
  scale_colour_discrete(breaks = c('c', 'd'), na.value = NA) +
  scale_colour_manual(values = tree_colors, na.translate = FALSE) +
  theme()
  #+guides(color = guide_legend(title = "Phylum", ncol = 1, order = 1))
ggsave("plot_biom_MAGs_tree_phyla.pdf", plot_tree_s15_phyla, height = 5, width = 7)

# Process QC attributes to match tree tips order
tree_tips <- tree_s15$tip.label
tree_com <- vector()
tree_con <- vector()
tree_st  <- vector()
tree_gs  <- vector()
tree_gc  <- vector()
tree_pg  <- vector()

# Match datasets
for (i in 1:nrow(tree_s15_df)) {
  for (j in 1:nrow(mags_qc)) {
    if (tree_s15_df$label[i] == mags_qc$bin[j]) {
      tree_com <- append(tree_com, mags_qc$completeness[j])
      tree_con <- append(tree_con, mags_qc$contamination[j])
      tree_st  <- append(tree_st,  mags_qc$strain_heterogeneity[j])
      tree_gs  <- append(tree_gs,  mags_qc$genome_size[j])
      tree_gc  <- append(tree_gc,  mags_qc$gc[j])
      tree_pg  <- append(tree_pg,  mags_qc$predicted_genes[j])
    }
  }
}

# Save attributes in a dataframe
tree_qc_df <- tibble(Bin = tree_tips, Completeness = tree_com,
                     Contamination = tree_con, Strain_Het = tree_st,
                     Genome_Size = tree_gs, GC = tree_gc, Pred_Genes = tree_pg)
rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg)

# Plot tree adding QC as multiple layers
plot_tree_s15_md <- plot_tree_s15_phyla + new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Completeness, fill = Completeness),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1),
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3),
             pwidth = 0.15, offset = 0.35) +
  scale_fill_gradient(low = "#1F4060", high = "#4CA8F6",
                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
                                           reverse = TRUE, order = 1)) +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Contamination, fill = Contamination),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1),
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1),
             pwidth = 0.15, offset = 0.1) +
  scale_fill_gradient(low = "#6A60A8", high = "#C4C2DE",
                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
                                           reverse = TRUE, order = 2)) +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Strain_Het, fill = Strain_Het),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1),
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3),
             pwidth = 0.15, offset = 0.1) +
  scale_fill_gradient(name = "Strain Heterogeneity",
                      low = "#C9543F", high = "#E9bAB2",
                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
                                           reverse = TRUE, order = 3)) +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Pred_Genes, fill = Pred_Genes),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1),
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1),
             pwidth = 0.15, offset = 0.1) +
  scale_fill_gradient(name = "Predicted Genes",
                      low = "#279854", high = "#A9DEDE",
                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
                                           reverse = TRUE, order = 4)) +
  theme(legend.title = element_text(size = 8),
        legend.text = element_text(size = 7),
        legend.key.height = unit(3.5, units = "mm"))
ggsave("plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 10)
```

## Figure 1

```{r fig1}
# Merge generated figures using patchwork
fig1_a <- {
  (plot_tree_s15_md | plot_depth_spec_eco)
  } + plot_layout(ncol = 2, widths = c(1, 0.5))
ggsave("figures/fig1_a.pdf", fig1_a, height = 7, width = 14)

fig1_b <- {
  (plot_mags_qc / plot_depth_sgbs_biom) |
    plot_depth_sgbs_both | plot_depth_kggbs_biom
  } + plot_layout(ncol = 3, widths = c(0.5, 2, 0.75))
ggsave("figures/fig1_b.pdf", fig1_b, height = 7, width = 14)

##fig1
fig1 <- (fig1_a / fig1_b) +
  plot_annotation(tag_levels = list(c("a", "e", "b", "c", "d", "", "f")))
ggsave("figures/fig1.pdf", fig1, height = 14, width = 14)
rm(fig1_a, fig1_b)

##fig2
```

## Functional Annotation: KEGG KOs from KOfamScan and Anvio Kofam

### Heatmaps

```{r annothmap}
library(ComplexHeatmap)

### 1. Read output file from KOfamScan hits
kegg_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/merge_hits/kofamscan_hits.tsv", sep = "\t")

# Remove uncharacterized and duplicates proteins
kegg_annot <- kegg_annot %>%
  filter(!grepl("uncharacterized", KO.definition) == TRUE) %>%
  distinct(KO.definition, .keep_all = TRUE)

# Extract remaining KO definitions and set KOs as row names
kegg_kos <- data.frame(KO = kegg_annot$KO, Definition = kegg_annot$KO.definition)
kegg_annot <- kegg_annot %>%
  select(-c("KO.definition")) %>%
  data.frame(row.names = 1)

# Set definitions for pathways of interest
kegg_paths <- kegg_kos %>%
  #"sulf": to group sulfur, sulfate, sulfane, sulfide, sulfo_
  #"soxr": SoxR 2Fe-2S cluster
  filter(grepl("iron|sulf|soxr|copper", Definition, ignore.case = TRUE) == TRUE) %>%
  select(KO, Definition)

## FUNCTIONS ##
## Function that returns the pathway based on the definitions
fx_setKOpathways <- function(def) {
  if (grepl("copper", def, ignore.case = TRUE) == TRUE)
    return("Copper")
  if (grepl("sulf|sox", def, ignore.case = TRUE) == TRUE)
    return("Sulfur")
  if (grepl("iron", def, ignore.case = TRUE) == TRUE)
    return("Iron")
  else
    return("Other")
}
## Function to define a list of key metabolism
fx_setKOmetabolism <- function(def) {
  #oxidation, reduction
  #fixation, hydrolysis, production, utilization, transport
  #sulfane is an oxidative process, as well as dehydrogenases
  if (grepl("oxida|oxidi|hydrog|hydrox|sulfane", def, ignore.case = TRUE) == TRUE)
    return("Oxidation")
  if (grepl("reduc", def, ignore.case = TRUE) == TRUE)
    return("Reduction")
  if (grepl("utili", def, ignore.case = TRUE) == TRUE)
    return("Utilization")
  if (grepl("hydrolase", def, ignore.case = TRUE) == TRUE)
    return("Hydrolysis")
  if (grepl("transferase", def, ignore.case = TRUE) == TRUE)
    return("Transferase")
  if (grepl("transport|binding", def, ignore.case = TRUE) == TRUE)
    return("Transport")
  if (grepl("resistance", def, ignore.case = TRUE) == TRUE)
    return("Resistance")
  if (grepl("carrier|chaperone", def, ignore.case = TRUE) == TRUE)
    return("Carrier")
  if (grepl("transcriptional|regulator", def, ignore.case = TRUE) == TRUE)
    return("Regulator")
  if (grepl("lyase", def, ignore.case = TRUE) == TRUE)
    return("Lyase")
  else
    return("Other")
}
## Function that returns the ecosystem given a MAG
fx_getMAGecosystem <- function(mag) {
  if (grepl("spades", mag, ignore.case = TRUE) == TRUE) {
    return("Biomining")
  }
  else {
    mags_df <- mags_metadata_bac %>%
      dplyr::mutate(bin = paste0(mag_name, ".", mag_id)) %>%
      dplyr::filter(bin == mag) %>%
      dplyr::select(ecosystem) %>%
      as.character()
  }
}
## END FUNCTIONS ##

# Save pathways and metabolism to dataframe
kegg_ko_paths <- vector()
kegg_ko_metab <- vector()
for (i in 1:nrow(kegg_paths)) {
	kegg_ko_paths <- append(kegg_ko_paths, fx_setKOpathways(kegg_paths$Definition[i]))
	kegg_ko_metab <- append(kegg_ko_metab, fx_setKOmetabolism(kegg_paths$Definition[i]))
}
kegg_paths['Pathway'] <- kegg_ko_paths
kegg_paths['Metabolism'] <- kegg_ko_metab

# Filter matrix to select key pathways
kegg_rows <- rowSums(kegg_annot)
#summary(kegg_rows)
#table(kegg_rows)
kegg_filt <- kegg_annot[row.names(kegg_annot) %in% kegg_paths$KO, ]

# Normalize filtered rows by z-score
kegg_filt_norm <- apply(kegg_filt, 1,
                        norm <- function(x){return (x - mean(x)) / sd(x)}) %>% t()

#(plot test)
heatmap(as.matrix(kegg_filt))
heatmap(kegg_filt_norm)

# Plot using ComplexHeatmap
kegg_hm_colors <- colorRamp2(c(-2, 0, 2), c("#5073AD", "white", "#D33D35"))

# Annotate rows: KO labels
#with high number of hits: row.names(kegg_filt[rowSums(kegg_filt) > 100, ])
# KOs with low variance (1/2 data %>% sort())
#kegg_hm_rows_lab <- row.names(kegg_filt[apply(kegg_filt_norm, 1, var) < 0.00015, ])
#kegg_hm_rows_at <- which(row.names(kegg_filt) %in% kegg_hm_rows_lab)
#kegg_hm_rows_kos <- HeatmapAnnotation(KOs = anno_mark(at = kegg_hm_rows_at,
#                                                      labels = kegg_hm_rows_lab,
#                                                      labels_gp = gpar(fontsize = 6)),
#                                      which = "row")
# Row names in the heatmap are not sorted!
#row.names(kegg_filt)[row_order(kegg_heatmap)]

# Annotate rows: KO pathways
kegg_hm_rows_path <- c("Iron" = "#666666", "Sulfur" = "#DDAD3B", "Copper" = "#CA6627")
kegg_hm_rows_path <- list(Def = kegg_paths$Pathway, Pathways = kegg_hm_rows_path)

kegg_hm_rows_ann_path <- HeatmapAnnotation(Pathways = kegg_paths$Pathway,
                                           col = kegg_hm_rows_path, border = FALSE,
                                           show_annotation_name = FALSE,
                                           show_legend = TRUE, which = "row")

# Annotate rows: KO metabolism
kegg_hm_rows_metab <- setNames(brewer.pal(length(unique(kegg_paths$Metabolism)), "Paired"), unique(kegg_paths$Metabolism)) #Set3
kegg_hm_rows_metab <- sample(kegg_hm_rows_metab)
kegg_hm_rows_metab <- list(Met = kegg_paths$Metabolism, Metabolism = kegg_hm_rows_metab)

kegg_hm_rows_ann_metab <- HeatmapAnnotation(Metabolism = kegg_paths$Metabolism,
                                            col = kegg_hm_rows_metab, border = FALSE,
                                            show_annotation_name = FALSE,
                                            show_legend = TRUE, which = "row")

# Annotate cols: separate biomining MAGs from other soil SMAG
kegg_mags <- colnames(kegg_filt) %>% as.data.frame()
colnames(kegg_mags) <- c("mag")

kegg_mags_eco <- vector()
for (i in 1:nrow(kegg_mags)) {
	kegg_mags_eco <- append(kegg_mags_eco, fx_getMAGecosystem(kegg_mags$mag[i]))
}
kegg_mags['ecosystem'] <- kegg_mags_eco
#table(kegg_mags$ecosystem)

kegg_hm_cols_mags <- setNames(brewer.pal(length(unique(kegg_mags$ecosystem)), "Paired"), unique(kegg_mags$ecosystem))
kegg_hm_cols_mags <- sample(kegg_hm_cols_mags)
kegg_hm_cols_mags <- list(Eco = kegg_mags$ecosystem, Ecosystem = kegg_hm_cols_mags)

kegg_hm_cols_ann_mags <- HeatmapAnnotation(Ecosystem = kegg_mags$ecosystem,
                                           col = kegg_hm_cols_mags, border = FALSE,
                                           show_annotation_name = FALSE,
                                           show_legend = TRUE, which = "column")

# Heatmap with annotations
pdf("plot_annot_heatmap_kofam.pdf", height = 9, width = 9)
kegg_heatmap <- Heatmap(kegg_filt_norm, name = "Abundance",
                        row_title = "KEGG KOs", column_title = "Ecosystem MAGs",
                        show_row_names = FALSE, show_column_names = FALSE,
                        col = kegg_hm_colors, border = FALSE,
                        row_split = kegg_paths$Pathway,
                        column_split = kegg_mags$ecosystem,
                        right_annotation = kegg_hm_rows_ann_metab,
                        left_annotation = kegg_hm_rows_ann_path,
                        top_annotation = kegg_hm_cols_ann_mags)
draw(kegg_heatmap, merge_legend = TRUE)
while (!is.null(dev.list())) dev.off()

### 2. Read output file from Anvio Kofam
anvio_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/results/anvio_kofam_hits.tsv", sep = "\t")

# Save KO definitions and set row names
#anvio_kos <- data.frame(KO = anvio_annot$ko,
#                        Modules = anvio_annot$modules_with_ko,
#                        Definition = anvio_annot$ko_definition)
#anvio_annot <- anvio_annot %>% select(-c("modules_with_ko", "ko_definition"))
#anvio_annot <- data.frame(anvio_annot, row.names = 1)
#
## Set definition for selected pathways
#anvio_paths <- anvio_kos %>%
#  filter(grepl("iron|sulfur|copper", Definition, ignore.case = TRUE) == TRUE) %>%
#  select(KO, Definition)
##which(anvio_paths$KO %in% kegg_paths$KO)
#
#anvio_def_paths <- vector()
#for (i in 1:nrow(anvio_paths)) {
#	anvio_def_paths <- append(anvio_def_paths, setKODefPathway(anvio_paths$Definition[i]))
#}
#anvio_paths['Pathway'] <- anvio_def_paths
#
## Filter matrix to select key pathways
## Preprocess matrix.. (filter KOs per row)
#anvio_rows <- rowSums(anvio_annot)
#summary(anvio_rows)
#table(anvio_rows)
#anvio_filt <- anvio_annot[row.names(anvio_annot) %in% anvio_paths$KO, ]
```

### PCA

```{r annotpca}
library(ggbiplot)
library(PCAtools)

### 1. KOfamScan

# Note: relevant cycles (iron, sulfur, copper) have low representation in row sums,
# so during filtration we need to consider them within the range of values.
# For example, rows with low variances could be kept but they are also too many..

# Preprocess matrix
kegg_pca_mat <- kegg_annot[kegg_rows > 1 & kegg_rows < round(summary(kegg_rows)[4], 0), ]

# Merge KO definitions
kegg_pca_kos <- data.frame(KO = row.names(kegg_pca_mat))
kegg_pca_kos <- left_join(kegg_pca_kos, kegg_kos, by = c("KO"))

# Pathways of interest to dataframe
kegg_pca_def <- vector()
for (i in 1:nrow(kegg_pca_kos)) {
	kegg_pca_def <- append(kegg_pca_def, fx_setKOpathways(kegg_pca_kos$Definition[i]))
}
kegg_pca_kos['Pathway'] <- kegg_pca_def

# Normalize data to z-score
kegg_pca_mat <- apply(kegg_pca_mat, 1,
                      norm_pca <- function(x){return (x - mean(x)) / sd(x)}) %>% t()

# PCA based on SVD
set.seed(12345)
kegg_pca <- prcomp(kegg_pca_mat, center = FALSE, scale. = FALSE)

# Print loadings (eigenvectors)
kegg_pca$rotation
kegg_pca$sdev

# Calculate variance and variance explained by each PC
kegg_pca_var <- kegg_pca$sdev ^ 2
kegg_pca_var_exp <- kegg_pca_var / sum(kegg_pca_var)

# Save as PCA object file
if (file.exists("kegg_pca.rds") == FALSE) {
  saveRDS(kegg_pca, "kegg_pca.rds")
}

# Screeplot
kegg_pca_scree_df <- data.frame(PC = 1:ncol(kegg_pca$rotation), Var = kegg_pca_var_exp)

kegg_pca_screeplot <- ggplot(kegg_pca_scree_df, aes(x = PC, y = Var)) +
  geom_bar(stat = "identity", fill = "#9cb8ca") +
  xlim(0, 50) + ylim(0, 0.06) +
  labs(title = "KEGG KOs (KOfamScan) Scree Plot", x = "PC", y = "Variance Explained") +
  theme_minimal()
ggsave("plot_annot_pca_kofam_screeplot.pdf", kegg_pca_screeplot, height = 4, width = 6)

# Biplot
kegg_pca_colors <- c("Copper" = "#CA6627", "Iron" = "#666666",
                     "Sulfur" = "#DDAD3B", "Other" = "#e6f0f6")

kegg_pca_biplot <- ggbiplot(kegg_pca, choices = c(1, 2), groups = kegg_pca_kos$Pathway,
                            obs.scale = 1, var.scale = 1, var.axes = FALSE,
                            labels.size = 5, labels = "*", #alpha = 0,
                            ellipse = TRUE, circle = FALSE) +
  scale_color_manual(name = "Pathway", values = kegg_pca_colors,
                     guide = guide_legend(override.aes = aes(label = ""))) +
  geom_vline(xintercept = 0, linetype = 3) +
  geom_hline(yintercept = 0, linetype = 3) +
  ylim(-8, 5) +
  #geom_point(aes(colour = kegg_pca_kos$Pathway), size = 1.5) +
  theme(legend.direction = "horizontal", legend.position = "right") +
  theme_bw()
ggsave("plot_annot_pca_kofam_biplot.pdf", kegg_pca_biplot, height = 9, width = 9)

# Set colors for ecosystem (standarize**)
kegg_colors_eco <- setNames(brewer.pal(length(unique(kegg_mags$ecosystem)), "Paired"), unique(kegg_mags$ecosystem))
kegg_colors_eco <- sample(kegg_colors_eco)

### PCAtools ###
set.seed(11122)
kegg_pca_obj <- pca(kegg_pca_mat, center = FALSE, scale = FALSE,
                    metadata = kegg_mags %>% data.frame(row.names = 1))

# Save PCA object
if (file.exists("kegg_pcatools.rds") == FALSE) {
  saveRDS(kegg_pca_obj, "kegg_pcatools.rds")
}

# Convert to a PCAtools object
# Same result, different y-scale!
#kegg_pca_obj <- list(rotated = data.frame(kegg_pca$rotation),
#                     loadings = data.frame(kegg_pca$x),
#                     variance = (kegg_pca_var_exp * 100) %>% set_names(colnames(kegg_pca$x)),
#                     sdev = kegg_pca$sdev,
#                     metadata = kegg_mags,
#                     xvars = kegg_pca_kos$KO,
#                     yvars = colnames(kegg_pca_mat),
#                     components = colnames(kegg_pca$x))
#row.names(kegg_pca_obj$rotated) <- kegg_pca_obj$yvars
#row.names(kegg_pca_obj$loadings) <- kegg_pca_obj$xvars
#names(kegg_pca_obj$variance) <- kegg_pca_obj$components
#class(kegg_pca_obj) <- "pca"

# Biplot
kegg_pca_obj_biplot <- biplot(kegg_pca_obj, showLoadings = FALSE, sizeLoadingsNames = FALSE,
                              drawConnectors = FALSE,  drawConnectorsLoadings = FALSE,
                              pointSize = 2, labSize = 1, legendPosition = "right",
                              colLegendTitle = "Ecosystem", colby = "ecosystem", 
                              colkey = kegg_colors_eco, ellipse = TRUE, ylim = c(-14, 12))
ggsave("plot_annot_pcatools_kofam_biplot.pdf", kegg_pca_obj_biplot, height = 9, width = 10)

# Pairsplot
kegg_pca_obj_pairs <- pairsplot(kegg_pca_obj, colby = "ecosystem",
                                colkey = kegg_colors_eco, pointSize = 1,
                                components = getComponents(kegg_pca_obj, c(1:4)))
ggsave("plot_annot_pcatools_kofam_pairsplot.pdf", kegg_pca_obj_pairs, height = 7, width = 10)
```

### Bubble plot

```{r annotbub}
#split KOs per copper, iron and sulfur
#for each KO, set all ecosystems to definition
#split columns per mag ecosystems
#print in row labels KO name + definition

### 1. KOfamScan

# Get unique ecosystems
kegg_bubble_eco <- unique(kegg_mags$ecosystem) %>% as.character()

# Prepare dataframe
kegg_bubble <- row.names(kegg_filt) %>%
  as.data.frame() %>%
  slice(rep(1:n(), each = length(kegg_bubble_eco)))
colnames(kegg_bubble) <- c("KO")
kegg_bubble['ecosystem'] <- c("")
kegg_bubble$ecosystem <- rep(kegg_bubble_eco, nrow(kegg_filt))
#optimize with pivot_longer()?
#kegg_bubble['count'] <- sample(100, size = nrow(kegg_bubble), replace = TRUE) #plot test

## Function to count matching KOs in all bins for each ecosystem
fx_countKOsbyeco <- function(ko) {
  #given a KO, retrieve all bins in the filtered matrix
  row <- kegg_filt[row.names(kegg_filt) == ko, ]
  #vectors to populate
  eco <- vector()
  sum <- vector()
  for (j in 1:ncol(row)) {
    #get the ecosystem for each bin (column) and sum uniques
    e <- kegg_mags[kegg_mags$mag == colnames(kegg_filt)[j], ] %>%
      select(ecosystem) %>%
      as.character()
    eco <- append(eco, e)
    sum <- append(sum, row[1, j])
  }
  #vectors to dataframe
  df <- data.frame(Ecosystem = eco, Counts = sum)
  #merge ecosystems for counting and return
  df <- aggregate(df$Counts, list(df$Ecosystem), FUN = sum)
  colnames(df)[1] <- "ecosystem"
  colnames(df)[2] <- "count"
  return(df)
}

# Count KOs
kegg_bubble_counts <- tibble(ecosystem = character(), count = numeric())
for (i in seq(1, nrow(kegg_bubble), length(kegg_bubble_eco))) {
  #loop step of 10 KOs
  #print(sprintf("%i -> KO: %s", i, kegg_bubble[i, 1]))
  #current KO
  kos_df <- fx_countKOsbyeco(kegg_bubble[i, 1])
  #merge ecosystem counts sorted from our stored list
  eco_df <- kegg_bubble_eco %>% as.data.frame()
  colnames(eco_df) <- c("ecosystem")
  counts_df <- left_join(eco_df, kos_df, by = c("ecosystem"))
  kegg_bubble_counts <- rbind(kegg_bubble_counts, counts_df)
}#runtime ~5 min!
#clean memory and add column of counts
rm(kos_df, eco_df, counts_df)
kegg_bubble <- cbind(kegg_bubble, counts = kegg_bubble_counts[, 2])

# Subset to cycles of interest (3):
# 1. get definitions and metabolism from kegg_paths
# 2. merge df with counts per ecosystem
# 3. remove biomining KOs with low counts (ie. sulfur: <= 5)
# 4. subset df with kept KOs
#todo: split definition in two columns by " [EC:"
kegg_bubble_copper <- kegg_paths %>%
  filter(Pathway == "Copper") %>%
  select(KO, Definition, Metabolism)
kegg_bubble_copper <- left_join(kegg_bubble_copper, kegg_bubble, by = c("KO"))
kegg_bubble_keep <- kegg_bubble_copper %>%
  filter(ecosystem == "Biomining" & counts > 1) %>%
  select(KO)
kegg_bubble_copper <- kegg_bubble_copper[kegg_bubble_copper$KO %in% kegg_bubble_keep$KO, ]

kegg_bubble_iron <- kegg_paths %>%
  filter(Pathway == "Iron") %>%
  select(KO, Definition, Metabolism)
kegg_bubble_iron <- left_join(kegg_bubble_iron, kegg_bubble, by = c("KO"))
kegg_bubble_keep <- kegg_bubble_iron %>%
  filter(ecosystem == "Biomining" & counts > 2) %>%
  select(KO)
kegg_bubble_iron <- kegg_bubble_iron[kegg_bubble_iron$KO %in% kegg_bubble_keep$KO, ] %>%
  filter(!KO == "K02014") #remove outlier with counts > 500

kegg_bubble_sulfur <- kegg_paths %>%
  filter(Pathway == "Sulfur") %>%
  select(KO, Definition, Metabolism)
kegg_bubble_sulfur <- left_join(kegg_bubble_sulfur, kegg_bubble, by = c("KO"))
kegg_bubble_keep <- kegg_bubble_sulfur %>%
  filter(ecosystem == "Biomining" & counts >= 5) %>%
  select(KO)
kegg_bubble_sulfur <- kegg_bubble_sulfur[kegg_bubble_sulfur$KO %in% kegg_bubble_keep$KO, ]

# Bubble plot
kegg_bubble_plot <- ggplot(kegg_bubble_iron,
                           aes(x = ecosystem, y = Definition,
                               size = counts, color = ecosystem)) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = kegg_colors_eco, guide = "none") +
  labs(title = "", x = "Ecosystem", y = "KEGG KO Definitions") +
  theme_classic() +
  theme(legend.title = element_text(size = 10),
        legend.position = "right",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
ggsave("plot_annot_bubble_kofam_copper.pdf", kegg_bubble_plot, height = 7, width = 9)
ggsave("plot_annot_bubble_kofam_iron.pdf", kegg_bubble_plot, height = 7, width = 7)
ggsave("plot_annot_bubble_kofam_sulfur.pdf", kegg_bubble_plot, height = 9, width = 9)
```

## SNPs

```{r snps}
library(VariantAnnotation)
library(GenomicRanges)
library(vcfR)

# Load VCF resulting from BCFTools
vcf_file <- readVcf("../../snps/S15.vcf")

# Header information
header(vcf_file)
samples(header(vcf_file))
geno(header(vcf_file))
info(header(vcf_file))

# Genomic positions (CHROM, POS, ID)
seqnames(vcf_file)
head(rowRanges(vcf_file), 3)
ref(vcf_file)[1:5]
qual(vcf_file)[1:5]
alt(vcf_file)[1:5]

# Genotype data (FORMAT)
geno(vcf_file)
dim(geno(vcf_file)$GT)

# Info data
info(vcf_file)[1:4, 1:5]

# Plot
#vcf_plot <- ggplot(data=d.fig5, aes(x=Output, y=bcftools, fill=Output)) +
#  geom_bar(aes(alpha=0.9), stat="identity", color="black", position=position_dodge()) +
#  theme_bw() + ylab("Similarity (%)") + xlab("") + expand_limits(y=c(0, 105)) + 
#  facet_grid(rows = vars(Label_bcftools), cols = vars(Assembler)) +
#  geom_text(aes(label=mean_bcftools), vjust=-0.3, color="black", size=2.5)
#ggsave("plot_annot_bubble_kofam_copper.pdf", kegg_bubble_plot, height = 7, width = 9)

# Read VCF file with vcfR package
vcf <- read.vcfR("../../snps/S15.vcf")

# Parse DP from the gt region
#https://onlinelibrary.wiley.com/doi/10.1111/1755-0998.12549
vcf_dp <- extract.gt(vcf, element = "DP", return.alleles = FALSE)
vcf_dpf <- melt(vcf_dp, varnames = c("Index", "Sample"), value.name = "Depth", na.rm = TRUE)

# Render a violin plot

# Create a chrom plot
vcf_chrom <- create.chromR(vcf, name = "S15 NODE_1") #subsetted to the first chromosome
vcf_chrom <- proc.chromR(vcf_chrom)
chromoqc(vcf_chrom)
```

### 2024-03-11 R/paper_filter_metadata.Rmd
Added datasets and R scripts. Updated until fig. 1

---
title: "Systemix Profiling Article: Filtering MAG-SMAG Metadata"
author: "Moises A. Rojas"
date: "2024-01-17"
output: html_document
---

```{r loadmd}
library(dplyr)

# Load Biomining MAGs and SMAG bacterial metadata
mags_md <- read.csv("4-mags-smag/mags_metadata_bac.tsv", sep = "\t",
                    header = TRUE, stringsAsFactors = FALSE)

# Choose a set of bins:
# 1. select other ecosystems keeping only high-quality bins;
# 2. sort them by completeness and group them by ecosystem
# 3. remove duplicates based on classification (after exploring its variability)
# 4. select top bins for each ecosystem
mags_md_sel <- mags_md %>%
  filter(ecosystem != "Biomining" & contamination <= 5) %>%
  arrange(desc(completeness)) %>%
  group_by(ecosystem) %>%
  select(mag_name, mag_id, completeness, contamination, ecosystem, phylum, class, order, family, genus, species) %>%
  distinct(order, family, genus, species, .keep_all = TRUE) %>%
  slice(1:70)

# Merge MAG name and id
mags_md_sel <- mags_md_sel %>%
  mutate(bin = paste0(mag_name, ".", mag_id)) %>%
  relocate(bin, .before = mag_name) %>%
  select(-c("mag_name", "mag_id"))

#explore selection..
#mags_md_sel %>% filter(species != "") %>% nrow()

# Save list of bins to file
write.table(mags_md_sel$bin, "4-mags-smag/smag_filtered_bins.tsv",
            quote = FALSE, row.names = FALSE, col.names = FALSE)

# Clean memory
rm(mags_md, mags_md_sel)
```

### 2024-03-11 R/parsers/checkm_bins.R

@@ -1,0 +1,67 @@
+#!/usr/bin/env Rscript
+
+#---------------------------------------------------------
+# Systemix MAGs: Get the QC from a list of -refined- bins
+# 
+# Moises A. Rojas 06/11/2023 - mod. 06/03/2024
+#---------------------------------------------------------
+
+# Set working dir
+setwd("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/2-statistics/")
+
+# Read bin list
+bins <- read.csv("bins_refined_list.tsv", sep = "\t", stringsAsFactors = FALSE)
+selected_bins <- bins$bin
+
+# QC parameters to keep
+completeness  <- vector()
+contamination <- vector()
+strain_het    <- vector()
+genome_size   <- vector()
+contigs       <- vector()
+n50_contigs   <- vector()
+gc            <- vector()
+pred_genes    <- vector()
+
+# Read the QC file from CheckM
+checkm_qc <- read.csv("checkm_qa.tsv", sep = "\t", stringsAsFactors = FALSE)
+
+# Save attributes for each bin
+checkm_bin <- checkm_qc$bin
+checkm_com <- checkm_qc$completeness
+checkm_con <- checkm_qc$contamination
+checkm_st  <- checkm_qc$strain_heterogeneity
+checkm_gs  <- checkm_qc$genome_size
+checkm_nco <- checkm_qc$contigs
+checkm_n50 <- checkm_qc$N50_contigs
+checkm_gc  <- checkm_qc$gc
+checkm_pg  <- checkm_qc$predicted_genes
+
+# Match positions for each bin, saving attributes
+for (i in 1:length(selected_bins)) { #428
+	for (j in 1:length(checkm_bin)) { #1840
+		if (selected_bins[i] == checkm_bin[j]) {
+			#print(j) #pos
+			completeness  <- append(completeness,  checkm_com[j])
+			contamination <- append(contamination, checkm_con[j])
+			strain_het    <- append(strain_het,    checkm_st[j])
+			genome_size   <- append(genome_size,   checkm_gs[j])
+			contigs       <- append(contigs,       checkm_nco[j])
+			n50_contigs   <- append(n50_contigs,   checkm_n50[j])
+			gc            <- append(gc,            checkm_gc[j])
+			pred_genes    <- append(pred_genes,    checkm_pg[j])
+		}
+	}
+}
+
+# Create output dataframe and save to file
[…]

### 2024-03-11 R/parsers/get_gtdb_taxa.R

@@ -1,0 +1,69 @@
+#!/usr/bin/env Rscript
+
+#-----------------------------------------------
+# Populate the GTDB-Tk of MAGs from a reference
+# Moises A. Rojas 13/10/2023
+#-----------------------------------------------
+
+setwd("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/1-taxonomy")
+
+# Read the list of bins (last nf-core/mag run: mags-all/)
+bins_all <- read.csv("bins_refined_list.tsv", sep = "\t", stringsAsFactors = FALSE)
+head(bins_all)
+
+bins_genome <- bins_all$bin
+
+# Read the GTDB-Tk classification file
+gtdb_file <- read.csv("gtdbtk_summary.tsv", sep = "\t", stringsAsFactors = FALSE)
+head(gtdb_file)
+
+# Save attributes
+gtdb_genome <- gtdb_file$user_genome
+gtdb_taxa   <- gtdb_file$classification
+
+# Set output attributes
+bin     <- vector()
+taxa    <- vector()
+kingdom <- vector()
+phylum  <- vector()
+class   <- vector()
+order   <- vector()
+family  <- vector()
+genus   <- vector()
+species <- vector()
+
+# Parse taxonomy array:
+# d__Bacteria;p__Nitrospirota_A;c__Leptospirillia;o__Leptospirillales;f__Leptospirillaceae;g__;s__
+taxa_split <- strsplit(gtdb_taxa, split = ";", fixed = TRUE)
+
+# Match positions for each bin
+# Then, save taxa attributes
+for (i in 1:length(bins_genome)) { # 428
+	for (j in 1:length(gtdb_genome)) { # 377
+		if (bins_genome[i] == gtdb_genome[j]) {
+			#print(j) #pos
+			bin     <- append(bin,     gtdb_genome[[j]])
+			taxa    <- append(taxa,    gtdb_taxa[[j]])
+			kingdom <- append(kingdom, taxa_split[[j]][1])
+			phylum  <- append(phylum,  taxa_split[[j]][2])
+			class   <- append(class,   taxa_split[[j]][3])
+			order   <- append(order,   taxa_split[[j]][4])
+			family  <- append(family,  taxa_split[[j]][5])
+			genus   <- append(genus,   taxa_split[[j]][6])
+			species <- append(species, taxa_split[[j]][7])
+		}
+	}
+}
+
+# Create output dataframe and save to file
[…]

### 2024-03-18 R/paper.Rmd
updated to fig. 2

@@
 suppressWarnings(library(colorspace))
-suppressWarnings(library(patchwork))
+suppressWarnings(library(patchwork)) #side-by-side plotting
 ```
@@
 
-plot_md_top_phyla <- ggplot(mags_md_top_phy, aes(x = reorder(phy_top, freq), y = freq)) +
+plot_depth_top_phyla <- ggplot(mags_md_top_phy, aes(x = reorder(phy_top, freq), y = freq)) +
   geom_bar(stat = "identity", fill = "#459682") +
   geom_text(aes(label = freq), vjust = 0.5, hjust = -0.5, size = 3) +
-  labs(x = "Phylum", y = "Count") +
+  labs(x = "Phylum", y = "Number") +
   scale_y_continuous(limits = c(0, 10000)) +
@@
   coord_flip()
-ggsave("plot_md_top_phyla.pdf", plot_md_top_phyla, height = 7, width = 6)
+ggsave("plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)
 
@@
   geom_bar(aes(x = species, fill = phylum)) +
-  labs(x = "Known biomining species", y = "Count") +
+  labs(x = "", y = "Number of known species") +
   scale_fill_manual(values = "#DD7E8B") +
@@
   geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.5, size = 3) +
-  labs(x = "Ecosystem", y = "Number of kSGBs / Number of MAGs") +
+  labs(x = "Ecosystem", y = "Number of kSGBs / MAGs") +
   scale_y_continuous(limits = c(0, 10)) +
@@
   scale_y_continuous(limits = c(0, 25)) +
-  labs(x = "Phylum", y = "Number of biomining species") +
+  labs(x = "Phylum", y = "Biomining species") +
   theme_classic() +
@@
 plot_tree_s15_phyla <- plot_tree_s15 +
-  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1, show.legend = FALSE) +
-  geom_tiplab(aes(label = phylum, color = phylum),
-              size = 1.5, hjust = -0.1, show.legend = FALSE) +
-  scale_colour_discrete(breaks = c('c', 'd'), na.value = NA) +
-  scale_colour_manual(values = tree_colors, na.translate = FALSE) +
+  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
+  geom_tiplab(aes(label = phylum, color = phylum), size = 1.75, hjust = -0.1) +
+  scale_colour_discrete(breaks = c("c", "d"), na.value = NA) +
+  scale_colour_manual(values = tree_colors, na.translate = FALSE, guide = "none") +
+  #guides(color = guide_legend(title = "Phyla", ncol = 1, order = 1)) +
   theme()
-  #+guides(color = guide_legend(title = "Phylum", ncol = 1, order = 1))
 ggsave("plot_biom_MAGs_tree_phyla.pdf", plot_tree_s15_phyla, height = 5, width = 7)
@@
 # Plot tree adding QC as multiple layers
-plot_tree_s15_md <- plot_tree_s15_phyla + new_scale_fill() + new_scale_color() +
+plot_tree_s15_md <- plot_tree_s15_phyla +
+  new_scale_fill() + new_scale_color() +
   geom_fruit(data = tree_qc_df, geom = geom_bar,
              mapping = aes(y = Bin, x = Completeness, fill = Completeness),
-             orientation = "y", stat = "identity", position = "auto",
-             grid.params = list(linetype = 1, size = 0.1),
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3),
-             pwidth = 0.15, offset = 0.35) +
-  scale_fill_gradient(low = "#1F4060", high = "#4CA8F6",
-                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
-                                           reverse = TRUE, order = 1)) +
+             orientation = "y", stat = "identity", position = "auto", width = 0.85,
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.6,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3,
+                                title = "Complet.", title.size = 1.6,
+                                title.height = 0.035, title.angle = 30)) +
+  scale_fill_gradient(low = "#1F4060", high = "#4CA8F6", guide = "none") +
+                      #guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
+                      #                     reverse = TRUE, order = 1)) +
+  ###Error in scale$guide == "none"
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Contamination, fill = Contamination),
-             orientation = "y", stat = "identity", position = "auto",
-             grid.params = list(linetype = 1, size = 0.1),
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1),
-             pwidth = 0.15, offset = 0.1) +
-  scale_fill_gradient(low = "#6A60A8", high = "#C4C2DE",
-                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
-                                           reverse = TRUE, order = 2)) +
+             orientation = "y", stat = "identity", position = "auto", width = 0.85,
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1,
+                                title = "Contam.", title.size = 1.6,
+                                title.height = 0.035, title.angle = 30)) +
+  scale_fill_gradient(low = "#6A60A8", high = "#C4C2DE", guide = "none") +
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Strain_Het, fill = Strain_Het),
-             orientation = "y", stat = "identity", position = "auto",
-             grid.params = list(linetype = 1, size = 0.1),
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3),
-             pwidth = 0.15, offset = 0.1) +
-  scale_fill_gradient(name = "Strain Heterogeneity",
-                      low = "#C9543F", high = "#E9bAB2",
-                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
-                                           reverse = TRUE, order = 3)) +
+             orientation = "y", stat = "identity", position = "auto", width = 0.85,
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3,
+                                title = "Strain Het.", title.size = 1.6,
+                                title.height = 0.035, title.angle = 30)) +
+  scale_fill_gradient(low = "#C9543F", high = "#E9BAB2",
+                      name = "Strain Heterogeneity", guide = "none") +
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Pred_Genes, fill = Pred_Genes),
-             orientation = "y", stat = "identity", position = "auto",
-             grid.params = list(linetype = 1, size = 0.1),
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1),
-             pwidth = 0.15, offset = 0.1) +
-  scale_fill_gradient(name = "Predicted Genes",
-                      low = "#279854", high = "#A9DEDE",
-                      guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
-                                           reverse = TRUE, order = 4)) +
-  theme(legend.title = element_text(size = 8),
-        legend.text = element_text(size = 7),
-        legend.key.height = unit(3.5, units = "mm"))
-ggsave("plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 10)
+             orientation = "y", stat = "identity", position = "auto", width = 0.85,
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1,
+                                title = "Pred. genes", title.size = 1.6,
+                                title.height = 0.035, title.angle = 30)) +
+  scale_fill_gradient(low = "#279854", high = "#A9DEDE",
+                      name = "Predicted Genes", guide = "none") +
+  theme()
+  #theme(legend.title = element_text(size = 8),
+  #      legend.text = element_text(size = 7),
+  #      legend.key.height = unit(3.5, units = "mm"),
+  #      legend.key.width = unit(2, units = "mm"))
+ggsave("plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 7) #x9
 ```
 
-## Figure 1
+## Figure 1 & supplementary 1
 
 ```{r fig1}
-# Merge generated figures using patchwork
-fig1_a <- {
-  (plot_tree_s15_md | plot_depth_spec_eco)
-  } + plot_layout(ncol = 2, widths = c(1, 0.5))
-ggsave("figures/fig1_a.pdf", fig1_a, height = 7, width = 14)
-
-fig1_b <- {
-  (plot_mags_qc / plot_depth_sgbs_biom) |
+# Merge plots using patchwork
+fig_1a <- { plot_tree_s15_md | plot_depth_spec_eco +
+  plot_layout(ncol = 2, widths = c(1, 0.5))
+  } + plot_annotation(tag_levels = list(c("a", "e"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("fig_1a.pdf", fig_1a, height = 7, width = 14)
+
+fig_1b <- { plot_mags_qc / plot_depth_sgbs_biom |
     plot_depth_sgbs_both | plot_depth_kggbs_biom
-  } + plot_layout(ncol = 3, widths = c(0.5, 2, 0.75))
-ggsave("figures/fig1_b.pdf", fig1_b, height = 7, width = 14)
-
-##fig1
-fig1 <- (fig1_a / fig1_b) +
-  plot_annotation(tag_levels = list(c("a", "e", "b", "c", "d", "", "f")))
-ggsave("figures/fig1.pdf", fig1, height = 14, width = 14)
-rm(fig1_a, fig1_b)
-
-##fig2
+  } + plot_layout(ncol = 3, widths = c(0.5, 2, 0.75)) +
+  plot_annotation(tag_levels = list(c("b", "c", "d", "", "f"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("fig_1b.pdf", fig_1b, height = 7, width = 14)
+
+fig_1 <- (fig_1a / fig_1b) +
+  plot_annotation(tag_levels = list(c("a", "e", "b", "c", "d", "", "f"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("fig_1.pdf", fig_1, height = 14, width = 14)
+rm(fig_1a, fig_1b)
+#ggsave("fig_1.png", fig_1, dpi = 300, dev = "png", height = 14, width = 14, units = "cm")
+
+##fig sup 1
+fig_s1 <- { plot_depth_top_phyla | plot_depth_spec } +
+  plot_layout(ncol = 2, widths = c(0.5, 1)) &
+  theme(axis.title.x = element_text(margin = margin(t = -75, unit = "pt")))
+fig_s1 <- { fig_s1 /
+    { (plot_spacer() + plot_depth_usgbs + plot_spacer()) +
+        plot_layout(ncol = 3, widths = c(0.5, 12, 2)) } #centered
+  } + plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("fig_s1.pdf", fig_s1, height = 14, width = 14)
+#ggsave("fig_s1.png", fig_s1, dpi = 300, dev = "png", height = 14, width = 14, units = "cm")
 ```
 
-## Functional Annotation: KEGG KOs from KOfamScan and Anvio Kofam
-
-### Heatmaps
+## Functional Annotation: KEGG KOs from KOfamScan
+
+### Heatmap
 
@@
 
-### 1. Read output file from KOfamScan hits
+# Read output file from KOfamScan
 kegg_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/merge_hits/kofamscan_hits.tsv", sep = "\t")
 
-# Remove uncharacterized and duplicates proteins
+# Remove uncharacterized and duplicated proteins
 kegg_annot <- kegg_annot %>%
-  filter(!grepl("uncharacterized", KO.definition) == TRUE) %>%
+  dplyr::filter(!grepl("uncharacterized", KO.definition) == TRUE) %>%
   distinct(KO.definition, .keep_all = TRUE)
 
-# Extract remaining KO definitions and set KOs as row names
+# Extract KO definitions and set KOs as row names
 kegg_kos <- data.frame(KO = kegg_annot$KO, Definition = kegg_annot$KO.definition)
 kegg_annot <- kegg_annot %>%
-  select(-c("KO.definition")) %>%
+  dplyr::select(-c("KO.definition")) %>%
   data.frame(row.names = 1)
@@
   #"soxr": SoxR 2Fe-2S cluster
-  filter(grepl("iron|sulf|soxr|copper", Definition, ignore.case = TRUE) == TRUE) %>%
-  select(KO, Definition)
+  dplyr::filter(grepl("iron|sulf|soxr|copper", Definition, ignore.case = TRUE) == TRUE) %>%
+  dplyr::select(KO, Definition)
 
 ## FUNCTIONS ##
-## Function that returns the pathway based on the definitions
+# Function that returns the pathway based on definitions
 fx_setKOpathways <- function(def) {
@@
 }
-## Function to define a list of key metabolism
+# Function to define types of metabolism
 fx_setKOmetabolism <- function(def) {
@@
 }
-## Function that returns the ecosystem given a MAG
+# Function that returns the ecosystem given a MAG
 fx_getMAGecosystem <- function(mag) {
@@
   else {
-    mags_df <- mags_metadata_bac %>%
+    mag_eco <- mags_metadata_bac %>%
       dplyr::mutate(bin = paste0(mag_name, ".", mag_id)) %>%
@@
       as.character()
+    return(mag_eco)
   }
@@
 
-# Save pathways and metabolism to dataframe
+# Populate both pathways and metabolism
 kegg_ko_paths <- vector()
@@
 
-# Filter matrix to select key pathways
-kegg_rows <- rowSums(kegg_annot)
-#summary(kegg_rows)
-#table(kegg_rows)
+# Keep only selected pathways from the annotation matrix
+# (this matrix will be used later for bubble plots)
 kegg_filt <- kegg_annot[row.names(kegg_annot) %in% kegg_paths$KO, ]
 
-# Normalize filtered rows by z-score
-kegg_filt_norm <- apply(kegg_filt, 1,
-                        norm <- function(x){return (x - mean(x)) / sd(x)}) %>% t()
-
-#(plot test)
-heatmap(as.matrix(kegg_filt))
-heatmap(kegg_filt_norm)
+# Row-wise variance, then filter both matrix and pathways
+kegg_filt_row_var <- apply(kegg_filt, 1, var)
+#save in another variable!**
+kegg_filt <- kegg_filt[kegg_filt_row_var >= summary(kegg_filt_row_var)[3], ] #median
+kegg_paths_sub <- kegg_paths[kegg_paths$KO %in% row.names(kegg_filt), ]
+
+# Normalize rows by z-score
+kegg_filt_norm <- apply(kegg_filt, 1, norm <- function(x){return (x - mean(x)) / sd(x)}) %>% t()
+
+#plot test
+#heatmap(as.matrix(kegg_filt))
+#heatmap(kegg_filt_norm)
 
@@
 kegg_hm_rows_path <- c("Iron" = "#666666", "Sulfur" = "#DDAD3B", "Copper" = "#CA6627")
-kegg_hm_rows_path <- list(Def = kegg_paths$Pathway, Pathways = kegg_hm_rows_path)
-
-kegg_hm_rows_ann_path <- HeatmapAnnotation(Pathways = kegg_paths$Pathway,
+kegg_hm_rows_path <- list(Def = kegg_paths_sub$Pathway, Pathways = kegg_hm_rows_path)
+
+kegg_hm_rows_ann_path <- HeatmapAnnotation(Pathways = kegg_paths_sub$Pathway,
                                            col = kegg_hm_rows_path, border = FALSE,
@@
 # Annotate rows: KO metabolism
-kegg_hm_rows_metab <- setNames(brewer.pal(length(unique(kegg_paths$Metabolism)), "Paired"), unique(kegg_paths$Metabolism)) #Set3
+kegg_hm_rows_metab <- setNames(brewer.pal(length(unique(kegg_paths_sub$Metabolism)), "Paired"), unique(kegg_paths_sub$Metabolism)) #Set3
 kegg_hm_rows_metab <- sample(kegg_hm_rows_metab)
-kegg_hm_rows_metab <- list(Met = kegg_paths$Metabolism, Metabolism = kegg_hm_rows_metab)
-
-kegg_hm_rows_ann_metab <- HeatmapAnnotation(Metabolism = kegg_paths$Metabolism,
+kegg_hm_rows_metab <- list(Met = kegg_paths_sub$Metabolism, Metabolism = kegg_hm_rows_metab)
+
+kegg_hm_rows_ann_metab <- HeatmapAnnotation(Metabolism = kegg_paths_sub$Metabolism,
                                             col = kegg_hm_rows_metab, border = FALSE,
@@
 
-# Annotate cols: separate biomining MAGs from other soil SMAG
+# Annotate cols: ecosystems
 kegg_mags <- colnames(kegg_filt) %>% as.data.frame()
@@
 kegg_mags['ecosystem'] <- kegg_mags_eco
+rm(kegg_mags_eco)
 #table(kegg_mags$ecosystem)
 
+#match with mags_colors_eco**
 kegg_hm_cols_mags <- setNames(brewer.pal(length(unique(kegg_mags$ecosystem)), "Paired"), unique(kegg_mags$ecosystem))
@@
                         col = kegg_hm_colors, border = FALSE,
-                        row_split = kegg_paths$Pathway,
+                        row_split = kegg_paths_sub$Pathway,
                         column_split = kegg_mags$ecosystem,
@@
 while (!is.null(dev.list())) dev.off()
-
-### 2. Read output file from Anvio Kofam
-anvio_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/results/anvio_kofam_hits.tsv", sep = "\t")
-
-# Save KO definitions and set row names
-#anvio_kos <- data.frame(KO = anvio_annot$ko,
-#                        Modules = anvio_annot$modules_with_ko,
-#                        Definition = anvio_annot$ko_definition)
-#anvio_annot <- anvio_annot %>% select(-c("modules_with_ko", "ko_definition"))
-#anvio_annot <- data.frame(anvio_annot, row.names = 1)
-#
-## Set definition for selected pathways
-#anvio_paths <- anvio_kos %>%
-#  filter(grepl("iron|sulfur|copper", Definition, ignore.case = TRUE) == TRUE) %>%
-#  select(KO, Definition)
-##which(anvio_paths$KO %in% kegg_paths$KO)
-#
-#anvio_def_paths <- vector()
-#for (i in 1:nrow(anvio_paths)) {
-#	anvio_def_paths <- append(anvio_def_paths, setKODefPathway(anvio_paths$Definition[i]))
-#}
-#anvio_paths['Pathway'] <- anvio_def_paths
-#
-## Filter matrix to select key pathways
-## Preprocess matrix.. (filter KOs per row)
-#anvio_rows <- rowSums(anvio_annot)
-#summary(anvio_rows)
-#table(anvio_rows)
-#anvio_filt <- anvio_annot[row.names(anvio_annot) %in% anvio_paths$KO, ]
 ```
@@
 
-### 1. KOfamScan
-
-# Note: relevant cycles (iron, sulfur, copper) have low representation in row sums,
+# NOTE: selected cycles (Fe, Cu, S) have low representation in row sums,
 # so during filtration we need to consider them within the range of values.
-# For example, rows with low variances could be kept but they are also too many..
-
-# Preprocess matrix
-kegg_pca_mat <- kegg_annot[kegg_rows > 1 & kegg_rows < round(summary(kegg_rows)[4], 0), ]
+# E.g., rows with low variances could be kept but they are also too many.
+#kegg_pca_annot_row_var <- apply(kegg_annot, 1, var)
+#kegg_pca_dat <- kegg_annot[kegg_pca_annot_row_var > summary(kegg_pca_annot_row_var)[3], ]
+kegg_pca_dat <- kegg_filt
 
 # Merge KO definitions
-kegg_pca_kos <- data.frame(KO = row.names(kegg_pca_mat))
+kegg_pca_kos <- data.frame(KO = row.names(kegg_pca_dat))
 kegg_pca_kos <- left_join(kegg_pca_kos, kegg_kos, by = c("KO"))
@@
 # Normalize data to z-score
-kegg_pca_mat <- apply(kegg_pca_mat, 1,
-                      norm_pca <- function(x){return (x - mean(x)) / sd(x)}) %>% t()
+kegg_pca_dat <- apply(kegg_pca_dat, 1, norm_pca <- function(x){return (x - mean(x)) / sd(x)}) %>% t()
 
 # PCA based on SVD
-set.seed(12345)
-kegg_pca <- prcomp(kegg_pca_mat, center = FALSE, scale. = FALSE)
+set.seed(122)
+kegg_pca <- prcomp(kegg_pca_dat, center = FALSE, scale. = FALSE)
 
@@
 kegg_pca_scree_df <- data.frame(PC = 1:ncol(kegg_pca$rotation), Var = kegg_pca_var_exp)
-
 kegg_pca_screeplot <- ggplot(kegg_pca_scree_df, aes(x = PC, y = Var)) +
   geom_bar(stat = "identity", fill = "#9cb8ca") +
+  labs(x = "PC", y = "Variance Explained") +
   xlim(0, 50) + ylim(0, 0.06) +
-  labs(title = "KEGG KOs (KOfamScan) Scree Plot", x = "PC", y = "Variance Explained") +
   theme_minimal()
-ggsave("plot_annot_pca_kofam_screeplot.pdf", kegg_pca_screeplot, height = 4, width = 6)
+ggsave("plot_annot_pca_kofam_screeplot.pdf", kegg_pca_screeplot, height = 4, width = 5)
 
 # Biplot
-kegg_pca_colors <- c("Copper" = "#CA6627", "Iron" = "#666666",
-                     "Sulfur" = "#DDAD3B", "Other" = "#e6f0f6")
-
+#kegg_pca_colors <- c("Copper" = "#CA6627", "Iron" = "#666666",
+#                     "Sulfur" = "#DDAD3B", "Other" = "#e6f0f6")
+kegg_pca_colors <- c("Copper" = "#CA6627", "Iron" = "#666666", "Sulfur" = "#DDAD3B")
 kegg_pca_biplot <- ggbiplot(kegg_pca, choices = c(1, 2), groups = kegg_pca_kos$Pathway,
@@
   geom_hline(yintercept = 0, linetype = 3) +
-  ylim(-8, 5) +
+  xlim(-100, 140) +
   #geom_point(aes(colour = kegg_pca_kos$Pathway), size = 1.5) +
@@
 
-# Set colors for ecosystem (standarize**)
-kegg_colors_eco <- setNames(brewer.pal(length(unique(kegg_mags$ecosystem)), "Paired"), unique(kegg_mags$ecosystem))
-kegg_colors_eco <- sample(kegg_colors_eco)
-
-### PCAtools ###
-set.seed(11122)
-kegg_pca_obj <- pca(kegg_pca_mat, center = FALSE, scale = FALSE,
+## PCAtools
+# Try to separate biomining from other ecosystems using selected pathways
+set.seed(221)
+kegg_pca_obj <- pca(kegg_pca_dat, center = FALSE, scale = FALSE,
                     metadata = kegg_mags %>% data.frame(row.names = 1))
@@
                               colLegendTitle = "Ecosystem", colby = "ecosystem", 
-                              colkey = kegg_colors_eco, ellipse = TRUE, ylim = c(-14, 12))
+                              colkey = mags_colors_eco, ellipse = TRUE,
+                              xlim = c(-100, 70), ylim = c(-20, 10))
 ggsave("plot_annot_pcatools_kofam_biplot.pdf", kegg_pca_obj_biplot, height = 9, width = 10)
@@
 kegg_pca_obj_pairs <- pairsplot(kegg_pca_obj, colby = "ecosystem",
-                                colkey = kegg_colors_eco, pointSize = 1,
+                                colkey = mags_colors_eco, pointSize = 1,
                                 components = getComponents(kegg_pca_obj, c(1:4)))
@@
 ```{r annotbub}
-#split KOs per copper, iron and sulfur
-#for each KO, set all ecosystems to definition
-#split columns per mag ecosystems
-#print in row labels KO name + definition
-
-### 1. KOfamScan
+library(agricolae)
+#library(ggpubr)
+
+# Plot design:
+# -split KOs per copper, iron and sulfur;
+# -for each KO, set all ecosystems to definition;
+# -split columns per mag ecosystems;
+# -print in row labels KO name + definition.
 
@@
   as.data.frame() %>%
-  slice(rep(1:n(), each = length(kegg_bubble_eco)))
+  dplyr::slice(rep(1:n(), each = length(kegg_bubble_eco)))
 colnames(kegg_bubble) <- c("KO")
@@
 
-## Function to count matching KOs in all bins for each ecosystem
+# Function to count matching KOs in all bins for each ecosystem
 fx_countKOsbyeco <- function(ko) {
@@
     e <- kegg_mags[kegg_mags$mag == colnames(kegg_filt)[j], ] %>%
-      select(ecosystem) %>%
+      dplyr::select(ecosystem) %>%
       as.character()
@@
 
-# Count KOs
+# Count KOs for selected pathways
+# (time-consuming loop: runtime ~5-6 min for 247*10 rows!)
 kegg_bubble_counts <- tibble(ecosystem = character(), count = numeric())
@@
   kegg_bubble_counts <- rbind(kegg_bubble_counts, counts_df)
-}#runtime ~5 min!
-#clean memory and add column of counts
+}
+# Clean memory and add column of counts
 rm(kos_df, eco_df, counts_df)
@@
 
-# Subset to cycles of interest (3):
-# 1. get definitions and metabolism from kegg_paths
-# 2. merge df with counts per ecosystem
-# 3. remove biomining KOs with low counts (ie. sulfur: <= 5)
-# 4. subset df with kept KOs
-#todo: split definition in two columns by " [EC:"
-kegg_bubble_copper <- kegg_paths %>%
-  filter(Pathway == "Copper") %>%
-  select(KO, Definition, Metabolism)
+# Combine counts and pathways datasets:
+# -get definitions and metabolism from kegg_paths
+# -merge df with counts per ecosystem
+# -remove biomining KOs with low counts (ie. sulfur: <= 5)
+# -subset df with kept KOs
+# *todo: split definition in two columns by " [EC:"
+#copper
+kegg_bubble_copper <- kegg_paths %>% #kegg_paths_sub
+  dplyr::filter(Pathway == "Copper") %>%
+  dplyr::select(KO, Definition, Metabolism)
 kegg_bubble_copper <- left_join(kegg_bubble_copper, kegg_bubble, by = c("KO"))
 kegg_bubble_keep <- kegg_bubble_copper %>%
-  filter(ecosystem == "Biomining" & counts > 1) %>%
-  select(KO)
+  dplyr::filter(ecosystem == "Biomining" & counts > 1) %>%
+  dplyr::select(KO)
 kegg_bubble_copper <- kegg_bubble_copper[kegg_bubble_copper$KO %in% kegg_bubble_keep$KO, ]
 
+#iron
 kegg_bubble_iron <- kegg_paths %>%
-  filter(Pathway == "Iron") %>%
-  select(KO, Definition, Metabolism)
+  dplyr::filter(Pathway == "Iron") %>%
+  dplyr::select(KO, Definition, Metabolism)
 kegg_bubble_iron <- left_join(kegg_bubble_iron, kegg_bubble, by = c("KO"))
 kegg_bubble_keep <- kegg_bubble_iron %>%
-  filter(ecosystem == "Biomining" & counts > 2) %>%
-  select(KO)
+  dplyr::filter(ecosystem == "Biomining" & counts > 1) %>%
+  dplyr::select(KO)
 kegg_bubble_iron <- kegg_bubble_iron[kegg_bubble_iron$KO %in% kegg_bubble_keep$KO, ] %>%
-  filter(!KO == "K02014") #remove outlier with counts > 500
-
+  dplyr::filter(!KO == "K02014") #remove outlier with counts > 500!
+
+#sulfur
 kegg_bubble_sulfur <- kegg_paths %>%
-  filter(Pathway == "Sulfur") %>%
-  select(KO, Definition, Metabolism)
+  dplyr::filter(Pathway == "Sulfur") %>%
+  dplyr::select(KO, Definition, Metabolism)
 kegg_bubble_sulfur <- left_join(kegg_bubble_sulfur, kegg_bubble, by = c("KO"))
 kegg_bubble_keep <- kegg_bubble_sulfur %>%
-  filter(ecosystem == "Biomining" & counts >= 5) %>%
-  select(KO)
+  dplyr::filter(ecosystem == "Biomining" & counts >= 10) %>%
+  dplyr::select(KO)
 kegg_bubble_sulfur <- kegg_bubble_sulfur[kegg_bubble_sulfur$KO %in% kegg_bubble_keep$KO, ]
 
-# Bubble plot
-kegg_bubble_plot <- ggplot(kegg_bubble_iron,
-                           aes(x = ecosystem, y = Definition,
-                               size = counts, color = ecosystem)) +
+# Check sorting. Conclusion: use reorder() instead fct_reorder()
+#kegg_bubble_copper %>% group_by(ecosystem) %>% dplyr::summarise(n = sum(counts))
+
+# ANOVA test:
+# H0 the means are equal for ecosystems (if p < 0.05, we take Ha)
+anova_copper <- aov(counts ~ ecosystem, data = kegg_bubble_copper)
+summary(anova_copper) #p = 0.00244 ** (significance 0.01)
+anova_iron <- aov(counts ~ ecosystem, data = kegg_bubble_iron)
+summary(anova_iron) #p = 0.00912 ** (signif. 0.01)
+anova_sulfur <- aov(counts ~ ecosystem, data = kegg_bubble_sulfur)
+summary(anova_sulfur) #p = 6.3e-10 *** (signif. 0.001)
+
+#for KO definitions (y-axis)
+anova_copper2 <- aov(counts ~ Definition, data = kegg_bubble_copper)
+summary(anova_copper2) #p = 4.26e-13 ***
+anova_iron2 <- aov(counts ~ Definition, data = kegg_bubble_iron)
+summary(anova_iron2) #p = <2e-16 ***
+anova_sulfur2 <- aov(counts ~ Definition, data = kegg_bubble_sulfur)
+summary(anova_sulfur2) #p = <2e-16 ***
+
+# Calculate mean and sd between ecosystems
+anova_copper_stats <- kegg_bubble_copper %>%
+  group_by(ecosystem) %>%
+  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+anova_iron_stats <- kegg_bubble_iron %>%
+  group_by(ecosystem) %>%
+  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+anova_sulfur_stats <- kegg_bubble_sulfur %>%
+  group_by(ecosystem) %>%
+  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+
+#mean and sd between KO definitions
+anova_copper_stats2 <- kegg_bubble_copper %>%
+  group_by(Definition) %>%
+  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+anova_iron_stats2 <- kegg_bubble_iron %>%
+  group_by(Definition) %>%
+  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+anova_sulfur_stats2 <- kegg_bubble_sulfur %>%
+  group_by(Definition) %>%
+  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+
+# FISHER LSD test: inspect which group means are different (ie. tukey)
+# H0: ecosystems are independent
+anova_copper_test <- LSD.test(anova_copper, "ecosystem")
+#print(anova_copper_test)
+anova_copper_test <- anova_copper_test$groups %>%
+  as.data.frame() %>%
+  rownames_to_column("ecosystem")
+anova_iron_test <- LSD.test(anova_iron, "ecosystem")
+anova_iron_test <- anova_iron_test$groups %>%
+  as.data.frame() %>%
+  rownames_to_column("ecosystem")
+anova_sulfur_test <- LSD.test(anova_sulfur, "ecosystem")
+anova_sulfur_test <- anova_sulfur_test$groups %>%
+  as.data.frame() %>%
+  rownames_to_column("ecosystem")
+
+#fisher for KO definitions
+anova_copper_test2 <- LSD.test(anova_copper2, "Definition")
+anova_copper_test2 <- anova_copper_test2$groups %>%
+  as.data.frame() %>%
+  rownames_to_column("Definition") #keep 5
+anova_iron_test2 <- LSD.test(anova_iron2, "Definition")
+anova_iron_test2 <- anova_iron_test2$groups %>%
+  as.data.frame() %>%
+  rownames_to_column("Definition") #keep 5
+anova_sulfur_test2 <- LSD.test(anova_sulfur2, "Definition")
+anova_sulfur_test2 <- anova_sulfur_test2$groups %>%
+  as.data.frame() %>%
+  rownames_to_column("Definition") #keep 7
+
+# Box plot: Cu, Fe, S
+anova_copper_plot <- ggplot(kegg_bubble_copper, aes(x = reorder(ecosystem, -counts),
+                                                    y = counts, color = ecosystem)) +
+  geom_boxplot() +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  geom_text(data = anova_copper_test, aes(label = groups),
+            y = -1.5, vjust = -0.5, size = 3, color = "black") +
+  labs(x = "Ecosystem", y = "KO counts for copper") +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+ggsave("plot_annot_bubble_anova_copper.pdf", anova_copper_plot, height = 7, width = 7)
+
+anova_iron_plot <- ggplot(kegg_bubble_iron, aes(x = reorder(ecosystem, -counts),
+                                                y = counts, color = ecosystem)) +
+  geom_boxplot() +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  geom_text(data = anova_iron_test, aes(label = groups),
+            y = -5, vjust = -0.5, size = 3, color = "black") +
+  labs(x = "Ecosystem", y = "KO counts for iron") +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+ggsave("plot_annot_bubble_anova_iron.pdf", anova_iron_plot, height = 7, width = 7)
+
+anova_sulfur_plot <- ggplot(kegg_bubble_sulfur, aes(x = reorder(ecosystem, -counts),
+                                                    y = counts, color = ecosystem)) +
+  geom_boxplot() +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  geom_text(data = anova_sulfur_test, aes(label = groups),
+            y = -4, vjust = -0.5, size = 3, color = "black") +
+  labs(x = "Ecosystem", y = "KO counts for sulfur") +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+ggsave("plot_annot_bubble_anova_sulfur.pdf", anova_sulfur_plot, height = 7, width = 7)
+
+# Bubble plot integrating stats: Cu, Fe, S
+kegg_bubble_plot_cu <- ggplot(kegg_bubble_copper,
+                              aes(x = reorder(ecosystem, -counts),
+                                  y = Definition, size = counts, color = ecosystem)) +
   geom_point(alpha = 0.7) +
-  scale_color_manual(values = kegg_colors_eco, guide = "none") +
-  labs(title = "", x = "Ecosystem", y = "KEGG KO Definitions") +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  #geom_text(aes(label = counts), vjust = -0.5, size = 2.5) +
+  scale_y_discrete(expand = c(0.06, 0)) + #expand y-axis 6% in both sides
+  geom_text(data = anova_copper_test, aes(x = ecosystem, y = 15.4, label = groups),
+            vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
+  geom_hline(yintercept = 15.4, linetype = 1, linewidth = 0.1) +
+  scale_x_discrete(expand = expansion(add = 0.85)) + #expand x-axis in units at both sides
+  geom_text(data = head(anova_copper_test2, 5),
+            aes(x = 10.45, y = Definition, label = groups),
+            vjust = -0.2, hjust = 0.25, size = 1.6, inherit.aes = FALSE) +
+  labs(title = "", subtitle = "p = 0.002 **",
+       x = "Ecosystem", y = "KO definitions for copper") +
   theme_classic() +
-  theme(legend.title = element_text(size = 10),
-        legend.position = "right",
+  theme(legend.title = element_text(size = 10), legend.position = "right",
+        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_kofam_copper.pdf", kegg_bubble_plot, height = 7, width = 9)
-ggsave("plot_annot_bubble_kofam_iron.pdf", kegg_bubble_plot, height = 7, width = 7)
-ggsave("plot_annot_bubble_kofam_sulfur.pdf", kegg_bubble_plot, height = 9, width = 9)
+ggsave("plot_annot_bubble_kofam_copper.pdf", kegg_bubble_plot_cu, height = 7, width = 9)
+#ggsave("plot_annot_bubble_kofam_copper.png", kegg_bubble_plot_cu, dpi = 300, dev = "png", height = 7, width = 9.5, units = "in")
+
+kegg_bubble_plot_fe <- ggplot(kegg_bubble_iron,
+                              aes(x = reorder(ecosystem, -counts),
+                                  y = Definition, size = counts, color = ecosystem)) +
+  geom_point(alpha = 0.7) +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  scale_y_discrete(expand = c(0.045, 0)) + #expand y-axis 4.5% in both sides
+  geom_text(data = anova_iron_test,
+            aes(x = ecosystem, y = 20.4, label = groups),
+            vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
+  geom_hline(yintercept = 20.4, linetype = 1, linewidth = 0.1) +
+  scale_x_discrete(expand = expansion(add = 0.85)) + #expand x-axis in units at both sides
+  geom_text(data = head(anova_iron_test2, 5),
+            aes(x = 10.45, y = Definition, label = groups),
+            vjust = -0.2, hjust = 0.25, size = 1.6, inherit.aes = FALSE) +
+  labs(title = "", subtitle = "p = 0.009 **",
+       x = "Ecosystem", y = "KO definitions for iron") +
+  theme_classic() +
+  theme(legend.title = element_text(size = 10), legend.position = "right",
+        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
+        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+ggsave("plot_annot_bubble_kofam_iron.pdf", kegg_bubble_plot_fe, height = 7, width = 7)
+
+kegg_bubble_plot_s <- ggplot(kegg_bubble_sulfur,
+                             aes(x = reorder(ecosystem, -counts),
+                                 y = Definition, size = counts, color = ecosystem)) +
+  geom_point(alpha = 0.7) +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  scale_y_discrete(expand = c(0.03, 0)) + #expand y-axis in both sides 3%
+  geom_text(data = anova_sulfur_test,
+            aes(x = ecosystem, y = 41.6, label = groups),
+            vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
+  geom_hline(yintercept = 41.6, linetype = 1, linewidth = 0.1) +
+  scale_x_discrete(expand = expansion(add = 0.8)) + #expand x-axis in units at both sides
+  geom_text(data = head(anova_sulfur_test2, 7),
+            aes(x = 10.45, y = Definition, label = groups),
+            vjust = -0.2, hjust = 0.25, size = 1.6, inherit.aes = FALSE) +
+  labs(title = "", subtitle = "p = 6.3e-10 ***",
+       x = "Ecosystem", y = "KO definitions for sulfur") +
+  theme_classic() +
+  theme(legend.title = element_text(size = 10), legend.position = "right",
+        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
+        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+ggsave("plot_annot_bubble_kofam_sulfur.pdf", kegg_bubble_plot_s, height = 8, width = 9)
+```
+
+## Figure 2
+
+```{r fig2}
+# Merge plots using patchwork
+fig_2 <- { { kegg_bubble_plot_cu / kegg_bubble_plot_fe +
+    plot_layout(ncol = 1, widths = c(0.75)) } | kegg_bubble_plot_s } +
+  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("figures/fig_2.pdf", fig_2, height = 14, width = 18)
 ```
@@
 ```
+
+## Figure 3
+
+```{r fig3}
+# Merge plots using patchwork
+
+```

### 2024-04-29 R/paper.Rmd
updated README (section 3), figures and Rdata

@@
 ---
-title: "Systemix Profiling Article: Combined MAG-SMAG Analysis"
+title: "Systemix Article: Combined MAG-SMAG Profiling"
 author: "Moises A. Rojas"
-date: "2023-12-01"
+date: "2023-12-01 // mod. Mar 2024"
 output: github_document
@@
 
-```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = '/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/')
+```{r setup}
+#knitr::opts_knit$set(root.dir = '/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/')
+knitr::opts_knit$set(root.dir = '/Users/moises/Downloads/mags-all/')
 
 suppressMessages(library(ape)) #tree basics
-suppressMessages(library(treeio))
-suppressMessages(library(ggtree))
-suppressMessages(library(ggplot2))
+suppressMessages(library(treeio)) #join trees
+suppressMessages(library(ggtree)) #tree plotting
+suppressMessages(library(ggplot2)) #plots
 suppressMessages(library(ggtreeExtra)) #geom_fruit
@@
 suppressMessages(library(ggstar)) #geom_star for ggplot2
-suppressMessages(library(phytools)) #MRCA
 suppressMessages(library(tidyverse)) #strings, dataframes processing
@@
 suppressMessages(library(tidytree)) #trees as dataframes
-suppressMessages(library(RColorBrewer)) #colors palette
-suppressWarnings(library(colorspace))
+suppressMessages(library(RColorBrewer)) #color palettes
+suppressWarnings(library(colorspace)) #colors manipulation
 suppressWarnings(library(patchwork)) #side-by-side plotting
+suppressWarnings(library(openxlsx)) #read excel
+suppressWarnings(library(ComplexHeatmap)) #heatmap
+suppressWarnings(library(PCAtools)) #pca
+suppressWarnings(library(ggbiplot)) #pca plots
+suppressWarnings(library(agricolae)) #fisher's test
+suppressWarnings(library(ggpubr)) #stats & plots
+suppressWarnings(library(dndscv)) #dN/dS of snps
+suppressWarnings(library(clusterProfiler)) #gene enrichment
 ```
@@
 
-```{r magsdata}
-library(openxlsx)
-
+```{r magdata}
 # Read Systemix MAGs taxonomy and QC files
@@
 
-# Parse and reorder columns
-mag_samples <- dplyr::select(mag_samples, -c("taxonomy", "gc", "contigs"))
+# Parse columns
+mag_samples <- dplyr::select(mag_samples, -c("taxonomy"))
 mag_samples <- mag_samples %>%
@@
 mag_samples <- mag_samples %>% separate(bin, "\\.", into = c("mag_name", "mag_id"))
-mag_samples <- mag_samples[, c(1, 2, 10, 11, 12, 13, 3, 4, 5, 6, 7, 8, 9)]
 
 # Define final attributes
+mag_samples['mag_num'] <- nrow(mag_samples)
 mag_samples['ecosystem'] <- "Biomining"
@@
 
-# Load SMAG metadata (tables 1 and 2)
+# Load SMAG catalog (metadata from tables 1-2)
 smag_md1 <- read.xlsx("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/SMAG/Extended_Data_Table_1.xlsx")
@@
 # Save attributes
-smag_md1_sel <- smag_md1 %>% dplyr::select("mag_name", "Ecosystem", "Continent")
+smag_md1_sel <- smag_md1 %>%
+  dplyr::select("mag_name", "mag_num", "Ecosystem", "Continent")
 smag_md2_sel <- smag_md2 %>%
   dplyr::select("user_genome", "completeness", "contamination",
-                "strain_heterogeneity", "N50", "classification",
+                "strain_heterogeneity", "length", "N50", "classification",
                 "phylum", "class", "order", "family", "genus", "species")
@@
 # Rename columns
+colnames(smag_data)[6] <- "genome_size"
+colnames(smag_data)[8] <- "kingdom"
+colnames(smag_data)[16] <- "ecosystem"
+colnames(smag_data)[17] <- "continent"
+
 colnames(smag_data)[7] <- "kingdom"
@@
 
-# Clean memory
-rm(smag_md1, smag_md2, smag_md1_sel, smag_md2_sel, mags_qc)
-
 # Merge metadata and save to file
@@
 if (file.exists("4-mags-smag/mags_metadata.tsv") == FALSE) {
-  write.table(mags_metadata, "4-mags-smag/mags_metadata.tsv",
-              sep = "\t", quote = FALSE, row.names = FALSE)
+  write.table(mags_metadata, "4-mags-smag/mags_metadata.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
 }
@@
 if (file.exists("4-mags-smag/smag_data.tsv") == FALSE) {
-  write.table(smag_data, "4-mags-smag/smag_data.tsv",
-              sep = "\t", quote = FALSE, row.names = FALSE)
-}
+  write.table(smag_data, "4-mags-smag/smag_data.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
+}
+
+# Save biomining metadata (S15)
+if (file.exists("4-mags-smag/mags_biomining.tsv") == FALSE) {
+  write.table(mag_samples, "4-mags-smag/mags_biomining.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
+}
+
+# Clean memory
+rm(smag_md1, smag_md2, smag_md1_sel, smag_md2_sel)
+#rm(list = ls(pattern = "^kegg_hm_cols_"))
 ```
 
-## In-depth taxonomic analysis of MAG-SMAG
-
-```{r magstaxa}
+## Section 1: In-depth taxonomy analysis
+
+### Biomining MAGs vs SMAG taxa comparison
+
+```{r magcombtaxa}
 # Biomining provides a new phylum -> Nitrospirota_A
@@
 mags_metadata %>%
-  filter(is.na(ecosystem)) %>%
-  group_by(mag_name) %>%
+  dplyr::filter(is.na(ecosystem)) %>%
+  dplyr::group_by(mag_name) %>%
   dplyr::summarise() #n=401
@@
 mags_metadata_bac <- mags_metadata %>%
-  filter(kingdom == "Bacteria" & !is.na(ecosystem)) #n=32975
-
-# Save to file
+  dplyr::filter(kingdom == "Bacteria" & !is.na(ecosystem)) #n=32975
+
+# Save bacterial metadata to file
 if (file.exists("4-mags-smag/mags_metadata_bac.tsv") == FALSE) {
-  write.table(mags_metadata_bac, "4-mags-smag/mags_metadata_bac.tsv",
-              sep = "\t", quote = FALSE, row.names = FALSE)
+  write.table(mags_metadata_bac, "4-mags-smag/mags_metadata_bac.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
 }
@@
   sort(TRUE) %>%
-  head(n = 30) %>%
+  head(n = 40) %>%
   as.data.frame() %>%
@@
 
-plot_depth_top_phyla <- ggplot(mags_md_top_phy, aes(x = reorder(phy_top, freq), y = freq)) +
+plot_depth_top_phyla <- ggplot(mags_md_top_phy, aes(x = freq, y = reorder(phy_top, freq))) +
   geom_bar(stat = "identity", fill = "#459682") +
-  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.5, size = 3) +
-  labs(x = "Phylum", y = "Number") +
-  scale_y_continuous(limits = c(0, 10000)) +
-  theme_classic() +
-  coord_flip()
+  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.5, size = 2.5) +
+  labs(x = "Number", y = "Phylum") +
+  scale_x_continuous(limits = c(0, 10000)) +
+  theme_classic()
 ggsave("plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)
 
-## Filtering metadata
-
+# Filter metadata..
 # Set colors for ecosystem
-#mags_colors_eco <- vector()
-#for (i in 1:length(kegg_colors_eco)) {
-#  mags_colors_eco <- append(mags_colors_eco, kegg_colors_eco[[i]])
-#}
 #mags_colors_eco <- colorRampPalette(brewer.pal(12, "Paired"))(length(unique(mags_metadata_bac$ecosystem)))
@@
   dplyr::select(mag_name, family, genus, species, ecosystem) %>%
-  filter(species != "" & ecosystem == "Biomining")
+  dplyr::filter(species != "" & ecosystem == "Biomining")
 
@@
   dplyr::select(mag_name, mag_id, phylum, family, genus, species, ecosystem, continent) %>%
-  group_by(ecosystem)
+  dplyr::group_by(ecosystem)
 
@@
   geom_bar(aes(x = species, fill = phylum)) +
-  labs(x = "", y = "Number of known species") +
-  scale_fill_manual(values = "#DD7E8B") +
+  labs(x = "", y = "Number of kSGBs from biomining") +
+  scale_fill_manual(values = "#78Acc2") +
   theme_classic() +
   theme(legend.title = element_blank(),
-        legend.position = c(0.85, 0.05),
-        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
-  facet_wrap(.~ecosystem)
-ggsave("plot_md_biom_species.pdf", plot_depth_spec, height = 6, width = 8)
+        legend.position = c(0.88, 0.12),
+        strip.background = element_blank()) +#axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  facet_grid(.~ecosystem) +
+  coord_flip()
+ggsave("plot_md_biom_species.pdf", plot_depth_spec, height = 2.5, width = 8)
 
@@
   dplyr::filter(species != "") %>%
-  group_by(species, ecosystem) %>%
+  dplyr::group_by(species, ecosystem) %>%
   dplyr::summarise(count_sp = table(species)) %>% #.groups = "drop"
   ungroup() %>%
-  group_by(ecosystem) %>%
+  dplyr::group_by(ecosystem) %>%
   dplyr::summarise(count_eco = sum(count_sp)) %>%
@@
 mags_depth_sgbs <- mags_metadata_bac %>%
-  group_by(phylum, species) %>%
+  dplyr::group_by(phylum, species) %>%
   dplyr::summarise(n_spe = n()) %>%
   ungroup() %>%
-  group_by(phylum) %>%
+  dplyr::group_by(phylum) %>%
   dplyr::mutate(t_phy = sum(n_spe)) %>% ###
@@
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"),
-                    values = c("#DD7E8B", "#459682")) +
+  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
   scale_y_continuous(limits = c(0, 8000)) +
@@
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"),
-                    values = c("#DD7E8B", "#459682")) +
+  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
   scale_y_continuous(limits = c(0, 100)) +
@@
         axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.line.y = element_blank()) +
+        axis.line.y = element_blank(),
+        axis.ticks.y = element_blank()) +
   coord_flip()
@@
   dplyr::filter(ecosystem == "Biomining") %>%
-  group_by(phylum, species) %>%
+  dplyr::group_by(phylum, species) %>%
   dplyr::summarise(n_spe = n()) %>%
   ungroup() %>%
-  group_by(phylum) %>%
+  dplyr::group_by(phylum) %>%
   dplyr::mutate(t_phy = sum(n_spe)) %>%
@@
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"),
-                    values = c("#DD7E8B", "#459682")) +
+  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
   scale_y_continuous(limits = c(0, 25)) +
@@
   dplyr::filter(species == "") %>%
-  group_by(phylum, ecosystem) %>%
+  dplyr::group_by(phylum, ecosystem) %>%
   dplyr::summarise(num = table(ecosystem)) %>%
-  dplyr::mutate(phylum = replace(phylum, num < 5, "Other")) #group lows as "other"
+  dplyr::mutate(phylum = replace(phylum, num < 5, "Other")) #group lows as "other" -> summarise?
 
 plot_depth_usgbs <- ggplot(mags_depth_usgbs_eco,
-                           aes(x = reorder(phylum, num), y = num, fill = ecosystem)) +
+                           aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
   geom_col() +
-  labs(x = "Phylum", y = "Number of uSGBs") +
-  scale_y_continuous(n.breaks = 3) +
+  #geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
+  labs(x = "Number of uSGBs", y = "Phylum") +
+  scale_x_continuous(n.breaks = 3) +
   scale_fill_manual(values = mags_colors_eco) +
@@
         strip.placement = "outside",
+        strip.background = element_blank(),
         panel.border = element_blank()) +
-  facet_grid(.~ecosystem, scales = "free") +
-  coord_flip()
+  facet_grid(.~ecosystem, scales = "free")
 ggsave("plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 8, width = 12)
@@
   dplyr::select(mag_name, mag_id, phylum, family, genus, ecosystem) %>%
-  group_by(genus) %>%
+  dplyr::group_by(genus) %>%
   dplyr::mutate(tot_gen = n()) %>%
@@
 mags_depth_kggbs_biom <- mags_depth_kggbs_biom %>%
-  group_by(genus, ecosystem) %>%
+  dplyr::group_by(genus, ecosystem) %>%
   dplyr::mutate(sum = n()) %>% #sum genus per ecosystem
@@
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_x_continuous(limits = c(0, 100), n.breaks = 5) +
+  scale_x_continuous(limits = c(0, 101), n.breaks = 5) + #check!
   scale_fill_manual(values = mags_colors_eco) +
@@
 
-### --OTHER/PENDING TAXA-- ###
-# Inspect unannotated family-level bins (uFGBs)? #1166
-#mags_depth_ufgbs
-
-# What about novel bacteria at genus level?
+### --EXPLORE OTHER TAXA-- ###
+#inspect unannotated family-level bins (uFGBs)? #1166
+#what about novel bacteria at genus level?
 mags_depth_novel_gen <- mags_metadata_bac %>%
   dplyr::filter(genus == "") %>%
-  group_by(phylum, genus, ecosystem) %>%
+  dplyr::group_by(phylum, genus, ecosystem) %>%
   dplyr::summarise(gen_unk = n()) %>%
-  #arrange(desc(gen)) %>%
+  #dplyr::arrange(desc(gen)) %>%
   ungroup() %>%
-  group_by(genus) %>%
+  dplyr::group_by(genus) %>%
   dplyr::mutate(tot_gen = sum(gen_unk)) %>%
@@
 
-# Get unique classes of unknown genus by ecosystem
-#Leptospirillia
-#Sulfobacillia
+#get unique classes of unknown genus by ecosystem
+#Leptospirillia, Sulfobacillia
 mags_md_unk_gen2 <- mags_metadata_bac %>%
   dplyr::filter(genus == "") %>%
-  group_by(class, ecosystem) %>%
+  dplyr::group_by(class, ecosystem) %>%
   tally() %>%
@@
 
-# Unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
+#unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
 unique(mag_samples[mag_samples[, "family"] != "", "genus"])
@@
 
-##plot
 plot_md_unk_cla <- ggplot(mags_md_unk_gen2[mags_md_unk_gen2$n > 10, ],
@@
   facet_wrap(~ecosystem)
-##
 ggsave("plot_md_unk_class_eco.pdf", plot_md_unk_cla, height = 7, width = 10)
 
-## Class of interest of acidic soils:
-# Alphaproteobacteria, Gammaproteobacteria
-# Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
+#class of interest for acidic soils:
+#Alphaproteobacteria, Gammaproteobacteria
+#Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
 table(mags_metadata_bac[mags_metadata_bac[, "phylum"] == "Proteobacteria", "class"])
@@
 
-## Family of interest reported for Systemix [ref 2022]:
-# Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
+#family of interest reported for Systemix [ref 2022]:
+#Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
 mags_md_fams_acid <- mags_metadata_bac %>%
-  filter(family == "Pseudomonadaceae") %>%
-  group_by(ecosystem) %>%
+  dplyr::filter(family == "Pseudomonadaceae") %>%
+  dplyr::group_by(ecosystem) %>%
   dplyr::summarise(count = n())
@@
 
-### QC (biomining)
-
-```{r magsqc}
+### QC
+
+```{r magbiomqc}
 # Read QC file from CheckM
@@
 # Plot
-plot_mags_qc <- ggplot(mags_qcs,
-                       aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
+plot_mags_qc <- ggplot(mags_qcs, aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
   geom_col(position = "dodge2") +
@@
 
-### Phylogenetic tree (biomining)
-
-```{r magstree}
+### Phylogenetic tree
+
+```{r magbiomtree}
 # Load GTDB taxonomy
@@
 rm(fig_1a, fig_1b)
-#ggsave("fig_1.png", fig_1, dpi = 300, dev = "png", height = 14, width = 14, units = "cm")
-
-##fig sup 1
-fig_s1 <- { plot_depth_top_phyla | plot_depth_spec } +
-  plot_layout(ncol = 2, widths = c(0.5, 1)) &
-  theme(axis.title.x = element_text(margin = margin(t = -75, unit = "pt")))
-fig_s1 <- { fig_s1 /
-    { (plot_spacer() + plot_depth_usgbs + plot_spacer()) +
-        plot_layout(ncol = 3, widths = c(0.5, 12, 2)) } #centered
-  } + plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+
+#fig sup 1
+fig_s1 <- { plot_depth_top_phyla |
+    (plot_depth_usgbs / plot_depth_spec + plot_layout(heights = c(0.85, 0.15))) } +
+  plot_layout(ncol = 2, widths = c(0.25, 1)) +
+  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
   theme(plot.tag =  element_text(face = "bold"))
-ggsave("fig_s1.pdf", fig_s1, height = 14, width = 14)
-#ggsave("fig_s1.png", fig_s1, dpi = 300, dev = "png", height = 14, width = 14, units = "cm")
+ggsave("fig_s1.pdf", fig_s1, height = 8, width = 16)
+#ggsave("fig_s1.png", fig_s1, height = 8, width = 16)
 ```
 
-## Functional Annotation: KEGG KOs from KOfamScan
-
-### Heatmap
-
-```{r annothmap}
-library(ComplexHeatmap)
-
-# Read output file from KOfamScan
-kegg_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/merge_hits/kofamscan_hits.tsv", sep = "\t")
+## Section 2: Functional Annotation
+
+### Pre-process matrix of KEGG KOs from KOfamScan
+
+```{r processkos}
+# Read output matrix from KOfamScan
+kegg_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/merge_hits/kofamscan_hits.tsv", sep = "\t", check.names = FALSE)
+colnames(kegg_annot)[2] <- "KO.definition"
 
@@
   dplyr::select(-c("KO.definition")) %>%
-  data.frame(row.names = 1)
+  data.frame(row.names = 1, check.names = FALSE)
 
@@
 kegg_paths <- kegg_kos %>%
-  #"sulf": to group sulfur, sulfate, sulfane, sulfide, sulfo_
-  #"soxr": SoxR 2Fe-2S cluster
-  dplyr::filter(grepl("iron|sulf|soxr|copper", Definition, ignore.case = TRUE) == TRUE) %>%
+  #sulf: to keep sulfur, sulfate, sulfane, sulfide, sulfo_
+  #soxr: SoxR 2Fe-2S cluster
+  #nitrogen: a negative marker (ubiquous; ex: MerR family)
+  dplyr::filter(grepl("iron|sulf|soxr|copper|nitrogen",
+                      Definition, ignore.case = TRUE) == TRUE) %>%
   dplyr::select(KO, Definition)
@@
     return("Copper")
+  if (grepl("iron", def, ignore.case = TRUE) == TRUE)
+    return("Iron")
   if (grepl("sulf|sox", def, ignore.case = TRUE) == TRUE)
     return("Sulfur")
-  if (grepl("iron", def, ignore.case = TRUE) == TRUE)
-    return("Iron")
+  if (grepl("nitrogen", def, ignore.case = TRUE) == TRUE)
+      return("Nitrogen")
   else
@@
 # Function to define types of metabolism
-fx_setKOmetabolism <- function(def) {
+fx_setKOmetabtypes <- function(def) {
   #oxidation, reduction
   #fixation, hydrolysis, production, utilization, transport
-  #sulfane is an oxidative process, as well as dehydrogenases
+  #sulfane involved in oxidative processes, as well as dehydrogenases
   if (grepl("oxida|oxidi|hydrog|hydrox|sulfane", def, ignore.case = TRUE) == TRUE)
@@
     return("Carrier")
-  if (grepl("transcriptional|regulator", def, ignore.case = TRUE) == TRUE)
+  if (grepl("regulat", def, ignore.case = TRUE) == TRUE)
     return("Regulator")
@@
 }
+# Function that returns the genome size given a MAG
+fx_getMAGlength <- function(mag) {
+  mag_size <- mags_metadata_bac %>%
+    dplyr::mutate(bin = paste0(mag_name, ".", mag_id)) %>%
+    dplyr::filter(bin == mag) %>%
+    dplyr::select(genome_size) %>%
+    as.numeric()
+  return(mag_size)
+}
 ## END FUNCTIONS ##
@@
 	kegg_ko_paths <- append(kegg_ko_paths, fx_setKOpathways(kegg_paths$Definition[i]))
-	kegg_ko_metab <- append(kegg_ko_metab, fx_setKOmetabolism(kegg_paths$Definition[i]))
+	kegg_ko_metab <- append(kegg_ko_metab, fx_setKOmetabtypes(kegg_paths$Definition[i]))
 }
@@
 kegg_paths['Metabolism'] <- kegg_ko_metab
+rm(kegg_ko_paths, kegg_ko_metab) #clean memory
 
 # Keep only selected pathways from the annotation matrix
-# (this matrix will be used later for bubble plots)
+# (NOTE: this matrix will be used later for bubble plots)
 kegg_filt <- kegg_annot[row.names(kegg_annot) %in% kegg_paths$KO, ]
 
+# Annotate ecosystem and genome size for MAGs
+kegg_mags <- colnames(kegg_filt) %>% as.data.frame()
+colnames(kegg_mags) <- c("mag")
+
+kegg_mags_eco <- vector()
+kegg_mags_length <- vector()
+for (i in 1:nrow(kegg_mags)) { #this loop takes a few secs to run!
+	kegg_mags_eco <- append(kegg_mags_eco, fx_getMAGecosystem(kegg_mags$mag[i]))
+	kegg_mags_length <- append(kegg_mags_length, fx_getMAGlength(kegg_mags$mag[i]))
+}
+kegg_mags['ecosystem'] <- kegg_mags_eco
+kegg_mags['length'] <- kegg_mags_length
+rm(kegg_mags_eco, kegg_mags_length) #clean memory
+
+# Explore ecosystems..
+table(kegg_mags$ecosystem)
+kegg_mags %>% dplyr::filter(ecosystem == "Biomining") %>% dplyr::summarise(total = sum(length))
+kegg_mags %>% dplyr::group_by(ecosystem) %>% dplyr::summarise(total = sum(length))
+```
+
+### Heatmap
+
+```{r heatmap}
 # Row-wise variance, then filter both matrix and pathways
-kegg_filt_row_var <- apply(kegg_filt, 1, var)
-#save in another variable!**
-kegg_filt <- kegg_filt[kegg_filt_row_var >= summary(kegg_filt_row_var)[3], ] #median
-kegg_paths_sub <- kegg_paths[kegg_paths$KO %in% row.names(kegg_filt), ]
+kegg_hm_filt_var <- apply(kegg_filt, 1, var)
+kegg_hm_data <- kegg_filt[kegg_hm_filt_var >= summary(kegg_hm_filt_var)[3], ] #median
+kegg_hm_paths <- kegg_paths[kegg_paths$KO %in% row.names(kegg_hm_data), ]
 
 # Normalize rows by z-score
-kegg_filt_norm <- apply(kegg_filt, 1, norm <- function(x){return (x - mean(x)) / sd(x)}) %>% t()
-
-#plot test
-#heatmap(as.matrix(kegg_filt))
-#heatmap(kegg_filt_norm)
+kegg_hm_data <- apply(kegg_hm_data, 1, norm <- function(x){return (x - mean(x)) / sd(x)}) %>% t()
 
@@
 # Annotate rows: KO labels
-#with high number of hits: row.names(kegg_filt[rowSums(kegg_filt) > 100, ])
+#with high number of hits: row.names(kegg_hm_data[rowSums(kegg_hm_data) > 100, ])
 # KOs with low variance (1/2 data %>% sort())
-#kegg_hm_rows_lab <- row.names(kegg_filt[apply(kegg_filt_norm, 1, var) < 0.00015, ])
-#kegg_hm_rows_at <- which(row.names(kegg_filt) %in% kegg_hm_rows_lab)
+#kegg_hm_rows_lab <- row.names(kegg_hm_data[apply(kegg_filt_norm, 1, var) < 0.00015, ])
+#kegg_hm_rows_at <- which(row.names(kegg_hm_data) %in% kegg_hm_rows_lab)
 #kegg_hm_rows_kos <- HeatmapAnnotation(KOs = anno_mark(at = kegg_hm_rows_at,
@@
 #                                      which = "row")
-# Row names in the heatmap are not sorted!
-#row.names(kegg_filt)[row_order(kegg_heatmap)]
 
@@
 kegg_hm_rows_path <- c("Iron" = "#666666", "Sulfur" = "#DDAD3B", "Copper" = "#CA6627")
-kegg_hm_rows_path <- list(Def = kegg_paths_sub$Pathway, Pathways = kegg_hm_rows_path)
-
-kegg_hm_rows_ann_path <- HeatmapAnnotation(Pathways = kegg_paths_sub$Pathway,
+kegg_hm_rows_path <- list(Def = kegg_hm_paths$Pathway, Pathways = kegg_hm_rows_path)
+
+kegg_hm_rows_ann_path <- HeatmapAnnotation(Pathways = kegg_hm_paths$Pathway,
                                            col = kegg_hm_rows_path, border = FALSE,
@@
 # Annotate rows: KO metabolism
-kegg_hm_rows_metab <- setNames(brewer.pal(length(unique(kegg_paths_sub$Metabolism)), "Paired"), unique(kegg_paths_sub$Metabolism)) #Set3
+kegg_hm_rows_metab <- setNames(brewer.pal(length(unique(kegg_hm_paths$Metabolism)), "Paired"), unique(kegg_hm_paths$Metabolism)) #Set3
 kegg_hm_rows_metab <- sample(kegg_hm_rows_metab)
-kegg_hm_rows_metab <- list(Met = kegg_paths_sub$Metabolism, Metabolism = kegg_hm_rows_metab)
-
-kegg_hm_rows_ann_metab <- HeatmapAnnotation(Metabolism = kegg_paths_sub$Metabolism,
+kegg_hm_rows_metab <- list(Met = kegg_hm_paths$Metabolism, Metabolism = kegg_hm_rows_metab)
+
+kegg_hm_rows_ann_metab <- HeatmapAnnotation(Metabolism = kegg_hm_paths$Metabolism,
                                             col = kegg_hm_rows_metab, border = FALSE,
@@
 # Annotate cols: ecosystems
-kegg_mags <- colnames(kegg_filt) %>% as.data.frame()
-colnames(kegg_mags) <- c("mag")
-
-kegg_mags_eco <- vector()
-for (i in 1:nrow(kegg_mags)) {
-	kegg_mags_eco <- append(kegg_mags_eco, fx_getMAGecosystem(kegg_mags$mag[i]))
-}
-kegg_mags['ecosystem'] <- kegg_mags_eco
-rm(kegg_mags_eco)
-#table(kegg_mags$ecosystem)
-
-#match with mags_colors_eco**
+#match colors with mags_colors_eco*
 kegg_hm_cols_mags <- setNames(brewer.pal(length(unique(kegg_mags$ecosystem)), "Paired"), unique(kegg_mags$ecosystem))
@@
 
-# Heatmap with annotations
-pdf("plot_annot_heatmap_kofam.pdf", height = 9, width = 9)
-kegg_heatmap <- Heatmap(kegg_filt_norm, name = "Abundance",
+# Plot heatmap with annotations
+pdf("plot_annot_heatmap.pdf", height = 9, width = 9)
+kegg_hm_plot <- Heatmap(kegg_filt_norm, name = "Abundance",
                         row_title = "KEGG KOs", column_title = "Ecosystem MAGs",
@@
                         col = kegg_hm_colors, border = FALSE,
-                        row_split = kegg_paths_sub$Pathway,
+                        row_split = kegg_hm_paths$Pathway,
                         column_split = kegg_mags$ecosystem,
@@
                         top_annotation = kegg_hm_cols_ann_mags)
-draw(kegg_heatmap, merge_legend = TRUE)
+draw(kegg_hm_plot, merge_legend = TRUE)
 while (!is.null(dev.list())) dev.off()
+# Row names in the heatmap are not sorted!
+#row.names(kegg_hm_data)[row_order(kegg_hm_plot)]
 ```
@@
 
-```{r annotpca}
-library(ggbiplot)
-library(PCAtools)
-
+```{r pca}
 # NOTE: selected cycles (Fe, Cu, S) have low representation in row sums,
@@
 #kegg_pca_dat <- kegg_annot[kegg_pca_annot_row_var > summary(kegg_pca_annot_row_var)[3], ]
-kegg_pca_dat <- kegg_filt
+kegg_pca_dat <- kegg_hm_data
 
@@
   theme_minimal()
-ggsave("plot_annot_pca_kofam_screeplot.pdf", kegg_pca_screeplot, height = 4, width = 5)
+ggsave("plot_annot_pca_screeplot.pdf", kegg_pca_screeplot, height = 4, width = 5)
 
@@
   theme_bw()
-ggsave("plot_annot_pca_kofam_biplot.pdf", kegg_pca_biplot, height = 9, width = 9)
+ggsave("plot_annot_pca_biplot.pdf", kegg_pca_biplot, height = 9, width = 9)
 
@@
                               xlim = c(-100, 70), ylim = c(-20, 10))
-ggsave("plot_annot_pcatools_kofam_biplot.pdf", kegg_pca_obj_biplot, height = 9, width = 10)
+ggsave("plot_annot_pcatools_biplot.pdf", kegg_pca_obj_biplot, height = 9, width = 10)
 
@@
                                 components = getComponents(kegg_pca_obj, c(1:4)))
-ggsave("plot_annot_pcatools_kofam_pairsplot.pdf", kegg_pca_obj_pairs, height = 7, width = 10)
+ggsave("plot_annot_pcatools_pairsplot.pdf", kegg_pca_obj_pairs, height = 7, width = 10)
 ```
 
-### Bubble plot
-
-```{r annotbub}
-library(agricolae)
-#library(ggpubr)
-
-# Plot design:
-# -split KOs per copper, iron and sulfur;
-# -for each KO, set all ecosystems to definition;
-# -split columns per mag ecosystems;
-# -print in row labels KO name + definition.
-
+### Bubble and box plots
+
+```{r bubble}
 # Get unique ecosystems
@@
 
-# Prepare dataframe
+# Take selected KOs and prepare dataset
 kegg_bubble <- row.names(kegg_filt) %>%
   as.data.frame() %>%
-  dplyr::slice(rep(1:n(), each = length(kegg_bubble_eco)))
+  slice(rep(1:n(), each = length(kegg_bubble_eco)))
 colnames(kegg_bubble) <- c("KO")
@@
 kegg_bubble$ecosystem <- rep(kegg_bubble_eco, nrow(kegg_filt))
-#optimize with pivot_longer()?
-#kegg_bubble['count'] <- sample(100, size = nrow(kegg_bubble), replace = TRUE) #plot test
-
-# Function to count matching KOs in all bins for each ecosystem
-fx_countKOsbyeco <- function(ko) {
-  #given a KO, retrieve all bins in the filtered matrix
+
+# Plot design/overview:
+# -split KOs per Cu, Fe and S (N is control)
+# -split x-axis per ecosystems, KO definitions in y-axis
+# -sum KO counts across ecosystems
+# -normalize them taking the genome size
+# -use statistical analysis (anova, groups)
+
+# Function that counts the matching KOs in each ecosystem
+# Also, retrieves the bins size in order to normalize counts later
+fx_countKOsByEco <- function(ko) {
+  #retrieve all bins in the filtered matrix for the given KO
+  #this row contains the counts across all ecosystems
   row <- kegg_filt[row.names(kegg_filt) == ko, ]
@@
   eco <- vector()
-  sum <- vector()
+  len <- vector()
+  cou <- vector()
+  #get the ecosystem, length and counts for each bin
   for (j in 1:ncol(row)) {
-    #get the ecosystem for each bin (column) and sum uniques
     e <- kegg_mags[kegg_mags$mag == colnames(kegg_filt)[j], ] %>%
-      dplyr::select(ecosystem) %>%
-      as.character()
+      dplyr::select(ecosystem) %>% as.character()
     eco <- append(eco, e)
-    sum <- append(sum, row[1, j])
+    l <- kegg_mags[kegg_mags$mag == colnames(kegg_filt)[j], ] %>%
+      dplyr::select(length) %>% as.numeric()
+    len <- append(len, l) #length
+    cou <- append(cou, row[1, j]) #counts
   }
   #vectors to dataframe
-  df <- data.frame(Ecosystem = eco, Counts = sum)
-  #merge ecosystems for counting and return
-  df <- aggregate(df$Counts, list(df$Ecosystem), FUN = sum)
-  colnames(df)[1] <- "ecosystem"
-  colnames(df)[2] <- "count"
+  df <- data.frame(ecosystem = eco, count = cou, length = len)
+  #sum counts and length by ecosystem
+  df <- df %>% dplyr::group_by(ecosystem) %>%
+    dplyr::summarise(across(everything(), sum))
   return(df)
@@
 
-# Count KOs for selected pathways
-# (time-consuming loop: runtime ~5-6 min for 247*10 rows!)
-kegg_bubble_counts <- tibble(ecosystem = character(), count = numeric())
+# Count total KOs and length in all across ecosystems for selected pathways
+# (time-consuming loop: runtime ~12 mins for rows in kegg_paths*)
+kegg_bubble_counts <- tibble(ecosystem = character(), count = numeric(), length = numeric())
+kos_df <- vector()
+eco_df <- vector()
+cou_df <- vector()
 for (i in seq(1, nrow(kegg_bubble), length(kegg_bubble_eco))) {
-  #loop step of 10 KOs
+  #current KO -> loop step of 10 KOs
   #print(sprintf("%i -> KO: %s", i, kegg_bubble[i, 1]))
-  #current KO
-  kos_df <- fx_countKOsbyeco(kegg_bubble[i, 1])
-  #merge ecosystem counts sorted from our stored list
+  kos_df <- fx_countKOsByEco(kegg_bubble[i, 1])
+  #merge ecosystem counts sorted from our list
   eco_df <- kegg_bubble_eco %>% as.data.frame()
   colnames(eco_df) <- c("ecosystem")
-  counts_df <- left_join(eco_df, kos_df, by = c("ecosystem"))
-  kegg_bubble_counts <- rbind(kegg_bubble_counts, counts_df)
-}
-# Clean memory and add column of counts
-rm(kos_df, eco_df, counts_df)
+  #save counts
+  cou_df <- left_join(eco_df, kos_df, by = c("ecosystem"))
+  kegg_bubble_counts <- rbind(kegg_bubble_counts, cou_df)
+}
+# Add column for counts
 kegg_bubble <- cbind(kegg_bubble, counts = kegg_bubble_counts[, 2])
+kegg_bubble <- cbind(kegg_bubble, length = kegg_bubble_counts[, 3])
+rm(kos_df, eco_df, cou_df) #clean memory
+
+# Normalize counts: length to mb and get counts per mb
+kegg_bubble <- kegg_bubble %>%
+  dplyr::mutate(length_mb = round(length / 1000000, 3)) %>%
+  dplyr::mutate(count_norm = round(counts / length_mb, 3))
+#explore
+kegg_bubble %>% dplyr::group_by(ecosystem) %>% dplyr::summarise(n = sum(count_norm))
 
@@
 # -merge df with counts per ecosystem
-# -remove biomining KOs with low counts (ie. sulfur: <= 5)
-# -subset df with kept KOs
-# *todo: split definition in two columns by " [EC:"
-#copper
-kegg_bubble_copper <- kegg_paths %>% #kegg_paths_sub
+# -remove KOs from biomining with low counts
+
+# Copper pathway
+kegg_bubble_cu <- kegg_paths %>%
   dplyr::filter(Pathway == "Copper") %>%
   dplyr::select(KO, Definition, Metabolism)
-kegg_bubble_copper <- left_join(kegg_bubble_copper, kegg_bubble, by = c("KO"))
-kegg_bubble_keep <- kegg_bubble_copper %>%
-  dplyr::filter(ecosystem == "Biomining" & counts > 1) %>%
+kegg_bubble_cu <- left_join(kegg_bubble_cu, kegg_bubble, by = c("KO"))
+#filter
+kegg_bubble_keep <- kegg_bubble_cu %>%
+  dplyr::filter(ecosystem == "Biomining" & count_norm > 0) %>%
   dplyr::select(KO)
-kegg_bubble_copper <- kegg_bubble_copper[kegg_bubble_copper$KO %in% kegg_bubble_keep$KO, ]
-
-#iron
-kegg_bubble_iron <- kegg_paths %>%
+kegg_bubble_cu <- kegg_bubble_cu[kegg_bubble_cu$KO %in% kegg_bubble_keep$KO, ]
+
+# Iron pathway
+kegg_bubble_fe <- kegg_paths %>%
   dplyr::filter(Pathway == "Iron") %>%
   dplyr::select(KO, Definition, Metabolism)
-kegg_bubble_iron <- left_join(kegg_bubble_iron, kegg_bubble, by = c("KO"))
-kegg_bubble_keep <- kegg_bubble_iron %>%
-  dplyr::filter(ecosystem == "Biomining" & counts > 1) %>%
+kegg_bubble_fe <- left_join(kegg_bubble_fe, kegg_bubble, by = c("KO"))
+#filter
+kegg_bubble_keep <- kegg_bubble_fe %>%
+  dplyr::filter(ecosystem == "Biomining" & count_norm > 0.01) %>%
   dplyr::select(KO)
-kegg_bubble_iron <- kegg_bubble_iron[kegg_bubble_iron$KO %in% kegg_bubble_keep$KO, ] %>%
-  dplyr::filter(!KO == "K02014") #remove outlier with counts > 500!
-
-#sulfur
-kegg_bubble_sulfur <- kegg_paths %>%
+kegg_bubble_fe <- kegg_bubble_fe[kegg_bubble_fe$KO %in% kegg_bubble_keep$KO, ] %>%
+  dplyr::filter(!KO == "K02013" & !KO == "K02014") #outliers
+
+# Sulfur pathway
+kegg_bubble_s <- kegg_paths %>%
   dplyr::filter(Pathway == "Sulfur") %>%
   dplyr::select(KO, Definition, Metabolism)
-kegg_bubble_sulfur <- left_join(kegg_bubble_sulfur, kegg_bubble, by = c("KO"))
-kegg_bubble_keep <- kegg_bubble_sulfur %>%
-  dplyr::filter(ecosystem == "Biomining" & counts >= 10) %>%
+kegg_bubble_s <- left_join(kegg_bubble_s, kegg_bubble, by = c("KO"))
+#filter
+kegg_bubble_keep <- kegg_bubble_s %>%
+  dplyr::filter(ecosystem == "Biomining" & count_norm > 0.05) %>%
   dplyr::select(KO)
-kegg_bubble_sulfur <- kegg_bubble_sulfur[kegg_bubble_sulfur$KO %in% kegg_bubble_keep$KO, ]
-
-# Check sorting. Conclusion: use reorder() instead fct_reorder()
-#kegg_bubble_copper %>% group_by(ecosystem) %>% dplyr::summarise(n = sum(counts))
+kegg_bubble_s <- kegg_bubble_s[kegg_bubble_s$KO %in% kegg_bubble_keep$KO, ]
+
+# Sort ecosystems (by median)
+anova_sort_cu <- kegg_bubble_cu %>%
+  dplyr::group_by(ecosystem) %>%
+  dplyr::summarise(n = median(count_norm)) %>%
+  dplyr::arrange(desc(n)) %>%
+  as.data.frame()
+anova_sort_fe <- kegg_bubble_fe %>%
+  dplyr::group_by(ecosystem) %>%
+  dplyr::summarise(n = median(count_norm)) %>%
+  dplyr::arrange(desc(n)) %>%
+  as.data.frame()
+anova_sort_s <- kegg_bubble_s %>%
+  dplyr::group_by(ecosystem) %>%
+  dplyr::summarise(n = median(count_norm)) %>%
+  dplyr::arrange(desc(n)) %>%
+  as.data.frame()
+
+# We need to re-sort the default colors..
+mags_ecosystem <- kegg_bubble_eco %>% sort()
+mags_colors <- tibble(ecosystem = mags_ecosystem, colors = mags_colors_eco)
+anova_sort_cu <- left_join(anova_sort_cu, mags_colors, by = c("ecosystem"))
+anova_sort_fe <- left_join(anova_sort_fe, mags_colors, by = c("ecosystem"))
+anova_sort_s <- left_join(anova_sort_s, mags_colors, by = c("ecosystem"))
+
+# Boxplots
+anova_boxplot_cu <- ggboxplot(kegg_bubble_cu, x = "ecosystem", y = "count_norm",
+                           color = "ecosystem", palette = anova_sort_cu$colors,
+                           order = anova_sort_cu$ecosystem, add = "jitter", legend = "none") +
+  geom_hline(yintercept = mean(kegg_bubble_cu$count_norm),
+             linetype = 2, linewidth = 0.2) +
+  labs(x = "Ecosystem", y = "KOs/Mb for copper") +
+  rotate_x_text(angle = 45) + #ylim(c(0, 0.25)) +
+  #stat_summary(fun = mean, geom = "point", shape = 18, size = 2.5, color = "#FC4E07") +
+  stat_compare_means(method = "anova", label.y = 0.25) + #add global annova p-value
+  stat_compare_means(label = "p.signif", method = "t.test",
+                     ref.group = ".all.", hide.ns = TRUE)
+ggsave("plot_annot_boxplot_copper.pdf", anova_boxplot_cu, height = 7, width = 7)
+
+anova_boxplot_fe <- ggboxplot(kegg_bubble_fe, x = "ecosystem", y = "count_norm",
+                           color = "ecosystem", palette = anova_sort_fe$colors,
+                           order = anova_sort_fe$ecosystem, add = "jitter", legend = "none") +
+  geom_hline(yintercept = mean(kegg_bubble_fe$count_norm),
+             linetype = 2, linewidth = 0.2) +
+  labs(x = "Ecosystem", y = "KOs/Mb for iron") +
+  rotate_x_text(angle = 45) +
+  stat_compare_means(method = "anova", label.y = 0.4) + #add global annova p-value
+  stat_compare_means(label = "p.signif", method = "t.test",
+                     ref.group = ".all.", hide.ns = TRUE, label.y = 0.35)
+ggsave("plot_annot_boxplot_iron.pdf", anova_boxplot_fe, height = 7, width = 7)
+
+anova_boxplot_s <- ggboxplot(kegg_bubble_s, x = "ecosystem", y = "count_norm",
+                             color = "ecosystem", palette = anova_sort_s$colors,
+                             order = anova_sort_s$ecosystem, add = "jitter", legend = "none") +
+  geom_hline(yintercept = mean(kegg_bubble_s$count_norm),
+             linetype = 2, linewidth = 0.2) +
+  labs(x = "Ecosystem", y = "KOs/Mb for sulfur") +
+  rotate_x_text(angle = 45) +
+  stat_compare_means(method = "anova", label.y = 0.7) + #add global annova p-value
+  stat_compare_means(label = "p.signif", method = "t.test",
+                     ref.group = ".all.", hide.ns = TRUE)
+ggsave("plot_annot_boxplot_sulfur.pdf", anova_boxplot_s, height = 7, width = 7)
+
+# Bubble plot with key KOs from biomining
+#multicopper, copper resistance (B,C,D), menaquinol
+#UDP−sulfoquinovose, high−affinity iron transporter, thiosulfate dehydrogenase
+#sulfide:quinone oxidoreductase, sulfate adenylyltransferase, mononucleotide sulfurtransferase, 
+#heterodisulfide C2, sulfolactate sulfo-lyase subunit, CysO sulfur-carrier
+kegg_biom_key <- c("K22552", "K07233", "K07156", "K07245", "K03886",
+                   "K06118", "K07243", "K19713", "K16937",
+                   "K17218", "K00958", "K06864", 
+                   "K03390", "K01023", "K16845", "K16846", "K21140", "K21148")
+
+kegg_bubbles_plot <- kegg_bubble %>%
+  filter(KO %in% kegg_biom_key) %>%
+  left_join(kegg_paths, by = c("KO")) %>% #add definitions, then plot
+  ggplot(aes(x = reorder(ecosystem, -count_norm),
+             y = Definition, size = count_norm, color = ecosystem)) +
+  geom_point(alpha = 0.7) +
+  scale_color_manual(values = mags_colors_eco, guide = "none") +
+  labs(title = "", x = "Ecosystem", y = "Highlighted KO definitions") +
+  guides(size = guide_legend(title = "Norm. counts")) +
+  theme_classic() +
+  theme(legend.title = element_text(size = 9), legend.position = "right",
+        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
+        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+ggsave("plot_annot_bubbles.pdf", kegg_bubbles_plot, height = 7, width = 8.5)
 
@@
 # H0 the means are equal for ecosystems (if p < 0.05, we take Ha)
-anova_copper <- aov(counts ~ ecosystem, data = kegg_bubble_copper)
-summary(anova_copper) #p = 0.00244 ** (significance 0.01)
-anova_iron <- aov(counts ~ ecosystem, data = kegg_bubble_iron)
-summary(anova_iron) #p = 0.00912 ** (signif. 0.01)
-anova_sulfur <- aov(counts ~ ecosystem, data = kegg_bubble_sulfur)
-summary(anova_sulfur) #p = 6.3e-10 *** (signif. 0.001)
-
-#for KO definitions (y-axis)
-anova_copper2 <- aov(counts ~ Definition, data = kegg_bubble_copper)
-summary(anova_copper2) #p = 4.26e-13 ***
-anova_iron2 <- aov(counts ~ Definition, data = kegg_bubble_iron)
-summary(anova_iron2) #p = <2e-16 ***
-anova_sulfur2 <- aov(counts ~ Definition, data = kegg_bubble_sulfur)
-summary(anova_sulfur2) #p = <2e-16 ***
-
-# Calculate mean and sd between ecosystems
-anova_copper_stats <- kegg_bubble_copper %>%
-  group_by(ecosystem) %>%
-  dplyr::summarise(mean = mean(counts), sd = sd(counts))
-anova_iron_stats <- kegg_bubble_iron %>%
-  group_by(ecosystem) %>%
-  dplyr::summarise(mean = mean(counts), sd = sd(counts))
-anova_sulfur_stats <- kegg_bubble_sulfur %>%
-  group_by(ecosystem) %>%
-  dplyr::summarise(mean = mean(counts), sd = sd(counts))
-
-#mean and sd between KO definitions
-anova_copper_stats2 <- kegg_bubble_copper %>%
-  group_by(Definition) %>%
-  dplyr::summarise(mean = mean(counts), sd = sd(counts))
-anova_iron_stats2 <- kegg_bubble_iron %>%
-  group_by(Definition) %>%
-  dplyr::summarise(mean = mean(counts), sd = sd(counts))
-anova_sulfur_stats2 <- kegg_bubble_sulfur %>%
-  group_by(Definition) %>%
-  dplyr::summarise(mean = mean(counts), sd = sd(counts))
-
-# FISHER LSD test: inspect which group means are different (ie. tukey)
-# H0: ecosystems are independent
-anova_copper_test <- LSD.test(anova_copper, "ecosystem")
-#print(anova_copper_test)
-anova_copper_test <- anova_copper_test$groups %>%
-  as.data.frame() %>%
-  rownames_to_column("ecosystem")
-anova_iron_test <- LSD.test(anova_iron, "ecosystem")
-anova_iron_test <- anova_iron_test$groups %>%
-  as.data.frame() %>%
-  rownames_to_column("ecosystem")
-anova_sulfur_test <- LSD.test(anova_sulfur, "ecosystem")
-anova_sulfur_test <- anova_sulfur_test$groups %>%
-  as.data.frame() %>%
-  rownames_to_column("ecosystem")
-
-#fisher for KO definitions
-anova_copper_test2 <- LSD.test(anova_copper2, "Definition")
-anova_copper_test2 <- anova_copper_test2$groups %>%
-  as.data.frame() %>%
-  rownames_to_column("Definition") #keep 5
-anova_iron_test2 <- LSD.test(anova_iron2, "Definition")
-anova_iron_test2 <- anova_iron_test2$groups %>%
-  as.data.frame() %>%
-  rownames_to_column("Definition") #keep 5
-anova_sulfur_test2 <- LSD.test(anova_sulfur2, "Definition")
-anova_sulfur_test2 <- anova_sulfur_test2$groups %>%
-  as.data.frame() %>%
-  rownames_to_column("Definition") #keep 7
-
-# Box plot: Cu, Fe, S
-anova_copper_plot <- ggplot(kegg_bubble_copper, aes(x = reorder(ecosystem, -counts),
-                                                    y = counts, color = ecosystem)) +
-  geom_boxplot() +
-  scale_color_manual(values = mags_colors_eco, guide = "none") +
-  geom_text(data = anova_copper_test, aes(label = groups),
-            y = -1.5, vjust = -0.5, size = 3, color = "black") +
-  labs(x = "Ecosystem", y = "KO counts for copper") +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_anova_copper.pdf", anova_copper_plot, height = 7, width = 7)
-
-anova_iron_plot <- ggplot(kegg_bubble_iron, aes(x = reorder(ecosystem, -counts),
-                                                y = counts, color = ecosystem)) +
-  geom_boxplot() +
-  scale_color_manual(values = mags_colors_eco, guide = "none") +
-  geom_text(data = anova_iron_test, aes(label = groups),
-            y = -5, vjust = -0.5, size = 3, color = "black") +
-  labs(x = "Ecosystem", y = "KO counts for iron") +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_anova_iron.pdf", anova_iron_plot, height = 7, width = 7)
-
-anova_sulfur_plot <- ggplot(kegg_bubble_sulfur, aes(x = reorder(ecosystem, -counts),
-                                                    y = counts, color = ecosystem)) +
-  geom_boxplot() +
-  scale_color_manual(values = mags_colors_eco, guide = "none") +
-  geom_text(data = anova_sulfur_test, aes(label = groups),
-            y = -4, vjust = -0.5, size = 3, color = "black") +
-  labs(x = "Ecosystem", y = "KO counts for sulfur") +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_anova_sulfur.pdf", anova_sulfur_plot, height = 7, width = 7)
-
-# Bubble plot integrating stats: Cu, Fe, S
-kegg_bubble_plot_cu <- ggplot(kegg_bubble_copper,
-                              aes(x = reorder(ecosystem, -counts),
-                                  y = Definition, size = counts, color = ecosystem)) +
+#anova_cu <- aov(count_norm ~ ecosystem, data = kegg_bubble_cu)
+#summary(anova_cu) #p = 0.0866
+#anova_iron <- aov(counts ~ ecosystem, data = kegg_bubble_iron)
+#summary(anova_iron) #p = 0.00912 ** (signif. 0.01)
+#anova_sulfur <- aov(counts ~ ecosystem, data = kegg_bubble_sulfur)
+#summary(anova_sulfur) #p = 6.3e-10 *** (signif. 0.001)
+
+## Calculate mean and sd between ecosystems
+#anova_copper_stats <- kegg_bubble_copper %>%
+#  group_by(ecosystem) %>%
+#  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+#anova_iron_stats <- kegg_bubble_iron %>%
+#  group_by(ecosystem) %>%
+#  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+#anova_sulfur_stats <- kegg_bubble_sulfur %>%
+#  group_by(ecosystem) %>%
+#  dplyr::summarise(mean = mean(counts), sd = sd(counts))
+#
+## FISHER LSD test: inspect which group means are different (ie. tukey)
+## H0: ecosystems are independent
+#anova_copper_test <- LSD.test(anova_copper, "ecosystem")
+##print(anova_copper_test)
+#anova_copper_test <- anova_copper_test$groups %>%
+#  as.data.frame() %>%
+#  rownames_to_column("ecosystem")
+#anova_iron_test <- LSD.test(anova_iron, "ecosystem")
+#anova_iron_test <- anova_iron_test$groups %>%
+#  as.data.frame() %>%
+#  rownames_to_column("ecosystem")
+#anova_sulfur_test <- LSD.test(anova_sulfur, "ecosystem")
+#anova_sulfur_test <- anova_sulfur_test$groups %>%
+#  as.data.frame() %>%
+#  rownames_to_column("ecosystem")
+
+# Box plots -> from checking use reorder()* instead fct_reorder()*!
+#anova_copper_plot <- ggplot(kegg_bubble_copper, aes(x = reorder(ecosystem, -counts),
+#                                                    y = counts, color = ecosystem)) +
+#  geom_boxplot() +
+#  scale_color_manual(values = mags_colors_eco, guide = "none") +
+#  geom_text(data = anova_copper_test, aes(label = groups),
+#            y = -1.5, vjust = -0.5, size = 3, color = "black") +
+#  labs(x = "Ecosystem", y = "KO counts for copper") +
+#  theme_bw() +
+#  theme(panel.grid.major = element_blank(),
+#        panel.grid.minor = element_blank(),
+#        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+#ggsave("plot_annot_bubble_anova_copper.pdf", anova_copper_plot, height = 7, width = 7)
+#
+#anova_iron_plot <- ggplot(kegg_bubble_iron, aes(x = reorder(ecosystem, -counts),
+#                                                y = counts, color = ecosystem)) +
+#  geom_boxplot() +
+#  scale_color_manual(values = mags_colors_eco, guide = "none") +
+#  geom_text(data = anova_iron_test, aes(label = groups),
+#            y = -5, vjust = -0.5, size = 3, color = "black") +
+#  labs(x = "Ecosystem", y = "KO counts for iron") +
+#  theme_bw() +
+#  theme(panel.grid.major = element_blank(),
+#        panel.grid.minor = element_blank(),
+#        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+#ggsave("plot_annot_bubble_anova_iron.pdf", anova_iron_plot, height = 7, width = 7)
+#
+#anova_sulfur_plot <- ggplot(kegg_bubble_sulfur, aes(x = reorder(ecosystem, -counts),
+#                                                    y = counts, color = ecosystem)) +
+#  geom_boxplot() +
+#  scale_color_manual(values = mags_colors_eco, guide = "none") +
+#  geom_text(data = anova_sulfur_test, aes(label = groups),
+#            y = -4, vjust = -0.5, size = 3, color = "black") +
+#  labs(x = "Ecosystem", y = "KO counts for sulfur") +
+#  theme_bw() +
+#  theme(panel.grid.major = element_blank(),
+#        panel.grid.minor = element_blank(),
+#        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
+#ggsave("plot_annot_bubble_anova_sulfur.pdf", anova_sulfur_plot, height = 7, width = 7)
+
+# Bubble plots for each pathway
+kegg_bubble_plot_cu <- ggplot(kegg_bubble_cu,
+                              aes(x = reorder(ecosystem, -count_norm),
+                                  y = Definition, size = count_norm, color = ecosystem)) +
   geom_point(alpha = 0.7) +
   scale_color_manual(values = mags_colors_eco, guide = "none") +
-  #geom_text(aes(label = counts), vjust = -0.5, size = 2.5) +
-  scale_y_discrete(expand = c(0.06, 0)) + #expand y-axis 6% in both sides
-  geom_text(data = anova_copper_test, aes(x = ecosystem, y = 15.4, label = groups),
-            vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
-  geom_hline(yintercept = 15.4, linetype = 1, linewidth = 0.1) +
-  scale_x_discrete(expand = expansion(add = 0.85)) + #expand x-axis in units at both sides
-  geom_text(data = head(anova_copper_test2, 5),
-            aes(x = 10.45, y = Definition, label = groups),
-            vjust = -0.2, hjust = 0.25, size = 1.6, inherit.aes = FALSE) +
-  labs(title = "", subtitle = "p = 0.002 **",
+  #scale_y_discrete(expand = c(0.06, 0)) + #expand y-axis 6% in both sides
+  #geom_text(data = anova_copper_test, aes(x = ecosystem, y = 15.4, label = groups),
+  #          vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
+  #geom_hline(yintercept = 15.4, linetype = 1, linewidth = 0.1) +
+  #scale_x_discrete(expand = expansion(add = 0.85)) + #expand x-axis in units at both sides
+  labs(title = "", subtitle = "p = 0.087",
        x = "Ecosystem", y = "KO definitions for copper") +
+  guides(size = guide_legend(title = "Norm. counts")) +
   theme_classic() +
-  theme(legend.title = element_text(size = 10), legend.position = "right",
+  theme(legend.title = element_text(size = 9), legend.position = "right",
         plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_kofam_copper.pdf", kegg_bubble_plot_cu, height = 7, width = 9)
-#ggsave("plot_annot_bubble_kofam_copper.png", kegg_bubble_plot_cu, dpi = 300, dev = "png", height = 7, width = 9.5, units = "in")
-
-kegg_bubble_plot_fe <- ggplot(kegg_bubble_iron,
-                              aes(x = reorder(ecosystem, -counts),
-                                  y = Definition, size = counts, color = ecosystem)) +
+ggsave("plot_annot_bubble_copper.pdf", kegg_bubble_plot_cu, height = 7, width = 9)
+
+kegg_bubble_plot_fe <- ggplot(kegg_bubble_fe,
+                              aes(x = reorder(ecosystem, -count_norm),
+                                  y = Definition, size = count_norm, color = ecosystem)) +
   geom_point(alpha = 0.7) +
   scale_color_manual(values = mags_colors_eco, guide = "none") +
-  scale_y_discrete(expand = c(0.045, 0)) + #expand y-axis 4.5% in both sides
-  geom_text(data = anova_iron_test,
-            aes(x = ecosystem, y = 20.4, label = groups),
-            vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
-  geom_hline(yintercept = 20.4, linetype = 1, linewidth = 0.1) +
-  scale_x_discrete(expand = expansion(add = 0.85)) + #expand x-axis in units at both sides
-  geom_text(data = head(anova_iron_test2, 5),
-            aes(x = 10.45, y = Definition, label = groups),
-            vjust = -0.2, hjust = 0.25, size = 1.6, inherit.aes = FALSE) +
-  labs(title = "", subtitle = "p = 0.009 **",
+  #scale_y_discrete(expand = c(0.045, 0)) + #expand y-axis 4.5% in both sides
+  #geom_text(data = anova_iron_test,
+  #          aes(x = ecosystem, y = 20.4, label = groups),
+  #          vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
+  #geom_hline(yintercept = 20.4, linetype = 1, linewidth = 0.1) +
+  labs(title = "", subtitle = "p = 0.99",
        x = "Ecosystem", y = "KO definitions for iron") +
+  guides(size = guide_legend(title = "Norm. counts")) +
   theme_classic() +
-  theme(legend.title = element_text(size = 10), legend.position = "right",
+  theme(legend.title = element_text(size = 9), legend.position = "right",
         plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_kofam_iron.pdf", kegg_bubble_plot_fe, height = 7, width = 7)
-
-kegg_bubble_plot_s <- ggplot(kegg_bubble_sulfur,
-                             aes(x = reorder(ecosystem, -counts),
-                                 y = Definition, size = counts, color = ecosystem)) +
+ggsave("plot_annot_bubble_iron.pdf", kegg_bubble_plot_fe, height = 7, width = 8)
+
+kegg_bubble_plot_s <- ggplot(kegg_bubble_s,
+                             aes(x = reorder(ecosystem, -count_norm),
+                                 y = Definition, size = count_norm, color = ecosystem)) +
   geom_point(alpha = 0.7) +
   scale_color_manual(values = mags_colors_eco, guide = "none") +
-  scale_y_discrete(expand = c(0.03, 0)) + #expand y-axis in both sides 3%
-  geom_text(data = anova_sulfur_test,
-            aes(x = ecosystem, y = 41.6, label = groups),
-            vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
-  geom_hline(yintercept = 41.6, linetype = 1, linewidth = 0.1) +
-  scale_x_discrete(expand = expansion(add = 0.8)) + #expand x-axis in units at both sides
-  geom_text(data = head(anova_sulfur_test2, 7),
-            aes(x = 10.45, y = Definition, label = groups),
-            vjust = -0.2, hjust = 0.25, size = 1.6, inherit.aes = FALSE) +
-  labs(title = "", subtitle = "p = 6.3e-10 ***",
+  #scale_y_discrete(expand = c(0.03, 0)) + #expand y-axis in both sides 3%
+  #geom_text(data = anova_sulfur_test,
+  #          aes(x = ecosystem, y = 41.6, label = groups),
+  #          vjust = -0.5, size = 2.5, inherit.aes = FALSE) +
+  #geom_hline(yintercept = 41.6, linetype = 1, linewidth = 0.1) +
+  labs(title = "", subtitle = "p = 0.062",
        x = "Ecosystem", y = "KO definitions for sulfur") +
+  guides(size = guide_legend(title = "Norm. counts")) +
   theme_classic() +
-  theme(legend.title = element_text(size = 10), legend.position = "right",
+  theme(legend.title = element_text(size = 9), legend.position = "right",
         plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
         axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubble_kofam_sulfur.pdf", kegg_bubble_plot_s, height = 8, width = 9)
+ggsave("plot_annot_bubble_sulfur.pdf", kegg_bubble_plot_s, height = 8, width = 9)
 ```
@@
 # Merge plots using patchwork
-fig_2 <- { { kegg_bubble_plot_cu / kegg_bubble_plot_fe +
-    plot_layout(ncol = 1, widths = c(0.75)) } | kegg_bubble_plot_s } +
-  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+fig_2 <- { { anova_boxplot_cu / anova_boxplot_fe +
+    plot_layout(ncol = 1, widths = c(0.75)) } | anova_boxplot_s / kegg_bubbles_plot } +
+  plot_annotation(tag_levels = list(c("a", "b", "c", "d"))) &
   theme(plot.tag =  element_text(face = "bold"))
-ggsave("figures/fig_2.pdf", fig_2, height = 14, width = 18)
+ggsave("fig_2.pdf", fig_2, height = 14, width = 18)
+#ggsave("fig_2.png", fig_2, height = 14, width = 18)
 ```
 
-## SNPs
+## Section 3: SNP Calling
+
+### dN/dS analysis
 
 ```{r snps}
-library(VariantAnnotation)
-library(GenomicRanges)
-library(vcfR)
-
-# Load VCF resulting from BCFTools
-vcf_file <- readVcf("../../snps/S15.vcf")
-
-# Header information
-header(vcf_file)
-samples(header(vcf_file))
-geno(header(vcf_file))
-info(header(vcf_file))
-
-# Genomic positions (CHROM, POS, ID)
-seqnames(vcf_file)
-head(rowRanges(vcf_file), 3)
-ref(vcf_file)[1:5]
-qual(vcf_file)[1:5]
-alt(vcf_file)[1:5]
-
-# Genotype data (FORMAT)
-geno(vcf_file)
-dim(geno(vcf_file)$GT)
-
-# Info data
-info(vcf_file)[1:4, 1:5]
-
-# Plot
-#vcf_plot <- ggplot(data=d.fig5, aes(x=Output, y=bcftools, fill=Output)) +
-#  geom_bar(aes(alpha=0.9), stat="identity", color="black", position=position_dodge()) +
-#  theme_bw() + ylab("Similarity (%)") + xlab("") + expand_limits(y=c(0, 105)) + 
-#  facet_grid(rows = vars(Label_bcftools), cols = vars(Assembler)) +
-#  geom_text(aes(label=mean_bcftools), vjust=-0.3, color="black", size=2.5)
-#ggsave("plot_annot_bubble_kofam_copper.pdf", kegg_bubble_plot, height = 7, width = 9)
-
-# Read VCF file with vcfR package
-vcf <- read.vcfR("../../snps/S15.vcf")
-
-# Parse DP from the gt region
-#https://onlinelibrary.wiley.com/doi/10.1111/1755-0998.12549
-vcf_dp <- extract.gt(vcf, element = "DP", return.alleles = FALSE)
-vcf_dpf <- melt(vcf_dp, varnames = c("Index", "Sample"), value.name = "Depth", na.rm = TRUE)
-
-# Render a violin plot
-
-# Create a chrom plot
-vcf_chrom <- create.chromR(vcf, name = "S15 NODE_1") #subsetted to the first chromosome
-vcf_chrom <- proc.chromR(vcf_chrom)
-chromoqc(vcf_chrom)
+# Explore CDS annotations (from Bakta)
+snp_refcds <- read.csv("5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+# Buildref testing for bacteria.. (1 bin)
+#https://github.com/im3sanger/dndscv/issues/62
+#buildref(cdsfile = "5-SNPs/dndscv/buildref/cds_node_25854.txt", genomefile = "5-SNPs/dndscv/buildref/node_25854.fa", outfile = "5-SNPs/dndscv/buildref/testref.rda")
+#load("5-SNPs/dndscv/buildref/testref.rda")
+
+# Buildref for all bins (very slow: ~2 hrs in building object)
+#buildref(cdsfile = "5-SNPs/dndscv/buildref/cds_table.txt", genomefile = "5-SNPs/dndscv/buildref/S15_bins.fa", outfile = "5-SNPs/dndscv/buildref/S15_refcds.rda")
+
+# Inspect refdb
+load("5-SNPs/dndscv/buildref/S15_refcds.rda")
+print(RefCDS[[1]])
+rm(gr_genes, RefCDS)
+
+# Get annotations from the VCF generated with SnpEff
+#sampleID chr pos ref mut
+snp_mutations <- read.csv("5-SNPs/dndscv/run/S15_mutations.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+rm(snp_mutations)
+
+# Run driver discovery or genes under positive selection (PS) (runtime ~5 min)
+#snp_dndsout <- dndscv(snp_mutations, refdb = "5-SNPs/dndscv/buildref/S15_refcds.rda", cv = NULL, max_muts_per_gene_per_sample = Inf, max_coding_muts_per_sample = Inf)
+#save(snp_dndsout, file = "5-SNPs/dndscv/run/S15_dndsout.rda")
+#load("5-SNPs/dndscv/run/S15_dndsout.rda")
+
+# Explore output object..
+#https://github.com/im3sanger/dndscv/issues/6
+#https://github.com/im3sanger/dndscv/issues/64
+head(snp_dndsout$sel_cv) #dN/dS ratios at gene-level
+head(snp_dndsout$annotmuts)
+print(snp_dndsout$globaldnds) #driver globals
+mean(snp_selcv$wmis_cv) #mis avg.
+
+# Save dN/dS per-gene matrix
+snp_selcv <- snp_dndsout$sel_cv
+rm(snp_dndsout)
+
+# Group global dN/dS ratios:
+#https://github.com/im3sanger/dndscv/issues/63
+#consider significant p-value -> pmis_cv < 0.05?
+#snp_selcv %>% dplyr::filter(pmis_cv < 0.05) %>% nrow() #n=109523
+snp_global_pos <- snp_selcv %>%
+  dplyr::filter(wmis_cv > 1) %>%
+  dplyr::summarise(genes = n())
+snp_global_pos$label = "dN/dS > 1"
+snp_global_neg <- snp_selcv %>%
+  dplyr::filter(wmis_cv < 1 & wmis_cv != 0) %>%
+  dplyr::summarise(genes = n())
+snp_global_neg$label = "dN/dS < 1"
+snp_global_neu <- snp_selcv %>%
+  dplyr::filter(wmis_cv == 0) %>%
+  dplyr::summarise(genes = n())
+snp_global_neu$label = "dN/dS = 0"
+snp_global_dnds <- rbind(snp_global_pos, snp_global_neg, snp_global_neu)
+snp_global_dnds <- snp_global_dnds %>%
+  dplyr::mutate(perc = round(genes * 100 / sum(snp_global_dnds$genes), 1)) %>% #add %
+  dplyr::mutate(perc2 = paste0(genes, " (", perc, "%)")) #label to plot
+rm(snp_global_pos, snp_global_neg, snp_global_neu) #clean memory
+
+# Filter genes under PS:
+#https://github.com/im3sanger/dndscv/issues/56
+#wmis_cv: dN/dS for missense mutations
+#pmis_cv: p-value < 0.05 is expected to occur by chance under neutrality (null hypothesis)
+#p-values for multiple testing (q-values)
+#qallsubs_cv = 144 genes; significant q-values (p-values adjusted for multiple testing)
+snp_psgenes <- snp_selcv %>%
+  dplyr::filter(wmis_cv > 1 & pmis_cv < 0.05) %>%
+  dplyr::mutate(dndsr = n_mis / n_syn)
+#snp_psgenes %>% dplyr::filter(dndsr == Inf) %>% nrow() #n=118
+
+# Load contig to bin matrix
+snp_bins <- read.csv("5-SNPs/dndscv/results/S15_DASTool_contig2bin.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+# Load gene name annotations (from genes.gff file)
+snp_genids <- read.csv("5-SNPs/dndscv/results/gene_id_name.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+## FUNCTIONS ##
+# Function that returns the contig where a gene was found
+fx_dndsGene2Contig <- function(g) {
+  c <- snp_refcds %>%
+    dplyr::filter(gene.name == g) %>%
+    dplyr::select(chr) %>%
+    as.character()
+  return(c)
+}
+# Function that returns the bin given a contig
+fx_dndsContig2Bin <- function(c) {
+  b <- snp_bins %>%
+    dplyr::filter(contig == c) %>%
+    dplyr::select(bin) %>%
+    as.character()
+  return(b)
+}
+# Function that gets gene names
+fx_dndsGeneName <- function(g) {
+  g <- snp_genids %>%
+    dplyr::filter(gene.id == g) %>%
+    dplyr::select(gene.name) %>%
+    as.character()
+  return(g)
+}
+# Function that returns a row with taxa given a bin
+fx_dndsTaxaBin <- function(b) {
+  t <- mag_samples %>%
+    dplyr::mutate(bin = paste0(mag_name, ".", mag_id)) %>%
+    dplyr::filter(bin == b) %>%
+    dplyr::select(phylum, genus, species, completeness, contamination)
+  return(t)
+}
+# Function that process the product tag of NS genes
+fx_dndsNsPathway <- function(p) {
+  if (grepl("copper", p, ignore.case = TRUE) == TRUE)
+    return("Copper")
+  if (grepl("nickel", p, ignore.case = TRUE) == TRUE)
+    return("Nickel")
+  if (grepl("sulfur", p, ignore.case = TRUE) == TRUE)
+    return("Sulfur")
+  if (grepl("mercury", p, ignore.case = TRUE) == TRUE)
+    return("Mercury")
+  if (grepl("nitrogen", p, ignore.case = TRUE) == TRUE)
+    return("Nitrogen")
+}
+# Function that gets dndscv stats for NS genes
+fx_dndsNsStats <- function(g) {
+  s <- snp_selcv %>%
+    dplyr::filter(gene_name == g) %>%
+    dplyr::select(gene_name, wmis_cv, pmis_cv)
+  return(s)
+}
+## END FUNCTIONS ##
+
+# Populate dataset with additional attributes per gene
+snp_gene2contig <- vector()
+snp_contig2bin <- vector()
+snp_genenames <- vector()
+for (i in 1:nrow(snp_psgenes)) {
+	snp_gene2contig <- append(snp_gene2contig, fx_dndsGene2Contig(snp_psgenes$gene_name[i]))
+	snp_contig2bin <- append(snp_contig2bin, fx_dndsContig2Bin(snp_gene2contig[i]))
+	snp_genenames <- append(snp_genenames, fx_dndsGeneName(snp_psgenes$gene_name[i]))
+}
+snp_psgenes['contig'] <- snp_gene2contig
+snp_psgenes['bin'] <- snp_contig2bin
+snp_psgenes['name'] <- snp_genenames
+rm(snp_refcds, snp_bins, snp_genids, snp_gene2contig, snp_contig2bin, snp_genenames) #clean memory
+
+# Select PS and add taxonomic classification
+snp_dnds <- snp_psgenes %>%
+  #dplyr::filter(dndsr > 1) %>%
+  dplyr::group_by(bin) %>%
+  dplyr::summarise(num = n()) %>%
+  dplyr::arrange(desc(num))
+
+snp_taxabins <- tibble(phylum = character(), genus = character(), species = character(),
+                       completeness = numeric(), contamination = numeric())
+for (i in 1:nrow(snp_dnds)) {
+  #print(sprintf("%i -> bin: %s", i, fx_dndsTaxaBin(snp_dnds$bin[i])))
+  snp_taxabins <- rbind(snp_taxabins, fx_dndsTaxaBin(snp_dnds$bin[i]))
+}
+snp_dnds['phylum'] <- snp_taxabins$phylum
+snp_dnds['genus'] <- snp_taxabins$genus
+snp_dnds['species'] <- snp_taxabins$species
+snp_dnds['completeness'] <- snp_taxabins$completeness
+snp_dnds['contamination'] <- snp_taxabins$contamination
+rm(snp_taxabins)
+
+# Set unknown genus/species (uGGBs and uSGBs)
+snp_dnds <- snp_dnds %>%
+  dplyr::mutate(genus = ifelse(genus == "", "uGGB", genus)) %>%
+  dplyr::mutate(species = ifelse(species == "", "uSGB", species))
+
+# Subset to plot mutations -> synonymous/missense (non-syn.)/nonsense
+snp_psgenes_sub <- snp_psgenes %>%
+  dplyr::filter(wmis_cv > 5.5) %>%
+  pivot_longer(cols = c(n_syn, n_mis, n_non), names_to = "mut", values_to = "n_mut")
+
+# Barplot global dN/dS ratios
+snp_plot_dndstot <- ggplot(snp_global_dnds, aes(x = genes, y = reorder(label, genes), fill = label)) +
+  geom_col(position = "dodge2", fill = "#4c6194") +
+  geom_text(aes(label = perc2), vjust = 0.5, hjust = -0.1, size = 2) +
+  #scale_fill_manual(values = c("#c56c65", "#778ab3", "#529548")) +
+  scale_x_continuous(limits = c(0, 125000)) +
+  labs(x = "Number of genes", y = "") +
+  theme_classic() +
+  theme(legend.position = "none")
+ggsave("5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndstot, height = 3, width = 5)
+
+# Plot global driver genes per taxa
+snp_plot_psbins <- ggplot(snp_dnds, aes(x = num, y = reorder(bin, num), fill = bin)) +
+  geom_bar(stat = "identity", fill = "#E31A1C") +
+  geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
+  labs(x = "Number of genes under PS", y = "Bin") +
+  scale_x_continuous(limits = c(0, 120)) +
+  theme_classic() +
+  theme(strip.background = element_blank()) +
+  facet_grid(.~species)
+ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 8.5, width = 9.5)
+
+# Lollipop plot of PS
+#ggdotchart(snp_psgenes_sub, x = "name", y = "wmis_cv", color = "#E31A1C", rotate = TRUE, sorting = "descending", add = "segments", ggtheme = theme_pubr())
+snp_plot_psgenes <- ggplot(snp_psgenes_sub,
+                           aes(x = wmis_cv, y = reorder(name, wmis_cv)), size = wmis_cv) +
+  geom_segment(aes(x = 0, xend = wmis_cv, y = name), color = "gray") +
+  geom_point(alpha = 0.7, color = "#E31A1C") +
+  geom_vline(xintercept = max(snp_psgenes$wmis_cv), linetype = 1, linewidth = 0.1) +
+  labs(x = "PS strength", y = "Gene CDS") +
+  theme_classic()
+ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 9, width = 7)
+
+# Plot mutations/genes per type
+snp_plot_muts <- ggplot(snp_psgenes_sub, aes(x = n_mut, y = reorder(name, wmis_cv), fill = mut)) +
+  geom_col(position = position_stack(reverse = TRUE)) +
+  scale_fill_manual(name = "", labels = c("mis", "non", "syn"),
+                    values = c("#5db4b2", "#b687b6", "#2374a4")) +
+  labs(x = "Number of mutations", y = "") +
+  theme_classic() +
+  theme(legend.position = c(0.85, 0.15),
+        strip.placement = "outside",
+        axis.text.y = element_blank(),
+        axis.line.y = element_blank(),
+        axis.ticks.y = element_blank())
+#ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psmuts.pdf", snp_plot_muts, height = 9, width = 7)
+
+snp_plot_psgenes <- (snp_plot_psgenes | snp_plot_muts) +
+  plot_layout(nrow = 1, ncol = 2, widths = c(0.55, 0.5))
+ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psgenmut.pdf", snp_plot_psgenes, height = 9, width = 9)
+
+# Negative selected (NS) genes with dN/dS < 1 (selected metals)
+snp_nsgenes <- read.csv("5-SNPs/dndscv/results/selected_cds_metals.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+# Parse product to pathway and add additional attributes (wmis_cv, pmis_cv)
+#add dndsr, contig, bin**?
+snp_nspath <- tibble(gene_name = character(), pathway = character())
+snp_nsstat <- tibble(wmis_cv = numeric(), pmis_cv = numeric())
+for (i in 1:nrow(snp_nsgenes)) {
+  #work using dataframes, not lists
+  snp_nspath <- snp_nspath %>%
+    add_row(gene_name = snp_nsgenes$gene_name[i], pathway = fx_dndsNsPathway(snp_nsgenes$product[i]))
+	snp_nsstat <- rbind(snp_nsstat, fx_dndsNsStats(snp_nsgenes$gene_name[i]))
+}
+snp_nsgenes <- left_join(snp_nsgenes, snp_nspath, by = c("gene_name"))
+snp_nsgenes <- left_join(snp_nsgenes, snp_nsstat, by = c("gene_name"))
+rm(snp_nspath, snp_nsstat)
+
+# Remove NA cells and significance: ie. BAOFFNEK_01342 not found
+snp_nsgenes <- snp_nsgenes %>%
+  dplyr::filter(!is.na(pathway) & !(is.na(wmis_cv))) %>%
+  dplyr::filter(wmis_cv != 0 & pmis_cv < 0.05)
+
+# Sort pathways and set colors
+snp_sort_ns <- snp_nsgenes %>%
+  dplyr::group_by(pathway) %>%
+  dplyr::summarise(n = median(wmis_cv)) %>%
+  dplyr::arrange(desc(n)) %>%
+  as.data.frame()
+snp_sort_ns$color <- c("#8c2421", "#4f4fad", "#ca6627", "#6a9b85")
+
+# Violin plot for NS genes with global mutations
+snp_plot_nsmut <- ggviolin(snp_nsgenes, x = "pathway", y = "wmis_cv", color = "pathway",
+                           shape = "pathway", xlab = "Pathway", ylab = "dN/dS",
+                           order = snp_sort_ns$pathway, add = "jitter", legend = "none",
+                           palette = snp_sort_ns$color) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
+  scale_y_continuous(limits = c(-0.2, 1)) +
+  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2)
+ggsave("5-SNPs/dndscv/results/plot/plot_dnds_nsmut.pdf", snp_plot_nsmut, height = 9, width = 7)
+
+# Merge PS and NS gene names
+snp_gmerge <- bind_rows(snp_psgenes, snp_nsgenes)
+snp_gmerge <- snp_gmerge %>%
+  dplyr::mutate(pathway = ifelse(is.na(pathway), "PS", pathway))
+#snp_color_gm <- c("#ddad3b", "#4f4fad", "#6a9b85", "#ca6627", "#8c2421", "black")
+snp_color_gm <- c("#8c2421", "#4f4fad", "#6a9b85", "#ca6627")
+
+# Lollipop plot dN/dS x-axis range 0-2, y-axis PS genes vs selected CDS
+snp_plot_gmerge <- ggdotchart(snp_nsgenes, x = "gene_name", y = "wmis_cv", color = "pathway",
+                              xlab = "Gene CDS", ylab = "dN/dS", y.text.col = FALSE, rotate = TRUE,
+                              palette = snp_color_gm, sorting = "descending", ggtheme = theme_pubr()) +
+  scale_y_continuous(limits = c(0, 0.4)) +
+  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
+  theme(axis.text.y = element_blank(),
+        axis.ticks.y = element_blank())
+ggsave("5-SNPs/dndscv/results/plot/plot_dnds_nspath.pdf", snp_plot_gmerge, height = 9, width = 7)
+
+# Genes for enrichment (we need KEGG KOs..)
+#snp_genenrich <- snp_psgenes$dndsr
+#names(snp_genenrich) <- snp_psgenes$name
+#snp_kegg <- enrichKEGG(snp_genenrich, "hsa", pvalueCutoff = 0.05)
+
 ```
@@
 # Merge plots using patchwork
-
+fig_3 <- { { anova_boxplot_cu / anova_boxplot_fe +
+    plot_layout(ncol = 1, widths = c(0.75)) } | anova_boxplot_s / kegg_bubbles_plot } +
+  plot_annotation(tag_levels = list(c("a", "b", "c", "d"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("fig_3.pdf", fig_3, height = 14, width = 18)
+ggsave("fig_3.png", fig_3, height = 14, width = 18)
 ```

### 2024-05-10 R/paper.Rmd
update to manuscript version 1

@@
 ```{r setup}
-#knitr::opts_knit$set(root.dir = '/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/')
-knitr::opts_knit$set(root.dir = '/Users/moises/Downloads/mags-all/')
+knitr::opts_knit$set(root.dir = "/Users/moises/Downloads/")
 
@@
 ggsave("fig_1.pdf", fig_1, height = 14, width = 14)
+#ggsave("fig_1.png", fig_1, height = 14, width = 14)
 rm(fig_1a, fig_1b)
@@
   theme(plot.tag =  element_text(face = "bold"))
+fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.title.y = element_text(vjust = 0))
+#fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.text.y = element_text(vjust = 0))
 ggsave("fig_s1.pdf", fig_s1, height = 8, width = 16)
@@
   theme(plot.tag =  element_text(face = "bold"))
+#adjust y-label margin for sulfur plot
+fig_2[[2]][[1]] <- fig_2[[2]][[1]] + theme(axis.title.y = element_text(vjust = -127))
 ggsave("fig_2.pdf", fig_2, height = 14, width = 18)
-#ggsave("fig_2.png", fig_2, height = 14, width = 18)
+#ggsave("fig_2.png", fig_2, height = 14, width = 18) vjust=120
 ```
 
-## Section 3: SNP Calling
+## Section 3: SNP Variant Calling
 
@@
 ```{r snps}
-# Explore CDS annotations (from Bakta)
-snp_refcds <- read.csv("5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+# Get CDS annotations (pre-processed from Bakta results)
+snp_refcds <- read.csv("mags-all/5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
@@
 #https://github.com/im3sanger/dndscv/issues/62
-#buildref(cdsfile = "5-SNPs/dndscv/buildref/cds_node_25854.txt", genomefile = "5-SNPs/dndscv/buildref/node_25854.fa", outfile = "5-SNPs/dndscv/buildref/testref.rda")
-#load("5-SNPs/dndscv/buildref/testref.rda")
+#buildref(cdsfile = "mags-all/5-SNPs/dndscv/buildref/cds_node_25854.txt",
+#         genomefile = "mags-all/5-SNPs/dndscv/buildref/node_25854.fa",
+#         outfile = "mags-all/5-SNPs/dndscv/buildref/testref.rda")
+#load("mags-all/5-SNPs/dndscv/buildref/testref.rda")
 
 # Buildref for all bins (very slow: ~2 hrs in building object)
-#buildref(cdsfile = "5-SNPs/dndscv/buildref/cds_table.txt", genomefile = "5-SNPs/dndscv/buildref/S15_bins.fa", outfile = "5-SNPs/dndscv/buildref/S15_refcds.rda")
+#buildref(cdsfile = "mags-all/5-SNPs/dndscv/buildref/cds_table.txt",
+#         genomefile = "mags-all/5-SNPs/dndscv/buildref/S15_bins.fa",
+#         outfile = "mags-all/5-SNPs/dndscv/buildref/S15_refcds.rda")
 
 # Inspect refdb
-load("5-SNPs/dndscv/buildref/S15_refcds.rda")
-print(RefCDS[[1]])
-rm(gr_genes, RefCDS)
-
-# Get annotations from the VCF generated with SnpEff
+#load("mags-all/5-SNPs/dndscv/buildref/S15_refcds.rda")
+#print(RefCDS[[1]])
+#rm(gr_genes, RefCDS)
+
+# Get annotations (parsed from the VCF generated with SnpEff)
 #sampleID chr pos ref mut
-snp_mutations <- read.csv("5-SNPs/dndscv/run/S15_mutations.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_mutations <- read.csv("mags-all/5-SNPs/dndscv/run/S15_mutations.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+# Run dndscv (runtime ~8 min, high memory usage) and save object
+#snp_dndsout <- dndscv(snp_mutations, refdb = "mags-all/5-SNPs/dndscv/buildref/S15_refcds.rda",
+#                      max_muts_per_gene_per_sample = Inf, max_coding_muts_per_sample = Inf,
+#                      cv = NULL, outmats = TRUE)
+#save(snp_dndsout, file = "mags-all/5-SNPs/dndscv/run/S15_dndsout.rda")
+#save(substmodel, file = "mags-all/5-SNPs/dndscv/run/S15_substmodel.rda")
+#load("mags-all/5-SNPs/dndscv/run/S15_dndsout.rda")
+#load("mags-all/5-SNPs/dndscv/run/S15_substmodel.rda")
 rm(snp_mutations)
-
-# Run driver discovery or genes under positive selection (PS) (runtime ~5 min)
-#snp_dndsout <- dndscv(snp_mutations, refdb = "5-SNPs/dndscv/buildref/S15_refcds.rda", cv = NULL, max_muts_per_gene_per_sample = Inf, max_coding_muts_per_sample = Inf)
-#save(snp_dndsout, file = "5-SNPs/dndscv/run/S15_dndsout.rda")
-#load("5-SNPs/dndscv/run/S15_dndsout.rda")
 
@@
 print(snp_dndsout$globaldnds) #driver globals
-mean(snp_selcv$wmis_cv) #mis avg.
 
 # Save dN/dS per-gene matrix
+#driver discovery or genes under positive selection (PS)
 snp_selcv <- snp_dndsout$sel_cv
@@
 #https://github.com/im3sanger/dndscv/issues/63
-#consider significant p-value -> pmis_cv < 0.05?
-#snp_selcv %>% dplyr::filter(pmis_cv < 0.05) %>% nrow() #n=109523
-snp_global_pos <- snp_selcv %>%
+snp_selcv %>% dplyr::filter(pmis_cv < 0.05) %>% nrow() #n=105328
+snp_global_ps <- snp_selcv %>%
   dplyr::filter(wmis_cv > 1) %>%
   dplyr::summarise(genes = n())
-snp_global_pos$label = "dN/dS > 1"
-snp_global_neg <- snp_selcv %>%
+snp_global_ps$label = "dN/dS > 1"
+snp_global_ns <- snp_selcv %>%
   dplyr::filter(wmis_cv < 1 & wmis_cv != 0) %>%
   dplyr::summarise(genes = n())
-snp_global_neg$label = "dN/dS < 1"
-snp_global_neu <- snp_selcv %>%
-  dplyr::filter(wmis_cv == 0) %>%
-  dplyr::summarise(genes = n())
-snp_global_neu$label = "dN/dS = 0"
-snp_global_dnds <- rbind(snp_global_pos, snp_global_neg, snp_global_neu)
+snp_global_ns$label = "dN/dS < 1"
+snp_global_dnds <- rbind(snp_global_ps, snp_global_ns)
 snp_global_dnds <- snp_global_dnds %>%
-  dplyr::mutate(perc = round(genes * 100 / sum(snp_global_dnds$genes), 1)) %>% #add %
+  dplyr::mutate(perc = round(genes * 100 / nrow(snp_selcv), 2)) %>% #add %
   dplyr::mutate(perc2 = paste0(genes, " (", perc, "%)")) #label to plot
-rm(snp_global_pos, snp_global_neg, snp_global_neu) #clean memory
+rm(snp_global_ps, snp_global_ns) #clean memory
 
@@
 #wmis_cv: dN/dS for missense mutations
-#pmis_cv: p-value < 0.05 is expected to occur by chance under neutrality (null hypothesis)
-#p-values for multiple testing (q-values)
-#qallsubs_cv = 144 genes; significant q-values (p-values adjusted for multiple testing)
+#pmis_cv: p-value (signif. < 0.05) is expected to occur by chance under neutrality (null hypothesis)
 snp_psgenes <- snp_selcv %>%
-  dplyr::filter(wmis_cv > 1 & pmis_cv < 0.05) %>%
-  dplyr::mutate(dndsr = n_mis / n_syn)
-#snp_psgenes %>% dplyr::filter(dndsr == Inf) %>% nrow() #n=118
+  dplyr::filter(wmis_cv > 1) %>%
+  dplyr::mutate(dndsr = round((n_mis / n_syn), 1))
 
 # Load contig to bin matrix
-snp_bins <- read.csv("5-SNPs/dndscv/results/S15_DASTool_contig2bin.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
-
-# Load gene name annotations (from genes.gff file)
-snp_genids <- read.csv("5-SNPs/dndscv/results/gene_id_name.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_bins <- read.csv("mags-all/5-SNPs/dndscv/results/S15_DASTool_contig2bin.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+# Load gene names (parsed from genes.gff file)
+snp_genids <- read.csv("mags-all/5-SNPs/dndscv/results/gene_id_name.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
@@
 }
-# Function that gets gene names
+# Function that gets gene names given an id
 fx_dndsGeneName <- function(g) {
@@
 }
-# Function that returns a row with taxa given a bin
+# Function that returns a row with the taxonomic classification of a bin
 fx_dndsTaxaBin <- function(b) {
@@
     dplyr::filter(bin == b) %>%
-    dplyr::select(phylum, genus, species, completeness, contamination)
+    dplyr::select(phylum, genus, species, predicted_genes)
   return(t)
@@
     return("Copper")
+  if (grepl("iron", p, ignore.case = TRUE) == TRUE)
+    return("Iron")
   if (grepl("nickel", p, ignore.case = TRUE) == TRUE)
@@
     return("Sulfur")
-  if (grepl("mercury", p, ignore.case = TRUE) == TRUE)
-    return("Mercury")
   if (grepl("nitrogen", p, ignore.case = TRUE) == TRUE)
@@
 
-# Select PS and add taxonomic classification
+# Group PS genes to add taxa & QC
 snp_dnds <- snp_psgenes %>%
-  #dplyr::filter(dndsr > 1) %>%
   dplyr::group_by(bin) %>%
@@
 snp_dnds['species'] <- snp_taxabins$species
-snp_dnds['completeness'] <- snp_taxabins$completeness
-snp_dnds['contamination'] <- snp_taxabins$contamination
+snp_dnds['predicted_genes'] <- snp_taxabins$predicted_genes
 rm(snp_taxabins)
@@
 
-# Subset to plot mutations -> synonymous/missense (non-syn.)/nonsense
+# Subset PS genes to plot mutations -> synonymous/missense (non-syn.)/nonsense
+#discard HPs (by name)
+#genes with at least 5 SNPs
+#mark genes reaching significance (q < 0.05)
+#mark genes: KJCAEPPB_04437 -> nickel pathway; NEBJFEFK_01910 -> sulfur pathway
 snp_psgenes_sub <- snp_psgenes %>%
-  dplyr::filter(wmis_cv > 5.5) %>%
+  dplyr::filter(nchar(name) < 14) %>%
+  dplyr::mutate(snps = n_syn + n_mis + n_non) %>%
+  dplyr::filter(snps >= 4) %>%
+  dplyr::mutate(signif = ifelse(pmis_cv < 0.05, "*", "")) %>%
+  dplyr::mutate(path = case_when(gene_name == "KJCAEPPB_04437" ~ "Ni",
+                                 gene_name == "NEBJFEFK_01910" ~ "S", TRUE ~ "")) %>%
+  distinct(name, .keep_all = TRUE) %>% #remove repeated names
   pivot_longer(cols = c(n_syn, n_mis, n_non), names_to = "mut", values_to = "n_mut")
 
-# Barplot global dN/dS ratios
-snp_plot_dndstot <- ggplot(snp_global_dnds, aes(x = genes, y = reorder(label, genes), fill = label)) +
+# Save PS gene names and ids to file
+if (file.exists("mags-all/5-SNPs/dndscv/results/ps_genes.txt") == FALSE) {
+  write.table(snp_psgenes_sub %>% dplyr::select(gene_name) %>% distinct(gene_name), "mags-all/5-SNPs/dndscv/results/ps_genes.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+if (file.exists("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt") == FALSE) {
+  write.table(snp_psgenes_sub %>% dplyr::select(name) %>% distinct(name), "mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+
+# Barplot with global dN/dS ratios
+snp_plot_dndsglob <- ggplot(snp_global_dnds, aes(x = genes, y = reorder(label, genes), fill = label)) +
   geom_col(position = "dodge2", fill = "#4c6194") +
   geom_text(aes(label = perc2), vjust = 0.5, hjust = -0.1, size = 2) +
-  #scale_fill_manual(values = c("#c56c65", "#778ab3", "#529548")) +
-  scale_x_continuous(limits = c(0, 125000)) +
+  scale_x_continuous(limits = c(0, 16000)) +
   labs(x = "Number of genes", y = "") +
@@
   theme(legend.position = "none")
-ggsave("5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndstot, height = 3, width = 5)
-
-# Plot global driver genes per taxa
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndsglob, height = 2, width = 5)
+
+# Plot global driver genes by species
 snp_plot_psbins <- ggplot(snp_dnds, aes(x = num, y = reorder(bin, num), fill = bin)) +
@@
   labs(x = "Number of genes under PS", y = "Bin") +
-  scale_x_continuous(limits = c(0, 120)) +
+  scale_x_continuous(limits = c(0, 255)) +
   theme_classic() +
@@
   facet_grid(.~species)
-ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 8.5, width = 9.5)
-
-# Lollipop plot of PS
-#ggdotchart(snp_psgenes_sub, x = "name", y = "wmis_cv", color = "#E31A1C", rotate = TRUE, sorting = "descending", add = "segments", ggtheme = theme_pubr())
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 8.5, width = 9.5)
+
+# Lollipop plot of PS genes
 snp_plot_psgenes <- ggplot(snp_psgenes_sub,
@@
   geom_point(alpha = 0.7, color = "#E31A1C") +
-  geom_vline(xintercept = max(snp_psgenes$wmis_cv), linetype = 1, linewidth = 0.1) +
+  geom_text(aes(label = signif, x = -0.15), size = 3, color = "#3c4d76") +
+  geom_text(aes(label = path, x = -0.15), size = 2, color = "#3c4d76") +
+  geom_vline(xintercept = 1, linetype = 1, linewidth = 0.2) +
+  geom_vline(xintercept = max(snp_psgenes_sub$wmis_cv), linetype = 1, linewidth = 0.2) +
   labs(x = "PS strength", y = "Gene CDS") +
   theme_classic()
-ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 9, width = 7)
-
-# Plot mutations/genes per type
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 11, width = 7)
+
+# Plot PS genes mutations per type
 snp_plot_muts <- ggplot(snp_psgenes_sub, aes(x = n_mut, y = reorder(name, wmis_cv), fill = mut)) +
@@
         axis.ticks.y = element_blank())
-#ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psmuts.pdf", snp_plot_muts, height = 9, width = 7)
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psmuts.pdf", snp_plot_muts, height = 11, width = 7)
 
@@
   plot_layout(nrow = 1, ncol = 2, widths = c(0.55, 0.5))
-ggsave("5-SNPs/dndscv/results/plot/plot_dnds_psgenmut.pdf", snp_plot_psgenes, height = 9, width = 9)
-
-# Negative selected (NS) genes with dN/dS < 1 (selected metals)
-snp_nsgenes <- read.csv("5-SNPs/dndscv/results/selected_cds_metals.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
-
-# Parse product to pathway and add additional attributes (wmis_cv, pmis_cv)
-#add dndsr, contig, bin**?
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psgenmuts.pdf", snp_plot_psgenes, height = 11, width = 11)
+rm(snp_plot_muts)
+
+# Load negative selected (NS) genes
+snp_nsgenes <- read.csv("mags-all/5-SNPs/dndscv/results/selected_cds_metals.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+
+# Infer pathway from product and add dnds attributes
 snp_nspath <- tibble(gene_name = character(), pathway = character())
@@
 for (i in 1:nrow(snp_nsgenes)) {
-  #work using dataframes, not lists
+  #use dataframes, not lists
   snp_nspath <- snp_nspath %>%
@@
 
-# Remove NA cells and significance: ie. BAOFFNEK_01342 not found
+# Filter NS genes not called in dndscv
+#omit significance (pmis_cv < 0.05)
+#discard nitrogen (not enough genes)
+#change %2C to comma
 snp_nsgenes <- snp_nsgenes %>%
-  dplyr::filter(!is.na(pathway) & !(is.na(wmis_cv))) %>%
-  dplyr::filter(wmis_cv != 0 & pmis_cv < 0.05)
-
-# Sort pathways and set colors
+  dplyr::filter(!(is.na(wmis_cv)) & wmis_cv < 1 & wmis_cv != 0 & pathway != "Nitrogen") %>%
+  dplyr::mutate(product = str_replace_all(product, "%2C", ","))
+
+# Save NS gene ids to file
+if (file.exists("mags-all/5-SNPs/dndscv/results/kofam/ns_genes_ids.txt") == FALSE) {
+  write.table(snp_nsgenes$gene_id, "mags-all/5-SNPs/dndscv/results/kofam/ns_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+
+# Sort pathways and set colors to plot
 snp_sort_ns <- snp_nsgenes %>%
   dplyr::group_by(pathway) %>%
-  dplyr::summarise(n = median(wmis_cv)) %>%
+  dplyr::summarise(n = mean(wmis_cv)) %>%
   dplyr::arrange(desc(n)) %>%
   as.data.frame()
-snp_sort_ns$color <- c("#8c2421", "#4f4fad", "#ca6627", "#6a9b85")
+snp_sort_ns$color <- c("#ddad3b", "#8c2421", "#666666", "#4f4fad") #"#6a9b85"
 
 # Violin plot for NS genes with global mutations
-snp_plot_nsmut <- ggviolin(snp_nsgenes, x = "pathway", y = "wmis_cv", color = "pathway",
+snp_plot_nsgenes <- ggviolin(snp_nsgenes, x = "pathway", y = "wmis_cv", color = "pathway",
                              shape = "pathway", xlab = "Pathway", ylab = "dN/dS",
@@
                              palette = snp_sort_ns$color) +
+  scale_y_continuous(limits = c(-0.2, 1.2)) +
   geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
-  scale_y_continuous(limits = c(-0.2, 1)) +
-  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2)
-ggsave("5-SNPs/dndscv/results/plot/plot_dnds_nsmut.pdf", snp_plot_nsmut, height = 9, width = 7)
-
+  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
+  stat_compare_means(method = "anova", label.y = 1.2) +
+  stat_compare_means(comparisons = list(c("Copper", "Iron"), c("Sulfur", "Nickel")),
+                     label = "p.format", method = "t.test")
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 9, width = 7)
+
+# Merge NS genes to unique products
+snp_nsgenes_sub <- snp_nsgenes %>%
+  dplyr::group_by(product, pathway) %>%
+  dplyr::summarise(dnds = mean(wmis_cv))
+snp_nsgenes_sub_col <- c("#8c2421", "#666666", "#4f4fad", "#ddad3b")
+
+# Lollipop plot of NS genes dN/dS x-axis range 0-2, y-axis PS genes vs selected CDS
+snp_plot_nsgensub <- ggdotchart(snp_nsgenes_sub, x = "product", y = "dnds", color = "pathway",
+                              xlab = "NS gene product", ylab = "dN/dS", y.text.col = FALSE,
+                              rotate = TRUE, palette = snp_nsgenes_sub_col,
+                              sorting = "descending", ggtheme = theme_pubr()) +
+  scale_y_continuous(limits = c(0, 1)) +
+  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2)
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenespath.pdf", snp_plot_nsgensub, height = 9, width = 10)
+
+# Count total SNPs per mutation type for phylum
+snp_topphyla <- snp_psgenes %>%
+  dplyr::select(gene_name, n_syn, n_mis, n_non, bin, name) %>%
+  dplyr::mutate(snps = n_syn + n_mis + n_non)
+snp_topphyla <- snp_topphyla %>%
+  dplyr::group_by(bin) %>%
+  dplyr::summarise(sum_nsyn = sum(n_syn),
+                   sum_nmis = sum(n_mis),
+                   sum_nnon = sum(n_non))
+snp_topphyla <- left_join(snp_topphyla, snp_dnds, by = c("bin")) %>%
+  dplyr::select(-c("num", "genus", "species")) %>%
+  pivot_longer(cols = c(sum_nsyn, sum_nmis, sum_nnon, predicted_genes),
+               names_to = "muts", values_to = "nmuts")
+#top 3 phyla and merge
+snp_topphyla <- snp_topphyla %>%
+  dplyr::filter(phylum == "Proteobacteria" | phylum == "Acidobacteriota" |
+                  phylum == "Actinobacteriota") %>%
+  dplyr::group_by(phylum, muts) %>%
+  dplyr::summarise(tot = sum(nmuts), .groups = "drop")
+
+# Barplot #snps/#genes
+snp_plot_topphy <- ggplot(snp_topphyla, aes(x = reorder(phylum, -tot), y = tot, fill = muts)) +
+  geom_col(position = position_stack(reverse = TRUE)) +
+  #geom_text(aes(label = tot), position = position_stack(reverse = TRUE), size = 2, vjust = 0.5) +
+  scale_fill_manual(name = "", labels = c("gen", "mis", "non", "syn"),
+                    values = c("#cba993", "#5db4b2", "#b687b6", "#2374a4")) +
+  scale_y_continuous(limits = c(0, 70000)) +
+  labs(x = "Phylum", y = "Number of PS mutations per type") +
+  theme_classic() +
+  theme(legend.position = c(0.85, 0.9),
+        strip.placement = "outside")
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_topphyla.pdf", snp_plot_topphy, height = 7, width = 5)
+
+### ###
 # Merge PS and NS gene names
-snp_gmerge <- bind_rows(snp_psgenes, snp_nsgenes)
-snp_gmerge <- snp_gmerge %>%
-  dplyr::mutate(pathway = ifelse(is.na(pathway), "PS", pathway))
+#snp_gmerge <- bind_rows(snp_psgenes, snp_nsgenes)
+#snp_gmerge <- snp_gmerge %>% dplyr::mutate(pathway = ifelse(is.na(pathway), "PS", pathway))
 #snp_color_gm <- c("#ddad3b", "#4f4fad", "#6a9b85", "#ca6627", "#8c2421", "black")
-snp_color_gm <- c("#8c2421", "#4f4fad", "#6a9b85", "#ca6627")
-
-# Lollipop plot dN/dS x-axis range 0-2, y-axis PS genes vs selected CDS
-snp_plot_gmerge <- ggdotchart(snp_nsgenes, x = "gene_name", y = "wmis_cv", color = "pathway",
-                              xlab = "Gene CDS", ylab = "dN/dS", y.text.col = FALSE, rotate = TRUE,
-                              palette = snp_color_gm, sorting = "descending", ggtheme = theme_pubr()) +
-  scale_y_continuous(limits = c(0, 0.4)) +
-  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
-  theme(axis.text.y = element_blank(),
-        axis.ticks.y = element_blank())
-ggsave("5-SNPs/dndscv/results/plot/plot_dnds_nspath.pdf", snp_plot_gmerge, height = 9, width = 7)
-
-# Genes for enrichment (we need KEGG KOs..)
-#snp_genenrich <- snp_psgenes$dndsr
-#names(snp_genenrich) <- snp_psgenes$name
-#snp_kegg <- enrichKEGG(snp_genenrich, "hsa", pvalueCutoff = 0.05)
-
+### ###
+
+# PS genes for enrichment (we need KEGG KOs from KOfamScan)
+snp_psenrich <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
+
+# Reduce to unique KOs, saving scores (eq. to foldchange?)
+snp_psenrich <- snp_psenrich %>%
+  dplyr::filter(V1 == "*") %>%
+  dplyr::group_by(V3, V7) %>%
+  dplyr::summarise(sum = n(), score = mean(log2(V4)), .groups = "drop") %>%
+  dplyr::rename(KO = V3, KO_definition = V7)
+#create a named vector
+#snp_kegglist <- snp_psenrich$score
+#names(snp_kegglist) <- snp_psenrich$KO
+
+# Get KEGG KO & pathway data
+snp_keggdata <- download_KEGG(species = "ko")
+snp_keggpat <- snp_keggdata$KEGGPATHID2NAME
+snp_keggkos <- snp_keggdata$KEGGPATHID2EXTID
+rm(snp_keggdata)
+
+# Enrich the genes by KO
+#https://www.biostars.org/p/494567/
+snp_keggenrich <- enricher(gene = snp_psenrich$KO, TERM2GENE = snp_keggkos,
+                           TERM2NAME = snp_keggpat, pAdjustMethod = "BH", pvalueCutoff = 0.05)
+
+# Dotplot
+pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichdot.pdf", height = 9, width = 7)
+dotplot(snp_keggenrich, showCategory = 30)
+dev.off()
+
+# Barplot
+pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichbar.pdf", height = 9, width = 7)
+barplot(snp_keggenrich, showCategory = 30)
+dev.off()
+
+# Category netplot
+pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichnet.pdf", height = 9, width = 13)
+cnetplot(snp_keggenrich, showCategory = 30, categorySize = "pvalue", foldChange = snp_psenrich$KO)
+dev.off()
+
+# Upset plot (overlaps among different genes)
+upsetplot(snp_keggenrich)
 ```
@@
 ```{r fig3}
+{ { anova_boxplot_cu / anova_boxplot_fe +
+    plot_layout(ncol = 1, widths = c(0.75)) } | anova_boxplot_s / kegg_bubbles_plot }
 # Merge plots using patchwork
-fig_3 <- { { anova_boxplot_cu / anova_boxplot_fe +
-    plot_layout(ncol = 1, widths = c(0.75)) } | anova_boxplot_s / kegg_bubbles_plot } +
-  plot_annotation(tag_levels = list(c("a", "b", "c", "d"))) &
+fig_3 <- { { snp_plot_dndsglob + snp_plot_psgenes + snp_plot_topphy +
+    plot_layout(ncol = 3, widths = c(0.25, 0.75, 0.75)) } |
+    snp_plot_nsgenes + snp_plot_nsgensub } +
+  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "e"))) &
   theme(plot.tag =  element_text(face = "bold"))
 ggsave("fig_3.pdf", fig_3, height = 14, width = 18)
-ggsave("fig_3.png", fig_3, height = 14, width = 18)
+
+#fig_s3 <- {snp_plot_psbins}
 ```

### 2024-05-10 data/dndscv/parse_cds.py

@@ -1,0 +1,66 @@
+#!/usr/bin/env python3
+"""
+Take as input a GFF3 file (ie. from Bakta) and parse the annotation of genes,
+printing a tab-delimited CDS table with defined columns as output.
+
+python3 parse_cds.py data/S15_bins/genes.gff > cds_table.txt
+"""
+
+import sys
+
+# Read GFF file from path
+gff_name = sys.argv[1]
+gff_file = open(gff_name, "r")
+
+# Lists for output..
+i = 0
+j = 0
+chrom = []
+chrom_start = []
+chrom_stop = []
+cds_start = []
+#cds_stop = []
+length = []
+strand = []
+gene_id = []
+gene_name = []
+
+# Read CDS annotation: genes, chrom (bins), positions, etc.
+# Since dNdScv is focused on eukaryotic genomes, we need to make adjustments for bacteria (no splicing)
+#https://github.com/im3sanger/dndscv/issues/62
+for line in gff_file:
+	line = line.strip()
+	#NODE_1370_length_37472_cov_9.854745	Prodigal:002006	CDS	1078	3408	.	+	0	ID=MBCHOMPH_01850;inference=ab initio prediction:Prodigal:002006;locus_tag=MBCHOMPH_01850;product=hypothetical protein
+	if line[:5] == "NODE_":
+		#i += 1
+		data = line.split("\t")
+		# Only features with type=CDS
+		if data[2] == "CDS":
+			#j += 1
+			chrom.append(data[0]) #chrom name
+			chrom_start.append(data[3]) #chrom start
+			chrom_stop.append(data[4]) #chrom stop
+			cds_start.append("1") #CDS start
+			#pos = data[0].split("_") #get max pos from range
+			#cds_stop.append(pos[3]) #chrom stop (from gff ##sequence-region, last col)
+			l = int(data[4]) - int(data[3]) + 1 #cast str to int
+			length.append(l)
+			if data[6] == "+": #strand + or -
+				strand.append("1")
+			else:
+				strand.append("-1")
+			att = data[8].split(";") #parse to ID=gene.name; Name=name;product=hypothetical|description
+			gene_id.append(att[0][3:])
+			ini = [i for i in range(len(att)) if att[i].startswith("Name=")] #get index
+			if ini:
+				gene_name.append(att[ini[0]][5:])
+			else:
+				gene_name.append(att[0][3:])
+
[…]

### 2024-05-10 data/dndscv/parse_snpeff.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python3
+"""
+Extract mutations from the VCF file obtained with SnpEff,
+generating a parsed file for dNdScv package as output
+
+python3 parse_snpeff.py S15.ann.vcf > S15_mutations.txt
+"""
+
+import sys
+
+# Read VCF file from path
+vcf_name = sys.argv[1]
+vcf_file = open(vcf_name, "r")
+
+i = 0
+chrom = []
+pos = []
+ref = []
+mut = []
+sid = []
+
+# Parse annotations
+#CHROM	POS	ID	REF	ALT	QUAL	FILTER	INFO	FORMAT	S15_sorted.bam
+for line in vcf_file:
+	line = line.strip()
+	if line[:2] != "##" and line[:6] != "#CHROM":
+		#i += 1
+		data = line.split("\t")
+		#keep attributes
+		chrom.append(data[0])
+		pos.append(data[1])
+		ref.append(data[3])
+		mut.append(data[4])
+		sid.append("S15") #data[2]
+		#info = data[7].split(";")
+		#print(len(info))
+		#last ; is ''ANN='', then split by '|'
+		#example: ANN=C|synonymous_variant|LOW|
+		#annot = info[-1].split("|")
+		#if info[0][:4] == "ANN=":
+
+vcf_file.close()
+
+# Print header and output columns
+print("sampleID\tchr\tpos\tref\tmut")
+for x in range(len(chrom)):
+	print(sid[x] + "\t" + chrom[x] + "\t" + pos[x] + "\t" + ref[x] + "\t" + mut[x])

### 2024-05-17 R/paper.Rmd
updated to manuscript version 1.1

@@
 suppressWarnings(library(clusterProfiler)) #gene enrichment
+suppressWarnings(library(circlize)) #circular viz
 ```
@@
 
-#fig sup 1
+#sup 1
 fig_s1 <- { plot_depth_top_phyla |
@@
 #fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.text.y = element_text(vjust = 0))
-ggsave("fig_s1.pdf", fig_s1, height = 8, width = 16)
-#ggsave("fig_s1.png", fig_s1, height = 8, width = 16)
+##use wrap_plots() to correct axis line/labels ??
+ggsave("mags-all/fig_s1.pdf", fig_s1, height = 8, width = 16)
+#ggsave("mags-all/fig_s1.png", fig_s1, height = 8, width = 16)
 ```
@@
 
-# Sort ecosystems (by median)
+# Sort ecosystems by median
 anova_sort_cu <- kegg_bubble_cu %>%
@@
                            order = anova_sort_cu$ecosystem, add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(kegg_bubble_cu$count_norm),
-             linetype = 2, linewidth = 0.2) +
+  geom_hline(yintercept = mean(kegg_bubble_cu$count_norm), linetype = 2, linewidth = 0.2) +
   labs(x = "Ecosystem", y = "KOs/Mb for copper") +
@@
                            order = anova_sort_fe$ecosystem, add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(kegg_bubble_fe$count_norm),
-             linetype = 2, linewidth = 0.2) +
+  geom_hline(yintercept = mean(kegg_bubble_fe$count_norm), linetype = 2, linewidth = 0.2) +
   labs(x = "Ecosystem", y = "KOs/Mb for iron") +
@@
                              order = anova_sort_s$ecosystem, add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(kegg_bubble_s$count_norm),
-             linetype = 2, linewidth = 0.2) +
+  geom_hline(yintercept = mean(kegg_bubble_s$count_norm), linetype = 2, linewidth = 0.2) +
   labs(x = "Ecosystem", y = "KOs/Mb for sulfur") +
@@
 
-## Figure 2
+## Figure 2 & supplementary 2
 
@@
 fig_2[[2]][[1]] <- fig_2[[2]][[1]] + theme(axis.title.y = element_text(vjust = -127))
-ggsave("fig_2.pdf", fig_2, height = 14, width = 18)
-#ggsave("fig_2.png", fig_2, height = 14, width = 18) vjust=120
+ggsave("mags-all/fig_2.pdf", fig_2, height = 14, width = 18)
+#ggsave("mags-all/fig_2.png", fig_2, height = 14, width = 18) #vjust=120
 ```
@@
 ```{r snps}
-# Get CDS annotations (pre-processed from Bakta results)
-snp_refcds <- read.csv("mags-all/5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
-
-# Buildref testing for bacteria.. (1 bin)
+# Buildref for 1 bin (testing for bacteria species..)
 #https://github.com/im3sanger/dndscv/issues/62
@@
 
-# Buildref for all bins (very slow: ~2 hrs in building object)
+# Buildref for all bins (very slow! ~2 hrs in building object)
 #buildref(cdsfile = "mags-all/5-SNPs/dndscv/buildref/cds_table.txt",
@@
 head(snp_dndsout$sel_cv) #dN/dS ratios at gene-level
-head(snp_dndsout$annotmuts)
+head(snp_dndsout$annotmuts) #list of annotated mutations
 print(snp_dndsout$globaldnds) #driver globals
 
-# Save dN/dS per-gene matrix
-#driver discovery or genes under positive selection (PS)
+# Save matrix with dN/dS: driver discovery or genes under positive selection (PS)
 snp_selcv <- snp_dndsout$sel_cv
@@
 #https://github.com/im3sanger/dndscv/issues/63
-snp_selcv %>% dplyr::filter(pmis_cv < 0.05) %>% nrow() #n=105328
 snp_global_ps <- snp_selcv %>%
@@
 
+# Barplot with global dN/dS ratios (not used- for reference only*)
+snp_plot_dndsglob <- ggplot(snp_global_dnds, aes(x = genes, y = reorder(label, genes), fill = label)) +
+  geom_col(position = "dodge2", fill = "#4c6194") +
+  geom_text(aes(label = perc2), vjust = 0.5, hjust = -0.1, size = 2) +
+  scale_x_continuous(limits = c(0, 16000)) +
+  labs(x = "Number of genes", y = "") +
+  theme_classic() +
+  theme(legend.position = "none")
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndsglob, height = 2, width = 5)
+
 # Filter genes under PS:
@@
   dplyr::mutate(dndsr = round((n_mis / n_syn), 1))
+
+# Get CDS annotations (pre-processed from Bakta results)
+snp_refcds <- read.csv("mags-all/5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
@@
     dplyr::filter(bin == b) %>%
-    dplyr::select(phylum, genus, species, predicted_genes)
+    dplyr::select(phylum, genus, species, completeness, contamination, predicted_genes)
   return(t)
@@
 }
-# Function that gets dndscv stats for NS genes
+# Function that gets dN/dS ratio and p-value for NS genes
 fx_dndsNsStats <- function(g) {
@@
 snp_taxabins <- tibble(phylum = character(), genus = character(), species = character(),
-                       completeness = numeric(), contamination = numeric())
+                       completeness = numeric(), contamination = numeric(), predicted_genes = numeric())
 for (i in 1:nrow(snp_dnds)) {
@@
 snp_dnds['species'] <- snp_taxabins$species
+snp_dnds['completeness'] <- snp_taxabins$completeness
+snp_dnds['contamination'] <- snp_taxabins$contamination
 snp_dnds['predicted_genes'] <- snp_taxabins$predicted_genes
@@
 #discard HPs (by name)
-#genes with at least 5 SNPs
+#genes with at least 4 SNPs
 #mark genes reaching significance (q < 0.05)
@@
 if (file.exists("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt") == FALSE) {
-  write.table(snp_psgenes_sub %>% dplyr::select(name) %>% distinct(name), "mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
-}
-
-# Barplot with global dN/dS ratios
-snp_plot_dndsglob <- ggplot(snp_global_dnds, aes(x = genes, y = reorder(label, genes), fill = label)) +
-  geom_col(position = "dodge2", fill = "#4c6194") +
-  geom_text(aes(label = perc2), vjust = 0.5, hjust = -0.1, size = 2) +
-  scale_x_continuous(limits = c(0, 16000)) +
-  labs(x = "Number of genes", y = "") +
-  theme_classic() +
-  theme(legend.position = "none")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndsglob, height = 2, width = 5)
-
-# Plot global driver genes by species
+  write.table(unique(snp_psgenes_sub$name), "mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+
+# Plot global driver genes by MAG/species
 snp_plot_psbins <- ggplot(snp_dnds, aes(x = num, y = reorder(bin, num), fill = bin)) +
@@
   geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
-  labs(x = "Number of genes under PS", y = "Bin") +
+  labs(x = "Number of genes under PS", y = "MAG Bin") +
   scale_x_continuous(limits = c(0, 255)) +
@@
   facet_grid(.~species)
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 8.5, width = 9.5)
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 9, width = 10)
+
+#add completeness & contamination barplots
+snp_plot_psbins2 <- ggplot(snp_dnds, aes(x = completeness, y = reorder(bin, num), fill = bin)) +
+  geom_col(fill = "#3c4d76") +
+  geom_vline(xintercept = 100, linetype = 1, linewidth = 0.2) +
+  labs(x = "Completeness", y = "") +
+  theme_classic() +
+  theme(axis.text.y = element_blank(),
+        axis.line.y = element_blank(),
+        axis.ticks.y = element_blank())
+snp_plot_psbins3 <- ggplot(snp_dnds, aes(x = contamination, y = reorder(bin, num), fill = bin)) +
+  geom_col(fill = "#763c4d") +
+  geom_vline(xintercept = max(snp_dnds$contamination), linetype = 1, linewidth = 0.2) +
+  labs(x = "Contamination", y = "") +
+  theme_classic() +
+  theme(axis.text.y = element_blank(),
+        axis.line.y = element_blank(),
+        axis.ticks.y = element_blank())
+
+#merge
+snp_plot_psbins <- wrap_plots(snp_plot_psbins, snp_plot_psbins2, snp_plot_psbins3,
+                              nrow = 1, ncol = 3, widths = c(1, 0.15, 0.15))
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 9, width = 12)
+rm(snp_plot_psbins2, snp_plot_psbins3)
 
@@
 
+#merge
 snp_plot_psgenes <- (snp_plot_psgenes | snp_plot_muts) +
@@
 
-# Load negative selected (NS) genes
+# Load negative selected (NS) genes from selected pathways
 snp_nsgenes <- read.csv("mags-all/5-SNPs/dndscv/results/selected_cds_metals.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
-# Infer pathway from product and add dnds attributes
+# Infer pathway from product and add dndscv attributes
 snp_nspath <- tibble(gene_name = character(), pathway = character())
@@
                              palette = snp_sort_ns$color) +
-  scale_y_continuous(limits = c(-0.2, 1.2)) +
+  scale_y_continuous(limits = c(-0.2, 1.3)) +
   geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
   geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
-  stat_compare_means(method = "anova", label.y = 1.2) +
-  stat_compare_means(comparisons = list(c("Copper", "Iron"), c("Sulfur", "Nickel")),
-                     label = "p.format", method = "t.test")
+  stat_compare_means(method = "anova", label.y = 1.3) +
+  stat_compare_means(comparisons = list(c("Copper", "Iron"), c("Sulfur", "Nickel"),
+                                        c("Copper", "Nickel")), label = "p.format", method = "t.test")
 ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 9, width = 7)
@@
 snp_plot_nsgensub <- ggdotchart(snp_nsgenes_sub, x = "product", y = "dnds", color = "pathway",
-                              xlab = "NS gene product", ylab = "dN/dS", y.text.col = FALSE,
+                              xlab = "NS gene product", ylab = "Mean dN/dS", y.text.col = FALSE,
                               rotate = TRUE, palette = snp_nsgenes_sub_col,
@@
   geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2)
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenespath.pdf", snp_plot_nsgensub, height = 9, width = 10)
-
-# Count total SNPs per mutation type for phylum
-snp_topphyla <- snp_psgenes %>%
-  dplyr::select(gene_name, n_syn, n_mis, n_non, bin, name) %>%
-  dplyr::mutate(snps = n_syn + n_mis + n_non)
-snp_topphyla <- snp_topphyla %>%
-  dplyr::group_by(bin) %>%
-  dplyr::summarise(sum_nsyn = sum(n_syn),
-                   sum_nmis = sum(n_mis),
-                   sum_nnon = sum(n_non))
-snp_topphyla <- left_join(snp_topphyla, snp_dnds, by = c("bin")) %>%
-  dplyr::select(-c("num", "genus", "species")) %>%
-  pivot_longer(cols = c(sum_nsyn, sum_nmis, sum_nnon, predicted_genes),
-               names_to = "muts", values_to = "nmuts")
-#top 3 phyla and merge
-snp_topphyla <- snp_topphyla %>%
-  dplyr::filter(phylum == "Proteobacteria" | phylum == "Acidobacteriota" |
-                  phylum == "Actinobacteriota") %>%
-  dplyr::group_by(phylum, muts) %>%
-  dplyr::summarise(tot = sum(nmuts), .groups = "drop")
-
-# Barplot #snps/#genes
-snp_plot_topphy <- ggplot(snp_topphyla, aes(x = reorder(phylum, -tot), y = tot, fill = muts)) +
-  geom_col(position = position_stack(reverse = TRUE)) +
-  #geom_text(aes(label = tot), position = position_stack(reverse = TRUE), size = 2, vjust = 0.5) +
-  scale_fill_manual(name = "", labels = c("gen", "mis", "non", "syn"),
-                    values = c("#cba993", "#5db4b2", "#b687b6", "#2374a4")) +
-  scale_y_continuous(limits = c(0, 70000)) +
-  labs(x = "Phylum", y = "Number of PS mutations per type") +
-  theme_classic() +
-  theme(legend.position = c(0.85, 0.9),
-        strip.placement = "outside")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_topphyla.pdf", snp_plot_topphy, height = 7, width = 5)
-
-### ###
-# Merge PS and NS gene names
-#snp_gmerge <- bind_rows(snp_psgenes, snp_nsgenes)
-#snp_gmerge <- snp_gmerge %>% dplyr::mutate(pathway = ifelse(is.na(pathway), "PS", pathway))
-#snp_color_gm <- c("#ddad3b", "#4f4fad", "#6a9b85", "#ca6627", "#8c2421", "black")
-### ###
-
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenesprod.pdf", snp_plot_nsgensub, height = 9, width = 10)
+
+# Global gene estimation of dn/ds (<1 & >1) per number of genes and phylum
+snp_estim <- snp_selcv %>%
+  dplyr::select(gene_name, wmis_cv, pmis_cv) %>%
+  dplyr::filter(wmis_cv != 0)
+
+# Populate global dataset (~2 min)
+snp_gene2contig <- vector()
+snp_contig2bin <- vector()
+snp_genenames <- vector()
+for (i in 1:nrow(snp_estim)) {
+	snp_gene2contig <- append(snp_gene2contig, fx_dndsGene2Contig(snp_estim$gene_name[i]))
+	snp_contig2bin <- append(snp_contig2bin, fx_dndsContig2Bin(snp_gene2contig[i]))
+	snp_genenames <- append(snp_genenames, fx_dndsGeneName(snp_estim$gene_name[i]))
+}
+snp_estim['contig'] <- snp_gene2contig
+snp_estim['bin'] <- snp_contig2bin
+snp_estim['name'] <- snp_genenames
+rm(snp_refcds, snp_bins, snp_genids, snp_gene2contig, snp_contig2bin, snp_genenames) #clean memory
+
+# Add taxa & QC
+snp_estim <- left_join(snp_estim, snp_dnds, by = c("bin")) %>%
+  dplyr::select(-c("contig", "num", "genus", "species"))
+
+# Normalize dN/dS per gene by MAG contamination, then group by phylum
+snp_estim_phy <- snp_estim %>%
+  dplyr::mutate(dnds_adj = wmis_cv * (1 - contamination / 100)) %>%
+  dplyr::group_by(phylum, bin) %>%
+  dplyr::summarise(snps = n(), dnds_sum = sum(dnds_adj)) %>%
+  ungroup() %>%
+  dplyr::mutate(dnds_glob = dnds_sum / snps) %>% #divide by total genes and top 3 phyla
+  dplyr::filter(phylum == "Proteobacteria" | phylum == "Acidobacteriota" | phylum == "Actinobacteriota")
+
+# Sort phylum by median and set color
+snp_estim_sort_phy <- snp_estim_phy %>%
+  dplyr::group_by(phylum) %>%
+  dplyr::summarise(n = median(dnds_glob)) %>%
+  dplyr::arrange(desc(n))
+snp_estim_sort_phy$color <- rep("#4c6194", 3)
+
+# Boxplot global dN/dS ratios by phylum
+snp_plot_estim_phy <- ggboxplot(snp_estim_phy, x = "phylum", y = "dnds_glob", color = "phylum",
+                                order = snp_estim_sort_phy$phylum, palette = snp_estim_sort_phy$color,
+                                add = "jitter", legend = "none") +
+  geom_hline(yintercept = mean(snp_estim_phy$dnds_glob), linetype = 2, linewidth = 0.2) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
+  labs(x = "Phylum", y = "Global normalized dN/dS ratio") +
+  stat_compare_means(method = "anova", label.y = 1.6) +
+  stat_compare_means(comparisons = list(c("Proteobacteria", "Actinobacteriota"),
+                                        c("Proteobacteria", "Acidobacteriota"),
+                                        c("Actinobacteriota", "Acidobacteriota")),
+                     label = "p.format", method = "t.test")
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_estim_phy.pdf", snp_plot_estim_phy, height = 9, width = 7)
+
+# Aggregate normalized dN/dS ratios by bin
+snp_estim_bin <- snp_estim %>%
+  dplyr::mutate(dnds_adj = wmis_cv * (1 - contamination / 100)) %>%
+  dplyr::group_by(phylum, bin) %>%
+  dplyr::summarise(snps = n(), dnds_sum = sum(dnds_adj)) %>%
+  ungroup() %>%
+  dplyr::mutate(dnds_glob = dnds_sum / snps) #divide by total genes
+
+# Sort bins
+snp_estim_sort_bin <- snp_estim_bin %>% dplyr::arrange(dnds_glob)
+
+# Barplot global dN/dS ratios by MAG bins
+snp_plot_estim_bin <- ggbarplot(snp_estim_bin, x = "bin", y = "dnds_glob", color = "phylum",
+                                order = snp_estim_sort_bin$bin, palette = tree_colors,
+                                fill = "phylum", rotate = TRUE) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
+  labs(x = "MAG bin", y = "Global normalized dN/dS ratio")
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_estim_bin.pdf", snp_plot_estim_bin, height = 9, width = 8)
+
+# Load lists of PS and NS genes (with KEGG KOs from KOfamScan)
+snp_kops <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
+snp_kons <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ns_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
+
+# Reduce to unique KOs, saving scores
+snp_kops <- snp_kops %>%
+  dplyr::group_by(V3, V7) %>%
+  dplyr::summarise(sum_ko = n(), score_avg = mean(V4), .groups = "drop") %>%
+  dplyr::rename(KO = V3, KO_definition = V7) %>%
+  dplyr::mutate(KO_def = paste0("(", KO, ") - ", KO_definition)) %>%
+  dplyr::mutate(dNdS = "PS")
+snp_kons <- snp_kons %>%
+  dplyr::group_by(V3, V7) %>%
+  dplyr::summarise(sum_ko = n(), score_avg = mean(V4), .groups = "drop") %>%
+  dplyr::rename(KO = V3, KO_definition = V7) %>%
+  dplyr::mutate(KO_def = paste0("(", KO, ") - ", KO_definition)) %>%
+  dplyr::mutate(dNdS = "NS")
+#merge dataframe and filter
+snp_kos <- bind_rows(snp_kops, snp_kons)
+snp_kos <- snp_kos %>%
+  dplyr::filter(sum_ko > 10)
+rm(snp_kops, snp_kons) #clean memory
+
+# Histogram for KO categories by dN/dS
+snp_plot_kos <- ggbarplot(snp_kos, x = "KO_def", y = "sum_ko", fill = "dNdS",
+                          color = "white", palette = "aaas", sort.val = "asc",
+                          rotate = TRUE, xlab = "KO definition", ylab = "Aggregated KOs")
+ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_kos.pdf", snp_plot_kos, height = 9, width = 12)
+
+# Filter genome bin with PS selection
+snp_cirmag <- snp_estim %>% dplyr::filter(grepl("S15.49", bin))
+
+# Set genomic attributes: gc, contigs
+snp_cirmagname <- "SPAdes-MetaBAT2Refined-S15.49"
+snp_cirsize <- c(0, 2703949)
+snp_cirran <- c(0, 0.5, 1, 1.5, 2, 2.5, 2.7) * 10^6
+
+# Export gene names
+if (file.exists("mags-all/5-SNPs/dndscv/results/circos/S15.49_genes.txt") == FALSE) {
+  write.table(snp_cirmag$gene_name, "mags-all/5-SNPs/dndscv/results/circos/S15.49_genes.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+
+# Read parsed gene annotation (*externally processed*)
+#chr	start	end	gene_name
+#NODE_4971_length_14023_cov_250.866767	2403	4742	BNHKIBDC_01907
+snp_cirgenes <- read.csv("mags-all/5-SNPs/dndscv/results/circos/S15.49_genes_pos.txt", sep = "\t", stringsAsFactors = FALSE)
+
+# Merge with genome selection (chr start end dnds gene)
+snp_cirgenes <- snp_cirgenes %>% dplyr::mutate(chr = snp_cirmagname)
+snp_cirgenes <- left_join(snp_cirgenes, snp_cirmag, by = c("gene_name"))
+snp_cirgenes <- snp_cirgenes %>%
+  dplyr::select(-c("pmis_cv", "bin", "phylum", "completeness", "contamination", "predicted_genes"))
+
+# Circos plot
+circos.clear()
+pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_circos.pdf", height = 5, width = 5)
+circos.par(cell.padding = c(0, 0))
+circos.initialize(sectors = snp_cirmagname, xlim = snp_cirsize)
+circos.track(snp_cirmagname, ylim = snp_cirsize, panel.fun = function(x, y) {
+  #circos.text(snp_cirsize[1], snp_cirsize[2], cex = 0.5, adj = c(-0.05, -1.5),
+  #            snp_cirmagname, facing = "bending.inside", niceFacing = TRUE)
+  text(0, 0, snp_cirmagname, cex = 0.65)
+  text(0, -0.08, paste("phylum Actinobacteriota | genus Acidithrix", 
+                       "168 contigs | N50 26686 | GC 47.7", sep = "\n"), cex = 0.55)
+  circos.axis(h = "top", major.at = snp_cirran, labels = paste0(round(snp_cirran/10^6, 1), " Mb"),
+              labels.cex = 0.35, lwd = 0.7, labels.facing = "clockwise")
+}, bg.col = "#e2edf2", bg.border = FALSE, track.height = 0.06)
+circos.genomicLabels(snp_cirgenes, labels.column = 6, cex = 0.35, line_lwd = 0.5,
+                     side = "inside", line_col = "black", col = "black",
+                     connection_height = 0.05, labels_height = 0.06)
+dev.off()
+
+### Enrichment ###
 # PS genes for enrichment (we need KEGG KOs from KOfamScan)
@@
 
-# Reduce to unique KOs, saving scores (eq. to foldchange?)
+# Reduce to unique KOs, saving scores (converted to foldchange)
 snp_psenrich <- snp_psenrich %>%
@@
 
-## Figure 3
+## Figure 3 & supplementary 3
 
 ```{r fig3}
-{ { anova_boxplot_cu / anova_boxplot_fe +
-    plot_layout(ncol = 1, widths = c(0.75)) } | anova_boxplot_s / kegg_bubbles_plot }
 # Merge plots using patchwork
-fig_3 <- { { snp_plot_dndsglob + snp_plot_psgenes + snp_plot_topphy +
-    plot_layout(ncol = 3, widths = c(0.25, 0.75, 0.75)) } |
-    snp_plot_nsgenes + snp_plot_nsgensub } +
-  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "e"))) &
+fig_3a <- wrap_plots(plot_spacer(), snp_plot_psgenes, snp_plot_estim_phy, plot_spacer(),
+                     nrow = 1, ncol = 4, widths = c(0.1, 0.85, 0.5, 0.4))
+fig_3b <- wrap_plots(snp_plot_estim_bin, snp_plot_nsgenes, snp_plot_nsgensub,
+                     nrow = 1, widths = c(1, 0.75, 1))
+fig_3 <- wrap_plots(fig_3a, fig_3b, nrow = 2) +
+  plot_annotation(tag_levels = list(c("a", "", "b", "c", "d", "e"))) &
   theme(plot.tag =  element_text(face = "bold"))
-ggsave("fig_3.pdf", fig_3, height = 14, width = 18)
-
-#fig_s3 <- {snp_plot_psbins}
+ggsave("mags-all/fig_3.pdf", fig_3, height = 20, width = 24)
+#ggsave("mags-all/fig_3.png", fig_3, height = 20, width = 24)
+rm(fig_3a, fig_3b)
+
+#sup 3
+fig_s3 <- wrap_plots(snp_plot_psbins, snp_plot_kos, nrow = 1, widths = c(1, 0.5)) +
+  plot_annotation(tag_levels = list(c("a", "", "", "b"))) &
+  theme(plot.tag =  element_text(face = "bold"))
+ggsave("mags-all/fig_s3.pdf", fig_s3, height = 10, width = 26)
+ggsave("mags-all/fig_s3.png", fig_s3, height = 10, width = 26)
 ```

### 2024-05-17 R/parsers/checkm_bins.R

@@ -1,67 +1,0 @@
-#!/usr/bin/env Rscript
-
-#---------------------------------------------------------
-# Systemix MAGs: Get the QC from a list of -refined- bins
-# 
-# Moises A. Rojas 06/11/2023 - mod. 06/03/2024
-#---------------------------------------------------------
-
-# Set working dir
-setwd("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/2-statistics/")
-
-# Read bin list
-bins <- read.csv("bins_refined_list.tsv", sep = "\t", stringsAsFactors = FALSE)
-selected_bins <- bins$bin
-
-# QC parameters to keep
-completeness  <- vector()
-contamination <- vector()
-strain_het    <- vector()
-genome_size   <- vector()
-contigs       <- vector()
-n50_contigs   <- vector()
-gc            <- vector()
-pred_genes    <- vector()
-
-# Read the QC file from CheckM
-checkm_qc <- read.csv("checkm_qa.tsv", sep = "\t", stringsAsFactors = FALSE)
-
-# Save attributes for each bin
-checkm_bin <- checkm_qc$bin
-checkm_com <- checkm_qc$completeness
-checkm_con <- checkm_qc$contamination
-checkm_st  <- checkm_qc$strain_heterogeneity
-checkm_gs  <- checkm_qc$genome_size
-checkm_nco <- checkm_qc$contigs
-checkm_n50 <- checkm_qc$N50_contigs
-checkm_gc  <- checkm_qc$gc
-checkm_pg  <- checkm_qc$predicted_genes
-
-# Match positions for each bin, saving attributes
-for (i in 1:length(selected_bins)) { #428
-	for (j in 1:length(checkm_bin)) { #1840
-		if (selected_bins[i] == checkm_bin[j]) {
-			#print(j) #pos
-			completeness  <- append(completeness,  checkm_com[j])
-			contamination <- append(contamination, checkm_con[j])
-			strain_het    <- append(strain_het,    checkm_st[j])
-			genome_size   <- append(genome_size,   checkm_gs[j])
-			contigs       <- append(contigs,       checkm_nco[j])
-			n50_contigs   <- append(n50_contigs,   checkm_n50[j])
-			gc            <- append(gc,            checkm_gc[j])
-			pred_genes    <- append(pred_genes,    checkm_pg[j])
-		}
-	}
-}
-
-# Create output dataframe and save to file
[…]

### 2024-05-17 R/parsers/get_gtdb_taxa.R

@@ -1,69 +1,0 @@
-#!/usr/bin/env Rscript
-
-#-----------------------------------------------
-# Populate the GTDB-Tk of MAGs from a reference
-# Moises A. Rojas 13/10/2023
-#-----------------------------------------------
-
-setwd("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/mags-all/analysis/1-taxonomy")
-
-# Read the list of bins (last nf-core/mag run: mags-all/)
-bins_all <- read.csv("bins_refined_list.tsv", sep = "\t", stringsAsFactors = FALSE)
-head(bins_all)
-
-bins_genome <- bins_all$bin
-
-# Read the GTDB-Tk classification file
-gtdb_file <- read.csv("gtdbtk_summary.tsv", sep = "\t", stringsAsFactors = FALSE)
-head(gtdb_file)
-
-# Save attributes
-gtdb_genome <- gtdb_file$user_genome
-gtdb_taxa   <- gtdb_file$classification
-
-# Set output attributes
-bin     <- vector()
-taxa    <- vector()
-kingdom <- vector()
-phylum  <- vector()
-class   <- vector()
-order   <- vector()
-family  <- vector()
-genus   <- vector()
-species <- vector()
-
-# Parse taxonomy array:
-# d__Bacteria;p__Nitrospirota_A;c__Leptospirillia;o__Leptospirillales;f__Leptospirillaceae;g__;s__
-taxa_split <- strsplit(gtdb_taxa, split = ";", fixed = TRUE)
-
-# Match positions for each bin
-# Then, save taxa attributes
-for (i in 1:length(bins_genome)) { # 428
-	for (j in 1:length(gtdb_genome)) { # 377
-		if (bins_genome[i] == gtdb_genome[j]) {
-			#print(j) #pos
-			bin     <- append(bin,     gtdb_genome[[j]])
-			taxa    <- append(taxa,    gtdb_taxa[[j]])
-			kingdom <- append(kingdom, taxa_split[[j]][1])
-			phylum  <- append(phylum,  taxa_split[[j]][2])
-			class   <- append(class,   taxa_split[[j]][3])
-			order   <- append(order,   taxa_split[[j]][4])
-			family  <- append(family,  taxa_split[[j]][5])
-			genus   <- append(genus,   taxa_split[[j]][6])
-			species <- append(species, taxa_split[[j]][7])
-		}
-	}
-}
-
-# Create output dataframe and save to file
[…]

### 2024-06-03 R/paper.Rmd
updated to manuscript version 1.2

@@
 mags_depth_spec_eco <- left_join(mags_depth_spec_eco, mags_depth_spec_eco_n, by = c("ecosystem"))
-rm(mags_depth_spec_eco_n)
+rm(mags_depth_spec_eco_n) #clean memory
 
@@
 ### --EXPLORE OTHER TAXA-- ###
-#inspect unannotated family-level bins (uFGBs)? #1166
-#what about novel bacteria at genus level?
-mags_depth_novel_gen <- mags_metadata_bac %>%
-  dplyr::filter(genus == "") %>%
-  dplyr::group_by(phylum, genus, ecosystem) %>%
-  dplyr::summarise(gen_unk = n()) %>%
-  #dplyr::arrange(desc(gen)) %>%
-  ungroup() %>%
-  dplyr::group_by(genus) %>%
-  dplyr::mutate(tot_gen = sum(gen_unk)) %>%
-  dplyr::mutate(perc = gen_unk * 100 / tot_gen)
-
-#get unique classes of unknown genus by ecosystem
-#Leptospirillia, Sulfobacillia
-mags_md_unk_gen2 <- mags_metadata_bac %>%
-  dplyr::filter(genus == "") %>%
-  dplyr::group_by(class, ecosystem) %>%
-  tally() %>%
-  #slice(n()) %>%
-  ungroup()
-unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Biomining", "class"])
-
-#unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
-unique(mag_samples[mag_samples[, "family"] != "", "genus"])
-mags_metadata_bac %>% filter(grepl("Acidithrix", genus, ignore.case = TRUE) == TRUE)
-
-plot_md_unk_cla <- ggplot(mags_md_unk_gen2[mags_md_unk_gen2$n > 10, ],
-                          aes(x = reorder(class, n), y = n, fill = ecosystem)) +
-  geom_bar(stat = "identity") +
-  labs(x = "Class", y = "Count") +
-  scale_y_continuous(limits = c(0, 250)) +
-  theme_minimal() +
-  coord_flip() +
-  facet_wrap(~ecosystem)
-ggsave("plot_md_unk_class_eco.pdf", plot_md_unk_cla, height = 7, width = 10)
-
-#class of interest for acidic soils:
-#Alphaproteobacteria, Gammaproteobacteria
-#Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
-table(mags_metadata_bac[mags_metadata_bac[, "phylum"] == "Proteobacteria", "class"])
-table(mags_metadata_bac[mags_metadata_bac[, "class"] == "Acidimicrobiia", "order"])
-
-#family of interest reported for Systemix [ref 2022]:
-#Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
-mags_md_fams_acid <- mags_metadata_bac %>%
-  dplyr::filter(family == "Pseudomonadaceae") %>%
-  dplyr::group_by(ecosystem) %>%
-  dplyr::summarise(count = n())
+##inspect unannotated family-level bins (uFGBs)? #1166
+##what about novel bacteria at genus level?
+#mags_depth_novel_gen <- mags_metadata_bac %>%
+#  dplyr::filter(genus == "") %>%
+#  dplyr::group_by(phylum, genus, ecosystem) %>%
+#  dplyr::summarise(gen_unk = n()) %>%
+#  #dplyr::arrange(desc(gen)) %>%
+#  ungroup() %>%
+#  dplyr::group_by(genus) %>%
+#  dplyr::mutate(tot_gen = sum(gen_unk)) %>%
+#  dplyr::mutate(perc = gen_unk * 100 / tot_gen)
+#
+##get unique classes of unknown genus by ecosystem
+##Leptospirillia, Sulfobacillia
+#mags_md_unk_gen2 <- mags_metadata_bac %>%
+#  dplyr::filter(genus == "") %>%
+#  dplyr::group_by(class, ecosystem) %>%
+#  tally() %>%
+#  #slice(n()) %>%
+#  ungroup()
+#unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Biomining", "class"])
+#
+##unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
+#unique(mag_samples[mag_samples[, "family"] != "", "genus"])
+#mags_metadata_bac %>% filter(grepl("Acidithrix", genus, ignore.case = TRUE) == TRUE)
+#
+#plot_md_unk_cla <- ggplot(mags_md_unk_gen2[mags_md_unk_gen2$n > 10, ],
+#                          aes(x = reorder(class, n), y = n, fill = ecosystem)) +
+#  geom_bar(stat = "identity") +
+#  labs(x = "Class", y = "Count") +
+#  scale_y_continuous(limits = c(0, 250)) +
+#  theme_minimal() +
+#  coord_flip() +
+#  facet_wrap(~ecosystem)
+#ggsave("plot_md_unk_class_eco.pdf", plot_md_unk_cla, height = 7, width = 10)
+#
+##class of interest for acidic soils:
+##Alphaproteobacteria, Gammaproteobacteria
+##Acidobacteriae, Thermoleophilia, Sulfobacillia, Acidimicrobiia
+#table(mags_metadata_bac[mags_metadata_bac[, "phylum"] == "Proteobacteria", "class"])
+#table(mags_metadata_bac[mags_metadata_bac[, "class"] == "Acidimicrobiia", "order"])
+#
+##family of interest reported for Systemix [ref 2022]:
+##Pseudomonadaceae (mayor), Flavobacteriaceae, Erwiniaceae
+#mags_md_fams_acid <- mags_metadata_bac %>%
+#  dplyr::filter(family == "Pseudomonadaceae") %>%
+#  dplyr::group_by(ecosystem) %>%
+#  dplyr::summarise(count = n())
 ```
@@
 # Read QC file from CheckM
-mags_qc <- read.csv("2-statistics/checkm_parsed_bins.tsv", sep = "\t")
-
-# Select current sample
+mags_qc <- read.csv("mags-all/2-statistics/checkm_parsed_bins.tsv", sep = "\t")
+
+# Select S15 sample
 mags_qc <- mags_qc %>% filter(grepl("Refined-S15", bin))
@@
   dplyr::filter(completeness >= 90 & contamination <= 5) %>%
-  dplyr::summarise(mags = n(), contigs = sum(contigs))
+  dplyr::summarise(mags = n(), contigs = sum(contigs),
+                   avg_comp = mean(completeness), avg_cont = mean(contamination))
 mags_qc_hq$label = "High Quality"
-
 mags_qc_mq <- mags_qc %>%
   dplyr::filter(completeness < 90 & completeness >= 50 & contamination <= 10) %>%
-  dplyr::summarise(mags = n(), contigs = sum(contigs))
+  dplyr::summarise(mags = n(), contigs = sum(contigs),
+                   avg_comp = mean(completeness), avg_cont = mean(contamination))
 mags_qc_mq$label = "Medium Quality"
-
 mags_qc_lq <- mags_qc %>%
   dplyr::filter(completeness < 50 & contamination <= 10) %>%
-  dplyr::summarise(mags = n(), contigs = sum(contigs))
+  dplyr::summarise(mags = n(), contigs = sum(contigs),
+                   avg_comp = mean(completeness), avg_cont = mean(contamination))
 mags_qc_lq$label = "Low Quality"
-
+#merge
 mags_qcs <- rbind(mags_qc_hq, mags_qc_mq, mags_qc_lq)
-rm(mags_qc_hq, mags_qc_mq, mags_qc_lq)
+rm(mags_qc_hq, mags_qc_mq, mags_qc_lq) #clean memory
 
@@
   theme(legend.position = "none")
-ggsave("plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 3, width = 5)
+ggsave("mags-all/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 3, width = 5)
 ```
@@
 }
-rm(tree)
+rm(tree) #clean memory
 
@@
                      Genome_Size = tree_gs, GC = tree_gc, Pred_Genes = tree_pg)
-rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg)
+rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg) #clean memory
 
@@
 #ggsave("fig_1.png", fig_1, height = 14, width = 14)
-rm(fig_1a, fig_1b)
+rm(fig_1a, fig_1b) #clean memory
 
@@
 #fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.text.y = element_text(vjust = 0))
-##use wrap_plots() to correct axis line/labels ??
-ggsave("mags-all/fig_s1.pdf", fig_s1, height = 8, width = 16)
-#ggsave("mags-all/fig_s1.png", fig_s1, height = 8, width = 16)
+#fig_s1 <- wrap_plots(plot_depth_usgbs, plot_depth_spec, nrow = 2, ncol = 1, heights = c(0.85, 0.15))
+ggsave("mags-all/fig_s1.pdf", fig_s1, height = 9, width = 18)
+#ggsave("mags-all/fig_s1.png", fig_s1, height = 9, width = 18)
 ```
@@
 #print(RefCDS[[1]])
-#rm(gr_genes, RefCDS)
+#rm(gr_genes, RefCDS) #clean memory
 
@@
 #load("mags-all/5-SNPs/dndscv/run/S15_substmodel.rda")
-rm(snp_mutations)
+rm(snp_mutations) #clean memory
 
@@
 snp_selcv <- snp_dndsout$sel_cv
-rm(snp_dndsout)
+rm(snp_dndsout) #clean memory
 
@@
 snp_dnds['predicted_genes'] <- snp_taxabins$predicted_genes
-rm(snp_taxabins)
+rm(snp_taxabins) #clean memory
 
@@
 # Subset PS genes to plot mutations -> synonymous/missense (non-syn.)/nonsense
-#discard HPs (by name)
+#discard hypothetical proteins (HPs: according to name)
 #genes with at least 4 SNPs
@@
 
-#add completeness & contamination barplots
+# Add completeness & contamination barplots, then merge
 snp_plot_psbins2 <- ggplot(snp_dnds, aes(x = completeness, y = reorder(bin, num), fill = bin)) +
@@
 
-#merge
 snp_plot_psbins <- wrap_plots(snp_plot_psbins, snp_plot_psbins2, snp_plot_psbins3,
@@
 ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 9, width = 12)
-rm(snp_plot_psbins2, snp_plot_psbins3)
+rm(snp_plot_psbins2, snp_plot_psbins3) #clean memory
 
@@
 
-# Plot PS genes mutations per type
+# Plot PS genes mutations per type, then merge
 snp_plot_muts <- ggplot(snp_psgenes_sub, aes(x = n_mut, y = reorder(name, wmis_cv), fill = mut)) +
@@
 
-#merge
 snp_plot_psgenes <- (snp_plot_psgenes | snp_plot_muts) +
@@
 ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psgenmuts.pdf", snp_plot_psgenes, height = 11, width = 11)
-rm(snp_plot_muts)
+rm(snp_plot_muts) #clean memory
 
@@
 snp_nsgenes <- left_join(snp_nsgenes, snp_nsstat, by = c("gene_name"))
-rm(snp_nspath, snp_nsstat)
+rm(snp_nspath, snp_nsstat) #clean memory
 
 # Filter NS genes not called in dndscv
-#omit significance (pmis_cv < 0.05)
+#skip significance (pmis_cv < 0.05)
 #discard nitrogen (not enough genes)
-#change %2C to comma
 snp_nsgenes <- snp_nsgenes %>%
   dplyr::filter(!(is.na(wmis_cv)) & wmis_cv < 1 & wmis_cv != 0 & pathway != "Nitrogen") %>%
-  dplyr::mutate(product = str_replace_all(product, "%2C", ","))
+  dplyr::mutate(product = str_replace_all(product, "%2C", ",")) #change %2C to comma
 
@@
 
-# Merge NS genes to unique products
+# Merge NS genes to unique selected gene products
 snp_nsgenes_sub <- snp_nsgenes %>%
@@
 
-# Lollipop plot of NS genes dN/dS x-axis range 0-2, y-axis PS genes vs selected CDS
+# Lollipop plot of grouped NS gene products
 snp_plot_nsgensub <- ggdotchart(snp_nsgenes_sub, x = "product", y = "dnds", color = "pathway",
-                              xlab = "NS gene product", ylab = "Mean dN/dS", y.text.col = FALSE,
-                              rotate = TRUE, palette = snp_nsgenes_sub_col,
+                                xlab = "NS gene product", ylab = "Mean dN/dS", y.text.col = TRUE,
+                                rotate = TRUE, palette = snp_nsgenes_sub_col, dot.size = 3,
                                 sorting = "descending", ggtheme = theme_pubr()) +
@@
 
-# Global gene estimation of dn/ds (<1 & >1) per number of genes and phylum
+# Global gene estimation of dN/dS (PS and NS)
 snp_estim <- snp_selcv %>%
@@
 snp_estim <- left_join(snp_estim, snp_dnds, by = c("bin")) %>%
-  dplyr::select(-c("contig", "num", "genus", "species"))
+  dplyr::select(-c("num", "genus", "species"))
 
@@
   ungroup() %>%
-  dplyr::mutate(dnds_glob = dnds_sum / snps) %>% #divide by total genes and top 3 phyla
+  dplyr::mutate(dnds_glob = dnds_sum / snps) %>% #divide by total genes and get top 3 phyla
   dplyr::filter(phylum == "Proteobacteria" | phylum == "Acidobacteriota" | phylum == "Actinobacteriota")
@@
   dplyr::group_by(phylum) %>%
-  dplyr::summarise(n = median(dnds_glob)) %>%
+  dplyr::summarise(n = median(dnds_glob)) %>% #by mean?
   dplyr::arrange(desc(n))
-snp_estim_sort_phy$color <- rep("#4c6194", 3)
+snp_estim_sort_phy$color <- rep("#4c6194", nrow(snp_estim_sort_phy))
 
@@
 
-# Load lists of PS and NS genes (with KEGG KOs from KOfamScan)
+# Load lists of PS and NS genes (externally* matched against KEGG KOs from KOfamScan hits)
 snp_kops <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
@@
   dplyr::mutate(dNdS = "NS")
-#merge dataframe and filter
+
+# Merge dataframe and filter by sum
 snp_kos <- bind_rows(snp_kops, snp_kons)
-snp_kos <- snp_kos %>%
-  dplyr::filter(sum_ko > 10)
+snp_kos <- snp_kos %>% dplyr::filter(sum_ko > 10)
 rm(snp_kops, snp_kons) #clean memory
@@
 
-# Filter genome bin with PS selection
+# Circos: filter the genome bin with global dN/dS >1
 snp_cirmag <- snp_estim %>% dplyr::filter(grepl("S15.49", bin))
 
-# Set genomic attributes: gc, contigs
+# Parse contigs by node id and sort dataframe
+snp_nodeid <- vector()
+for (i in 1:nrow(snp_cirmag)) {
+  snp_nodeid <- append(snp_nodeid, str_split(snp_cirmag$contig, "_")[[i]][2])
+}
+snp_cirmag['node_id'] <- snp_nodeid
+snp_cirmag$node_id <- as.integer(snp_cirmag$node_id)
+snp_cirmag <- snp_cirmag %>% dplyr::arrange(node_id)
+rm(snp_nodeid) #clean memory
+
+# Set genomic attributes
 snp_cirmagname <- "SPAdes-MetaBAT2Refined-S15.49"
@@
 
-# Export gene names
+# Export gene names & contigs
 if (file.exists("mags-all/5-SNPs/dndscv/results/circos/S15.49_genes.txt") == FALSE) {
@@
 }
-
-# Read parsed gene annotation (*externally processed*)
-#chr	start	end	gene_name
-#NODE_4971_length_14023_cov_250.866767	2403	4742	BNHKIBDC_01907
+if (file.exists("mags-all/5-SNPs/dndscv/results/circos/contigs/circos_S15.49.txt") == FALSE) {
+  write.table(snp_cirmag$contig, "mags-all/5-SNPs/dndscv/results/circos/contigs/circos_S15.49.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+
+# Read parsed gene annotation (externally* processed)
 snp_cirgenes <- read.csv("mags-all/5-SNPs/dndscv/results/circos/S15.49_genes_pos.txt", sep = "\t", stringsAsFactors = FALSE)
 
-# Merge with genome selection (chr start end dnds gene)
+# Merge with genome selection <chr start end dnds gene>
 snp_cirgenes <- snp_cirgenes %>% dplyr::mutate(chr = snp_cirmagname)
@@
 
+# Use global positions for genes (externally* processed)
+snp_cirgenpos <- read.csv("mags-all/5-SNPs/dndscv/results/circos/contigs/circos_S15.49_relpos.txt", sep = "\t", stringsAsFactors = FALSE)
+snp_cirgenes <- left_join(snp_cirgenes, snp_cirgenpos, by = c("contig"))
+snp_cirgenes <- snp_cirgenes %>% dplyr::mutate(start = start + rel_pos, end = end + rel_pos)
+rm(snp_cirgenpos) #clean memory
+
 # Circos plot
 circos.clear()
-pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_circos.pdf", height = 5, width = 5)
+pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_circos.pdf", height = 4, width = 4)
 circos.par(cell.padding = c(0, 0))
@@
   text(0, 0, snp_cirmagname, cex = 0.65)
-  text(0, -0.08, paste("phylum Actinobacteriota | genus Acidithrix", 
+  text(0, -0.14, paste("Genome size 2703949 bp", "phylum Actinobacteriota | genus Acidithrix",
                        "168 contigs | N50 26686 | GC 47.7", sep = "\n"), cex = 0.55)
-  circos.axis(h = "top", major.at = snp_cirran, labels = paste0(round(snp_cirran/10^6, 1), " Mb"),
+  circos.axis(h = "top", major.at = snp_cirran, labels = paste0(round(snp_cirran/10^6, 1)), #, " Mb"
               labels.cex = 0.35, lwd = 0.7, labels.facing = "clockwise")
-}, bg.col = "#e2edf2", bg.border = FALSE, track.height = 0.06)
-circos.genomicLabels(snp_cirgenes, labels.column = 6, cex = 0.35, line_lwd = 0.5,
+}, bg.col = "#6a9b85", bg.border = FALSE, track.height = 0.06)
+circos.genomicLabels(snp_cirgenes, labels.column = 7, cex = 0.35, line_lwd = 0.5,
                      side = "inside", line_col = "black", col = "black",
@@
 ### Enrichment ###
-# PS genes for enrichment (we need KEGG KOs from KOfamScan)
-snp_psenrich <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
-
-# Reduce to unique KOs, saving scores (converted to foldchange)
-snp_psenrich <- snp_psenrich %>%
-  dplyr::filter(V1 == "*") %>%
-  dplyr::group_by(V3, V7) %>%
-  dplyr::summarise(sum = n(), score = mean(log2(V4)), .groups = "drop") %>%
-  dplyr::rename(KO = V3, KO_definition = V7)
-#create a named vector
-#snp_kegglist <- snp_psenrich$score
-#names(snp_kegglist) <- snp_psenrich$KO
-
-# Get KEGG KO & pathway data
-snp_keggdata <- download_KEGG(species = "ko")
-snp_keggpat <- snp_keggdata$KEGGPATHID2NAME
-snp_keggkos <- snp_keggdata$KEGGPATHID2EXTID
-rm(snp_keggdata)
-
-# Enrich the genes by KO
-#https://www.biostars.org/p/494567/
-snp_keggenrich <- enricher(gene = snp_psenrich$KO, TERM2GENE = snp_keggkos,
-                           TERM2NAME = snp_keggpat, pAdjustMethod = "BH", pvalueCutoff = 0.05)
-
-# Dotplot
-pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichdot.pdf", height = 9, width = 7)
-dotplot(snp_keggenrich, showCategory = 30)
-dev.off()
-
-# Barplot
-pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichbar.pdf", height = 9, width = 7)
-barplot(snp_keggenrich, showCategory = 30)
-dev.off()
-
-# Category netplot
-pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichnet.pdf", height = 9, width = 13)
-cnetplot(snp_keggenrich, showCategory = 30, categorySize = "pvalue", foldChange = snp_psenrich$KO)
-dev.off()
-
-# Upset plot (overlaps among different genes)
-upsetplot(snp_keggenrich)
+## PS genes for enrichment (we need KEGG KOs from KOfamScan)
+#snp_psenrich <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = #FALSE, stringsAsFactors = FALSE)
+#
+## Reduce to unique KOs, saving scores (converted to foldchange)
+#snp_psenrich <- snp_psenrich %>%
+#  dplyr::filter(V1 == "*") %>%
+#  dplyr::group_by(V3, V7) %>%
+#  dplyr::summarise(sum = n(), score = mean(log2(V4)), .groups = "drop") %>%
+#  dplyr::rename(KO = V3, KO_definition = V7)
+##create a named vector
+##snp_kegglist <- snp_psenrich$score
+##names(snp_kegglist) <- snp_psenrich$KO
+#
+## Get KEGG KO & pathway data
+#snp_keggdata <- download_KEGG(species = "ko")
+#snp_keggpat <- snp_keggdata$KEGGPATHID2NAME
+#snp_keggkos <- snp_keggdata$KEGGPATHID2EXTID
+#rm(snp_keggdata) #clean
+#
+## Enrich the genes by KO
+##https://www.biostars.org/p/494567/
+#snp_keggenrich <- enricher(gene = snp_psenrich$KO, TERM2GENE = snp_keggkos,
+#                           TERM2NAME = snp_keggpat, pAdjustMethod = "BH", pvalueCutoff = 0.05)
+#
+## Dotplot
+#pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichdot.pdf", height = 9, width = 7)
+#dotplot(snp_keggenrich, showCategory = 30)
+#dev.off()
+#
+## Barplot
+#pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichbar.pdf", height = 9, width = 7)
+#barplot(snp_keggenrich, showCategory = 30)
+#dev.off()
+#
+## Category netplot
+#pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichnet.pdf", height = 9, width = 13)
+#cnetplot(snp_keggenrich, showCategory = 30, categorySize = "pvalue", foldChange = snp_psenrich$KO)
+#dev.off()
+#
+## Upset plot (overlaps among different genes)
+#upsetplot(snp_keggenrich)
 ```
@@
 # Merge plots using patchwork
-fig_3a <- wrap_plots(plot_spacer(), snp_plot_psgenes, snp_plot_estim_phy, plot_spacer(),
-                     nrow = 1, ncol = 4, widths = c(0.1, 0.85, 0.5, 0.4))
-fig_3b <- wrap_plots(snp_plot_estim_bin, snp_plot_nsgenes, snp_plot_nsgensub,
-                     nrow = 1, widths = c(1, 0.75, 1))
-fig_3 <- wrap_plots(fig_3a, fig_3b, nrow = 2) +
-  plot_annotation(tag_levels = list(c("a", "", "b", "c", "d", "e"))) &
+fig_3a <- wrap_plots(plot_spacer(), snp_plot_psgenes, snp_plot_estim_phy,
+                     snp_plot_estim_bin, plot_spacer(),
+                     nrow = 1, ncol = 5, widths = c(0.05, 0.75, 0.55, 0.65, 0.15))
+fig_3b <- wrap_plots(snp_plot_nsgenes, snp_plot_nsgensub, snp_plot_kos,
+                     nrow = 1, ncol = 3, widths = c(0.85, 0.85, 1))
+fig_3 <- wrap_plots(fig_3a, fig_3b, nrow = 2, heights = c(11.5, 10.5)) +
+  plot_annotation(tag_levels = list(c("a", "", "b", "c", "d", "e", "f"))) &
   theme(plot.tag =  element_text(face = "bold"))
-ggsave("mags-all/fig_3.pdf", fig_3, height = 20, width = 24)
-#ggsave("mags-all/fig_3.png", fig_3, height = 20, width = 24)
-rm(fig_3a, fig_3b)
+ggsave("mags-all/fig_3.pdf", fig_3, height = 22, width = 28)
+#ggsave("mags-all/fig_3.png", fig_3, height = 22, width = 28)
+rm(fig_3a, fig_3b) #clean memory
 
 #sup 3
-fig_s3 <- wrap_plots(snp_plot_psbins, snp_plot_kos, nrow = 1, widths = c(1, 0.5)) +
-  plot_annotation(tag_levels = list(c("a", "", "", "b"))) &
+fig_s3 <- wrap_plots(snp_plot_psbins, widths = c(1)) +
+  plot_annotation(tag_levels = list(c("a", "", ""))) &
   theme(plot.tag =  element_text(face = "bold"))
-ggsave("mags-all/fig_s3.pdf", fig_s3, height = 10, width = 26)
-ggsave("mags-all/fig_s3.png", fig_s3, height = 10, width = 26)
+ggsave("mags-all/fig_s3.pdf", fig_s3, height = 9, width = 13)
+#ggsave("mags-all/fig_s3.png", fig_s3, height = 9, width = 13)
 ```

### 2024-07-04 R/paper.Rmd
added latest changes

@@
 ---
-title: "Systemix Article: Combined MAG-SMAG Profiling"
+title: "Systemix MAG-SMAG Profiling"
 author: "Moises A. Rojas"
-date: "2023-12-01 // mod. Mar 2024"
+date: "2023-12-01 // mod. Mar 2024 // rev. Jun 2024"
 output: github_document
@@
 # Read Systemix MAGs taxonomy and QC files
-gtdb_taxonomy <- read.csv("1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
-mags_qc <- read.csv("2-statistics/checkm_parsed_bins.tsv", sep = "\t")
+gtdb_taxonomy <- read.csv("article_profiling/1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
+mags_qc <- read.csv("article_profiling/2-statistics/checkm_parsed_bins.tsv", sep = "\t")
 
 # Inspect completeness and contamination per sample, then select one
-filter(mags_qc, grepl("\\-S15\\.", bin)) %>%
-  dplyr::select(2:3) %>%
-  summary()
+filter(mags_qc, grepl("\\-S15\\.", bin)) %>% dplyr::select(2:3) %>% summary()
 mag_samples <- filter(gtdb_taxonomy, grepl("\\-S15\\.", genome))
@@
 # Define final attributes
-mag_samples['mag_num'] <- nrow(mag_samples)
-mag_samples['ecosystem'] <- "Biomining"
-mag_samples['continent'] <- "South America"
+mag_samples["mag_num"] <- nrow(mag_samples)
+mag_samples["ecosystem"] <- "Biomining"
+mag_samples["continent"] <- "South America"
 
@@
 # Save attributes
-smag_md1_sel <- smag_md1 %>%
-  dplyr::select("mag_name", "mag_num", "Ecosystem", "Continent")
+smag_md1_sel <- smag_md1 %>% dplyr::select("mag_name", "mag_num", "Ecosystem", "Continent")
 smag_md2_sel <- smag_md2 %>%
@@
 colnames(smag_data)[17] <- "continent"
-
 colnames(smag_data)[7] <- "kingdom"
@@
 
-## Section 1: In-depth taxonomy analysis
-
-### Biomining MAGs vs SMAG taxa comparison
+## Section 1: Taxonomic analysis
+
+### In-depth biomining MAGs vs SMAG comparison
 
@@
 
-# Keep only bacteria and defined ecosystems (discard archaea kingdom and non-NA eco.)
-mags_metadata_bac <- mags_metadata %>%
-  dplyr::filter(kingdom == "Bacteria" & !is.na(ecosystem)) #n=32975
+# Keep only bacteria and known ecosystems (n = 32975)
+mags_metadata_bac <- mags_metadata %>% dplyr::filter(kingdom == "Bacteria" & !is.na(ecosystem))
 
@@
 
-plot_depth_top_phyla <- ggplot(mags_md_top_phy, aes(x = freq, y = reorder(phy_top, freq))) +
+plot_depth_top_phyla <- mags_md_top_phy %>%
+  ggplot(aes(x = freq, y = reorder(phy_top, freq))) +
   geom_bar(stat = "identity", fill = "#459682") +
-  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.5, size = 2.5) +
+  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.25, size = 3) +
   labs(x = "Number", y = "Phylum") +
   scale_x_continuous(limits = c(0, 10000)) +
-  theme_classic()
-ggsave("plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)
+  theme_classic() +
+  theme(axis.text = element_text(size = 14),
+        axis.title = element_text(size = 16))
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)
 
@@
   labs(x = "", y = "Number of kSGBs from biomining") +
-  scale_fill_manual(values = "#78Acc2") +
+  scale_fill_manual(values = "#78acc2") +
+  coord_flip() +
   theme_classic() +
@@
         legend.position = c(0.88, 0.12),
-        strip.background = element_blank()) +#axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
-  facet_grid(.~ecosystem) +
-  coord_flip()
-ggsave("plot_md_biom_species.pdf", plot_depth_spec, height = 2.5, width = 8)
+        strip.background = element_blank(),
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14),
+        strip.text.x = element_text(size = 10)) +
+  facet_grid(.~ecosystem)
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_biom_species.pdf", plot_depth_spec, height = 3, width = 9)
 
@@
 
-plot_depth_spec_eco <- ggplot(mags_depth_spec_eco, aes(x = reorder(ecosystem, ratio),
-                                                       y = ratio, fill = ecosystem)) +
+plot_depth_spec_eco <- mags_depth_spec_eco %>%
+  ggplot(aes(x = reorder(ecosystem, ratio), y = ratio, fill = ecosystem)) +
   geom_bar(stat = "identity") +
-  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.5, size = 3) +
-  labs(x = "Ecosystem", y = "Number of kSGBs / MAGs") +
+  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.25, size = 3) +
+  labs(x = "Ecosystem", y = "Number of kSGBs/MAGs") +
   scale_y_continuous(limits = c(0, 10)) +
   scale_fill_manual(values = mags_colors_eco) +
+  coord_flip() +
   theme_classic() +
-  theme(legend.position = "none") +
-  coord_flip()
-ggsave("plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)
+  theme(legend.position = "none",
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)
 
@@
 
-plot_depth_sgbs <- ggplot(mags_depth_sgbs[1:60, ],
-                          aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
+plot_depth_sgbs <- mags_depth_sgbs[1:60, ] %>%
+  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
+  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
   scale_y_continuous(limits = c(0, 8000)) +
   labs(x = "Phylum", y = "Number of species") +
+  coord_flip() +
   theme_classic() +
-  theme(legend.position = c(0.85, 0.15), strip.placement = "outside") +
-  coord_flip()
-#ggsave("plot_md_SGBs_phylum.pdf", plot_depth_sgbs, height = 7, width = 6)
+  theme(legend.position = c(0.85, 0.15),
+        strip.placement = "outside",
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
 
@@
 
-plot_depth_sgbs_p <- ggplot(mags_depth_sgbs_p[1:60, ],
-                            aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
+plot_depth_sgbs_p <- mags_depth_sgbs_p[1:60, ] %>%
+  ggplot(aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
+  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
   scale_y_continuous(limits = c(0, 100)) +
   labs(x = "", y = "% of species") +
+  coord_flip() +
   theme_classic() +
@@
         axis.line.y = element_blank(),
-        axis.ticks.y = element_blank()) +
-  coord_flip()
-#ggsave("plot_md_SGBs_phylum_p.pdf", plot_depth_sgbs_p, height = 7, width = 5)
-
-plot_depth_sgbs_both <- (plot_depth_sgbs | plot_depth_sgbs_p) +
-  plot_layout(nrow = 1, ncol = 2, widths = c(0.55, 0.5))
-ggsave("plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)
+        axis.ticks.y = element_blank(),
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
+
+#merge
+plot_depth_sgbs_both <- wrap_plots(plot_depth_sgbs, plot_depth_sgbs_p,
+                                   nrow = 1, ncol = 2, widths = c(0.55, 0.5))
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)
 
@@
 
-plot_depth_sgbs_biom <- ggplot(mags_depth_sgbs_biom,
-                               aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
+plot_depth_sgbs_biom <- mags_depth_sgbs_biom %>%
+  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
+  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
   scale_y_continuous(limits = c(0, 25)) +
   labs(x = "Phylum", y = "Biomining species") +
+  coord_flip() +
   theme_classic() +
-  theme(legend.position = c(0.75, 0.2), strip.placement = "outside") +
-  coord_flip()
-ggsave("plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 5, width = 4)
+  theme(aspect.ratio = 0.875,
+        legend.position = c(0.75, 0.3),
+        strip.placement = "outside",
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 4, width = 6)
 
@@
 
-plot_depth_usgbs <- ggplot(mags_depth_usgbs_eco,
-                           aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
+plot_depth_usgbs <- mags_depth_usgbs_eco %>%
+  ggplot(aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
   geom_col() +
@@
         strip.background = element_blank(),
-        panel.border = element_blank()) +
+        panel.border = element_blank(),
+        axis.text = element_text(size = 14),
+        axis.title = element_text(size = 16),
+        strip.text.x = element_text(size = 10)) +
   facet_grid(.~ecosystem, scales = "free")
-ggsave("plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 8, width = 12)
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 9, width = 14)
 
@@
 
-plot_depth_kggbs_biom <- ggplot(mags_depth_kggbs_biom,
-                                aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
+plot_depth_kggbs_biom <- mags_depth_kggbs_biom %>%
+  ggplot(aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
   geom_col(position = position_stack(reverse = TRUE)) +
-  scale_x_continuous(limits = c(0, 101), n.breaks = 5) + #check!
+  scale_x_continuous(limits = c(0, 101), n.breaks = 5) +
   scale_fill_manual(values = mags_colors_eco) +
@@
   theme_classic() +
-  theme(legend.title = element_blank())
-ggsave("plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)
+  theme(legend.title = element_blank(),
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)
 
@@
 # Read QC file from CheckM
-mags_qc <- read.csv("mags-all/2-statistics/checkm_parsed_bins.tsv", sep = "\t")
+mags_qc <- read.csv("article_profiling/2-statistics/checkm_parsed_bins.tsv", sep = "\t")
 
@@
 # Plot
-plot_mags_qc <- ggplot(mags_qcs, aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
+plot_mags_qc <- mags_qcs %>%
+  ggplot(aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
   geom_col(position = "dodge2") +
-  scale_fill_manual(values = c("#DB7971", "#5CA650", "#859AC7")) +
+  #geom_text(aes(label = mags), vjust = 0.5, hjust = -0.25, size = 3) +
+  scale_fill_manual(values = c("#db7971", "#5ca650", "#859ac7")) +
   labs(x = "Number of MAGs", y = "") +
   theme_classic() +
-  theme(legend.position = "none")
-ggsave("mags-all/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 3, width = 5)
+  theme(aspect.ratio = 0.875,
+        axis.text = element_text(size = 12),
+        axis.title.x = element_text(size = 14),
+        legend.position = "none")
+ggsave("article_profiling/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 4, width = 6)
 ```
@@
 # Load GTDB taxonomy
-gtdb_taxonomy <- read.csv("1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
+gtdb_taxonomy <- read.csv("article_profiling/1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
 
@@
 
-# Read full tree (S15)
-tree <- ape::read.tree("1-taxonomy/trees-gtdbtk/gtdbtk.DASTool-S15.tree")
-
-# Subset to biomining placements
+# Read tree (S15 sample), subset it to biomining placements and save to file
+tree <- ape::read.tree("article_profiling/1-taxonomy/raw/gtdbtk-trees/gtdbtk.DASTool-S15.tree")
 tree_s15 <- keep.tip(tree, as.character(gtdb_taxonomy$genome))
 tree_s15 #44 tips
-
-# Export tree to file
-if (file.exists("1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree") == FALSE) {
-  ape::write.tree(tree_s15, "1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree")
+if (file.exists("article_profiling/1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree") == FALSE) {
+  ape::write.tree(tree_s15, "article_profiling/1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree")
 }
@@
 
-# Rename column {genome -> label} (to match with tree) from GTDB data
+# Rename column from GTDB data to match with the tree
 colnames(gtdb_taxonomy)[1] <- "label"
@@
 tree_s15_df <- tidytree::as_tibble(tree_s15)
-tree_s15_df <- treeio::full_join(tree_s15_df, gtdb_taxonomy[c("label", "phylum", "family", "genus", "species")], by = "label")
+tree_s15_df <- treeio::full_join(tree_s15_df, gtdb_taxonomy[c("label", "phylum", "class")], by = "label")
+
+# Summarise phyla
+tree_s15_sum_df <- tree_s15_df %>%
+  filter(!is.na(phylum)) %>%
+  dplyr::group_by(phylum) %>%
+  summarise(n = n()) %>%
+  dplyr::mutate(phylum_n = paste0(phylum, " (", n, ")"))
+tree_s15_df <- left_join(tree_s15_df, tree_s15_sum_df, by = c("phylum"))
+rm(tree_s15_sum_df) #clean
 
@@
 # Plot circular tree
-plot_tree_s15 <- ggtree(tree_s15_md, layout = "circular", size = 0.2)
-
+#plot_tree_s15 <- ggtree(tree_s15_md, layout = "circular", size = 0.2)
 # Plot tree adding phylum
-plot_tree_s15_phyla <- plot_tree_s15 +
-  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
-  geom_tiplab(aes(label = phylum, color = phylum), size = 1.75, hjust = -0.1) +
-  scale_colour_discrete(breaks = c("c", "d"), na.value = NA) +
-  scale_colour_manual(values = tree_colors, na.translate = FALSE, guide = "none") +
+#plot_tree_s15_phyla <- plot_tree_s15 +
+#  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
+#  geom_tiplab(aes(label = phylum, color = phylum), size = 1.75, hjust = -0.1) +
+#  scale_colour_discrete(breaks = c("c", "d"), na.value = NA) +
+#  scale_colour_manual(values = tree_colors, na.translate = FALSE, guide = "none") +
   #guides(color = guide_legend(title = "Phyla", ncol = 1, order = 1)) +
-  theme()
-ggsave("plot_biom_MAGs_tree_phyla.pdf", plot_tree_s15_phyla, height = 5, width = 7)
+#  theme()
+#ggsave("plot_biom_MAGs_tree_phyla.pdf", plot_tree_s15_phyla, height = 5, width = 7)
 
@@
 tree_pg  <- vector()
-
-# Match datasets
 for (i in 1:nrow(tree_s15_df)) {
@@
                      Genome_Size = tree_gs, GC = tree_gc, Pred_Genes = tree_pg)
-rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg) #clean memory
+rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg, i, j) #clean memory
 
 # Plot tree adding QC as multiple layers
-plot_tree_s15_md <- plot_tree_s15_phyla +
+plot_tree_s15_md <- ggtree(tree_s15_md, layout = "circular", size = 0.1) +
+  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
+  geom_tiplab(aes(label = phylum, color = phylum), size = 2, hjust = -0.2) +
+  geom_treescale(x = -0.15, y = 0, offset = 0.2, fontsize = 1.4, linesize = 0.1, width = 0.1) +
+  scale_colour_manual(values = tree_colors, guide = "none") +
+  #add QC as layers
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Completeness, fill = Completeness),
-             orientation = "y", stat = "identity", position = "auto", width = 0.85,
-             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.6,
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3,
-                                title = "Complet.", title.size = 1.6,
-                                title.height = 0.035, title.angle = 30)) +
-  scale_fill_gradient(low = "#1F4060", high = "#4CA8F6", guide = "none") +
-                      #guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
-                      #                     reverse = TRUE, order = 1)) +
-  ###Error in scale$guide == "none"
+             orientation = "y", stat = "identity", position = "auto",
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 1,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3)) +
+  scale_fill_gradient(name = "Completeness (%)", low = "#1f4060", high = "#4497dd",
+                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 1)) +
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Contamination, fill = Contamination),
-             orientation = "y", stat = "identity", position = "auto", width = 0.85,
-             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1,
-                                title = "Contam.", title.size = 1.6,
-                                title.height = 0.035, title.angle = 30)) +
-  scale_fill_gradient(low = "#6A60A8", high = "#C4C2DE", guide = "none") +
+             orientation = "y", stat = "identity", position = "auto",
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 0.15,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3)) +
+  scale_fill_gradient(name = "Contamination (%)", low = "#6a60a8", high = "#c4c2de",
+                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 2)) +
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Strain_Het, fill = Strain_Het),
-             orientation = "y", stat = "identity", position = "auto", width = 0.85,
-             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3,
-                                title = "Strain Het.", title.size = 1.6,
-                                title.height = 0.035, title.angle = 30)) +
-  scale_fill_gradient(low = "#C9543F", high = "#E9BAB2",
-                      name = "Strain Heterogeneity", guide = "none") +
+             orientation = "y", stat = "identity", position = "auto",
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 0.15,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3)) +
+  scale_fill_gradient(name = "Strain Heterogeneity", low = "#c9543f", high = "#e9bab2",
+                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 3)) +
   new_scale_fill() + new_scale_color() +
@@
              mapping = aes(y = Bin, x = Pred_Genes, fill = Pred_Genes),
-             orientation = "y", stat = "identity", position = "auto", width = 0.85,
-             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
-             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1,
-                                title = "Pred. genes", title.size = 1.6,
-                                title.height = 0.035, title.angle = 30)) +
-  scale_fill_gradient(low = "#279854", high = "#A9DEDE",
-                      name = "Predicted Genes", guide = "none") +
-  theme()
-  #theme(legend.title = element_text(size = 8),
-  #      legend.text = element_text(size = 7),
-  #      legend.key.height = unit(3.5, units = "mm"),
-  #      legend.key.width = unit(2, units = "mm"))
-ggsave("plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 7) #x9
+             orientation = "y", stat = "identity", position = "auto",
+             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 0.15,
+             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 2)) +
+  scale_fill_gradient(name = "Predicted Genes", low = "#279854", high = "#a9dede",
+                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 4)) +
+  theme(legend.title = element_text(size = 10),
+        legend.text = element_text(size = 9),
+        legend.key.height = unit(4, units = "mm"),
+        legend.key.width = unit(3, units = "mm"))
+ggsave("article_profiling/1-taxonomy/plots/plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 9)
 ```
@@
 # Merge plots using patchwork
-fig_1a <- { plot_tree_s15_md | plot_depth_spec_eco +
-  plot_layout(ncol = 2, widths = c(1, 0.5))
-  } + plot_annotation(tag_levels = list(c("a", "e"))) &
+fig_1a <- wrap_plots(plot_tree_s15_md, plot_mags_qc, plot_depth_sgbs_biom,
+                     design = "\nAAAB\nAAAC", heights = c(1, 0.85))
+#ggsave("article_profiling/fig_1a.pdf", fig_1a, height = 8, width = 15)
+fig_1b <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco, plot_depth_kggbs_biom,
+                     design = "\nAAABBCC\n")
+#ggsave("article_profiling/fig_1b.pdf", fig_1b, height = 8, width = 15)
+
+fig_1 <- wrap_plots(fig_1a, fig_1b, nrow = 2, heights = c(1, 0.85)) +
+  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "", "e", "f"))) &
   theme(plot.tag = element_text(face = "bold"))
-ggsave("fig_1a.pdf", fig_1a, height = 7, width = 14)
-
-fig_1b <- { plot_mags_qc / plot_depth_sgbs_biom |
-    plot_depth_sgbs_both | plot_depth_kggbs_biom
-  } + plot_layout(ncol = 3, widths = c(0.5, 2, 0.75)) +
-  plot_annotation(tag_levels = list(c("b", "c", "d", "", "f"))) &
-  theme(plot.tag =  element_text(face = "bold"))
-ggsave("fig_1b.pdf", fig_1b, height = 7, width = 14)
-
-fig_1 <- (fig_1a / fig_1b) +
-  plot_annotation(tag_levels = list(c("a", "e", "b", "c", "d", "", "f"))) &
-  theme(plot.tag =  element_text(face = "bold"))
-ggsave("fig_1.pdf", fig_1, height = 14, width = 14)
-#ggsave("fig_1.png", fig_1, height = 14, width = 14)
-rm(fig_1a, fig_1b) #clean memory
+ggsave("article_profiling/fig_1.pdf", fig_1, height = 16, width = 16)
+#ggsave("article_profiling/fig_1.png", fig_1, height = 16, width = 16)
+rm(fig_1a, fig_1b) #clean
 
@@
   plot_annotation(tag_levels = list(c("a", "b", "c"))) &
-  theme(plot.tag =  element_text(face = "bold"))
+  theme(plot.tag = element_text(size = 20, face = "bold"))
 fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.title.y = element_text(vjust = 0))
-#fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.text.y = element_text(vjust = 0))
-#fig_s1 <- wrap_plots(plot_depth_usgbs, plot_depth_spec, nrow = 2, ncol = 1, heights = c(0.85, 0.15))
-ggsave("mags-all/fig_s1.pdf", fig_s1, height = 9, width = 18)
-#ggsave("mags-all/fig_s1.png", fig_s1, height = 9, width = 18)
+fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.line.y.left = element_blank(),
+                                             axis.ticks.y.left = element_blank())#,
+                                             #axis.text.y = element_text(hjust = 2.2))
+ggsave("article_profiling/fig_s1.pdf", fig_s1, height = 13, width = 22)
+#ggsave("article_profiling/fig_s1.png", fig_s1, height = 13, width = 22)
 ```
@@
 }
-kegg_paths['Pathway'] <- kegg_ko_paths
-kegg_paths['Metabolism'] <- kegg_ko_metab
+kegg_paths["Pathway"] <- kegg_ko_paths
+kegg_paths["Metabolism"] <- kegg_ko_metab
 rm(kegg_ko_paths, kegg_ko_metab) #clean memory
@@
 }
-kegg_mags['ecosystem'] <- kegg_mags_eco
-kegg_mags['length'] <- kegg_mags_length
+kegg_mags["ecosystem"] <- kegg_mags_eco
+kegg_mags["length"] <- kegg_mags_length
 rm(kegg_mags_eco, kegg_mags_length) #clean memory
@@
 }
-kegg_pca_kos['Pathway'] <- kegg_pca_def
+kegg_pca_kos["Pathway"] <- kegg_pca_def
 
@@
 
-### Bubble and box plots
+### Bubble and boxplots
 
@@
 colnames(kegg_bubble) <- c("KO")
-kegg_bubble['ecosystem'] <- c("")
+kegg_bubble["ecosystem"] <- c("")
 kegg_bubble$ecosystem <- rep(kegg_bubble_eco, nrow(kegg_filt))
@@
                            order = anova_sort_cu$ecosystem, add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(kegg_bubble_cu$count_norm), linetype = 2, linewidth = 0.2) +
-  labs(x = "Ecosystem", y = "KOs/Mb for copper") +
-  rotate_x_text(angle = 45) + #ylim(c(0, 0.25)) +
+  geom_hline(yintercept = mean(kegg_bubble_cu$count_norm), linetype = 2, linewidth = 0.5) +
+  labs(x = "Ecosystem", y = "Copper pathway (KOs/Mb)") +
+  rotate_x_text(angle = 45) + ylim(c(0, 0.3)) +
   #stat_summary(fun = mean, geom = "point", shape = 18, size = 2.5, color = "#FC4E07") +
-  stat_compare_means(method = "anova", label.y = 0.25) + #add global annova p-value
-  stat_compare_means(label = "p.signif", method = "t.test",
-                     ref.group = ".all.", hide.ns = TRUE)
-ggsave("plot_annot_boxplot_copper.pdf", anova_boxplot_cu, height = 7, width = 7)
+  stat_compare_means(method = "anova", label.y = 0.3, size = 5) + #add global annova p-value
+  stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE) +
+  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))
+ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_boxplot_copper.pdf", anova_boxplot_cu, height = 7, width = 7)
 
@@
                            order = anova_sort_fe$ecosystem, add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(kegg_bubble_fe$count_norm), linetype = 2, linewidth = 0.2) +
-  labs(x = "Ecosystem", y = "KOs/Mb for iron") +
-  rotate_x_text(angle = 45) +
-  stat_compare_means(method = "anova", label.y = 0.4) + #add global annova p-value
+  geom_hline(yintercept = mean(kegg_bubble_fe$count_norm), linetype = 2, linewidth = 0.5) +
+  labs(x = "Ecosystem", y = "Iron pathway (KOs/Mb)") +
+  rotate_x_text(angle = 45) + ylim(c(0, 0.4)) +
+  stat_compare_means(method = "anova", label.y = 0.4, size = 5) + #add global annova p-value
   stat_compare_means(label = "p.signif", method = "t.test",
-                     ref.group = ".all.", hide.ns = TRUE, label.y = 0.35)
-ggsave("plot_annot_boxplot_iron.pdf", anova_boxplot_fe, height = 7, width = 7)
+                     ref.group = ".all.", hide.ns = TRUE, label.y = 0.35) +
+  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))
+ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_boxplot_iron.pdf", anova_boxplot_fe, height = 7, width = 7)
 
@@
                              order = anova_sort_s$ecosystem, add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(kegg_bubble_s$count_norm), linetype = 2, linewidth = 0.2) +
-  labs(x = "Ecosystem", y = "KOs/Mb for sulfur") +
-  rotate_x_text(angle = 45) +
-  stat_compare_means(method = "anova", label.y = 0.7) + #add global annova p-value
-  stat_compare_means(label = "p.signif", method = "t.test",
-                     ref.group = ".all.", hide.ns = TRUE)
-ggsave("plot_annot_boxplot_sulfur.pdf", anova_boxplot_s, height = 7, width = 7)
+  geom_hline(yintercept = mean(kegg_bubble_s$count_norm), linetype = 2, linewidth = 0.5) +
+  labs(x = "Ecosystem", y = "Sulfur pathway (KOs/Mb)") +
+  rotate_x_text(angle = 45) + ylim(c(0, 0.8)) +
+  stat_compare_means(method = "anova", label.y = 0.8, size = 5) + #add global annova p-value
+  stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.",
+                     hide.ns = TRUE, label.y = 0.75, size = 6) +
+  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))
+ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_boxplot_sulfur.pdf", anova_boxplot_s, height = 7, width = 7)
 
@@
              y = Definition, size = count_norm, color = ecosystem)) +
-  geom_point(alpha = 0.7) +
+  geom_point(alpha = 0.5) +
+  scale_size(range = c(0, 10)) +
+  scale_y_discrete(position = "right") +
   scale_color_manual(values = mags_colors_eco, guide = "none") +
-  labs(title = "", x = "Ecosystem", y = "Highlighted KO definitions") +
+  labs(title = "", x = "Ecosystem", y = "") +
   guides(size = guide_legend(title = "Norm. counts")) +
   theme_classic() +
-  theme(legend.title = element_text(size = 9), legend.position = "right",
-        plot.subtitle = element_text(size = 8, hjust = 0.5, face = "italic"),
-        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
-ggsave("plot_annot_bubbles.pdf", kegg_bubbles_plot, height = 7, width = 8.5)
+  theme(legend.title = element_text(size = 12), legend.position = "top",
+        axis.text.x = element_text(size = 14, angle = 90, vjust = 1, hjust = 1),
+        axis.text.y = element_text(size = 14), axis.title.x = element_text(size = 16),
+        axis.line.y.right = element_blank(), axis.ticks.y.right = element_blank())
+ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_bubbles.pdf", kegg_bubbles_plot, height = 7, width = 12)
 
@@
 # Merge plots using patchwork
-fig_2 <- { { anova_boxplot_cu / anova_boxplot_fe +
-    plot_layout(ncol = 1, widths = c(0.75)) } | anova_boxplot_s / kegg_bubbles_plot } +
-  plot_annotation(tag_levels = list(c("a", "b", "c", "d"))) &
-  theme(plot.tag =  element_text(face = "bold"))
-#adjust y-label margin for sulfur plot
-fig_2[[2]][[1]] <- fig_2[[2]][[1]] + theme(axis.title.y = element_text(vjust = -127))
-ggsave("mags-all/fig_2.pdf", fig_2, height = 14, width = 18)
-#ggsave("mags-all/fig_2.png", fig_2, height = 14, width = 18) #vjust=120
+fig_2 <- wrap_plots(anova_boxplot_cu, anova_boxplot_fe, anova_boxplot_s, design = "AABBCC") +
+  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+  theme(plot.tag = element_text(size = 22, face = "bold"))
+ggsave("article_profiling/fig_2.pdf", fig_2, height = 7, width = 19)
+ggsave("article_profiling/fig_2.png", fig_2, height = 7, width = 19)
+
+#sup 2
+fig_s2 <- wrap_plots(kegg_bubbles_plot, design = "AAA") +
+  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+ggsave("article_profiling/fig_s2a.pdf", fig_s2, height = 7, width = 12)
+#ggsave("article_profiling/fig_s2a.png", fig_s2, height = 7, width = 12)
 ```
 
-## Section 3: SNP Variant Calling
-
-### dN/dS analysis
+## Section 3: Evolutionary analysis
+
+### dN/dS
 
@@
 #https://github.com/im3sanger/dndscv/issues/62
-#buildref(cdsfile = "mags-all/5-SNPs/dndscv/buildref/cds_node_25854.txt",
-#         genomefile = "mags-all/5-SNPs/dndscv/buildref/node_25854.fa",
-#         outfile = "mags-all/5-SNPs/dndscv/buildref/testref.rda")
-#load("mags-all/5-SNPs/dndscv/buildref/testref.rda")
+#buildref(cdsfile = "article_profiling/5-SNPs/dndscv/buildref/cds_node_25854.txt",
+#         genomefile = "article_profiling/5-SNPs/dndscv/buildref/node_25854.fa",
+#         outfile = "article_profiling/5-SNPs/dndscv/buildref/testref.rda")
+#load("article_profiling/5-SNPs/dndscv/buildref/testref.rda")
 
 # Buildref for all bins (very slow! ~2 hrs in building object)
-#buildref(cdsfile = "mags-all/5-SNPs/dndscv/buildref/cds_table.txt",
-#         genomefile = "mags-all/5-SNPs/dndscv/buildref/S15_bins.fa",
-#         outfile = "mags-all/5-SNPs/dndscv/buildref/S15_refcds.rda")
+#buildref(cdsfile = "article_profiling/5-SNPs/dndscv/buildref/cds_table.txt",
+#         genomefile = "article_profiling/5-SNPs/dndscv/buildref/S15_bins.fa",
+#         outfile = "article_profiling/5-SNPs/dndscv/buildref/S15_refcds.rda")
 
 # Inspect refdb
-#load("mags-all/5-SNPs/dndscv/buildref/S15_refcds.rda")
+#load("article_profiling/5-SNPs/dndscv/buildref/S15_refcds.rda")
 #print(RefCDS[[1]])
@@
 #sampleID chr pos ref mut
-snp_mutations <- read.csv("mags-all/5-SNPs/dndscv/run/S15_mutations.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_mutations <- read.csv("article_profiling/5-SNPs/dndscv/run/S15_mutations.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
 # Run dndscv (runtime ~8 min, high memory usage) and save object
-#snp_dndsout <- dndscv(snp_mutations, refdb = "mags-all/5-SNPs/dndscv/buildref/S15_refcds.rda",
+#snp_dndsout <- dndscv(snp_mutations, refdb = "article_profiling/5-SNPs/dndscv/buildref/S15_refcds.rda",
 #                      max_muts_per_gene_per_sample = Inf, max_coding_muts_per_sample = Inf,
 #                      cv = NULL, outmats = TRUE)
-#save(snp_dndsout, file = "mags-all/5-SNPs/dndscv/run/S15_dndsout.rda")
-#save(substmodel, file = "mags-all/5-SNPs/dndscv/run/S15_substmodel.rda")
-#load("mags-all/5-SNPs/dndscv/run/S15_dndsout.rda")
-#load("mags-all/5-SNPs/dndscv/run/S15_substmodel.rda")
+#save(snp_dndsout, file = "article_profiling/5-SNPs/dndscv/run/S15_dndsout.rda")
+#save(substmodel, file = "article_profiling/5-SNPs/dndscv/run/S15_substmodel.rda")
+#load("article_profiling/5-SNPs/dndscv/run/S15_dndsout.rda")
+#load("article_profiling/5-SNPs/dndscv/run/S15_substmodel.rda")
 rm(snp_mutations) #clean memory
@@
   theme(legend.position = "none")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndsglob, height = 2, width = 5)
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_global.pdf", snp_plot_dndsglob, height = 2, width = 5)
 
@@
 # Get CDS annotations (pre-processed from Bakta results)
-snp_refcds <- read.csv("mags-all/5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_refcds <- read.csv("article_profiling/5-SNPs/dndscv/buildref/cds_table.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
 # Load contig to bin matrix
-snp_bins <- read.csv("mags-all/5-SNPs/dndscv/results/S15_DASTool_contig2bin.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_bins <- read.csv("article_profiling/5-SNPs/dndscv/results/S15_DASTool_contig2bin.tsv", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
 # Load gene names (parsed from genes.gff file)
-snp_genids <- read.csv("mags-all/5-SNPs/dndscv/results/gene_id_name.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_genids <- read.csv("article_profiling/5-SNPs/dndscv/results/gene_id_name.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
@@
 }
-snp_psgenes['contig'] <- snp_gene2contig
-snp_psgenes['bin'] <- snp_contig2bin
-snp_psgenes['name'] <- snp_genenames
+snp_psgenes["contig"] <- snp_gene2contig
+snp_psgenes["bin"] <- snp_contig2bin
+snp_psgenes["name"] <- snp_genenames
 rm(snp_refcds, snp_bins, snp_genids, snp_gene2contig, snp_contig2bin, snp_genenames) #clean memory
@@
 }
-snp_dnds['phylum'] <- snp_taxabins$phylum
-snp_dnds['genus'] <- snp_taxabins$genus
-snp_dnds['species'] <- snp_taxabins$species
-snp_dnds['completeness'] <- snp_taxabins$completeness
-snp_dnds['contamination'] <- snp_taxabins$contamination
-snp_dnds['predicted_genes'] <- snp_taxabins$predicted_genes
+snp_dnds["phylum"] <- snp_taxabins$phylum
+snp_dnds["genus"] <- snp_taxabins$genus
+snp_dnds["species"] <- snp_taxabins$species
+snp_dnds["completeness"] <- snp_taxabins$completeness
+snp_dnds["contamination"] <- snp_taxabins$contamination
+snp_dnds["predicted_genes"] <- snp_taxabins$predicted_genes
 rm(snp_taxabins) #clean memory
@@
 # Save PS gene names and ids to file
-if (file.exists("mags-all/5-SNPs/dndscv/results/ps_genes.txt") == FALSE) {
-  write.table(snp_psgenes_sub %>% dplyr::select(gene_name) %>% distinct(gene_name), "mags-all/5-SNPs/dndscv/results/ps_genes.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
-}
-if (file.exists("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt") == FALSE) {
-  write.table(unique(snp_psgenes_sub$name), "mags-all/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+if (file.exists("article_profiling/5-SNPs/dndscv/results/ps_genes.txt") == FALSE) {
+  write.table(snp_psgenes_sub %>% dplyr::select(gene_name) %>% distinct(gene_name), "article_profiling/5-SNPs/dndscv/results/ps_genes.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+if (file.exists("article_profiling/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt") == FALSE) {
+  write.table(unique(snp_psgenes_sub$name), "article_profiling/5-SNPs/dndscv/results/kofam/ps_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
 }
@@
 # Plot global driver genes by MAG/species
-snp_plot_psbins <- ggplot(snp_dnds, aes(x = num, y = reorder(bin, num), fill = bin)) +
-  geom_bar(stat = "identity", fill = "#E31A1C") +
-  geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
+snp_plot_psbins <- snp_dnds %>%
+  dplyr::mutate(bin = str_replace_all(bin, "SPAdes-MaxBin2Refined-", "")) %>% #trim bin names
+  dplyr::mutate(bin = str_replace_all(bin, "SPAdes-MetaBAT2Refined-", "")) %>%
+  ggplot(aes(x = num, y = reorder(bin, num), fill = bin)) +
+  geom_bar(stat = "identity", fill = "#e31a1c") +
+  geom_text(aes(label = num), vjust = 0.5, hjust = -0.25, size = 4) +
   labs(x = "Number of genes under PS", y = "MAG Bin") +
@@
   theme_classic() +
-  theme(strip.background = element_blank()) +
+  theme(strip.background = element_blank(), axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18), strip.text.x = element_text(size = 14)) +
   facet_grid(.~species)
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 9, width = 10)
-
-# Add completeness & contamination barplots, then merge
-snp_plot_psbins2 <- ggplot(snp_dnds, aes(x = completeness, y = reorder(bin, num), fill = bin)) +
+#ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 11, width = 12)
+#add completeness & contamination barplots
+snp_plot_psbins2 <- snp_dnds %>%
+  ggplot(aes(x = completeness, y = reorder(bin, num), fill = bin)) +
   geom_col(fill = "#3c4d76") +
-  geom_vline(xintercept = 100, linetype = 1, linewidth = 0.2) +
+  geom_vline(xintercept = 100, linetype = 1, linewidth = 0.4) +
   labs(x = "Completeness", y = "") +
@@
         axis.line.y = element_blank(),
-        axis.ticks.y = element_blank())
-snp_plot_psbins3 <- ggplot(snp_dnds, aes(x = contamination, y = reorder(bin, num), fill = bin)) +
+        axis.ticks.y = element_blank(),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+snp_plot_psbins3 <- snp_dnds %>%
+  ggplot(aes(x = contamination, y = reorder(bin, num), fill = bin)) +
   geom_col(fill = "#763c4d") +
-  geom_vline(xintercept = max(snp_dnds$contamination), linetype = 1, linewidth = 0.2) +
+  geom_vline(xintercept = max(snp_dnds$contamination), linetype = 1, linewidth = 0.4) +
   labs(x = "Contamination", y = "") +
@@
         axis.line.y = element_blank(),
-        axis.ticks.y = element_blank())
-
+        axis.ticks.y = element_blank(),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+#merge
 snp_plot_psbins <- wrap_plots(snp_plot_psbins, snp_plot_psbins2, snp_plot_psbins3,
                               nrow = 1, ncol = 3, widths = c(1, 0.15, 0.15))
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 9, width = 12)
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psbins.pdf", snp_plot_psbins, height = 11, width = 17)
 rm(snp_plot_psbins2, snp_plot_psbins3) #clean memory
@@
 # Lollipop plot of PS genes
-snp_plot_psgenes <- ggplot(snp_psgenes_sub,
-                           aes(x = wmis_cv, y = reorder(name, wmis_cv)), size = wmis_cv) +
+snp_plot_psgenes <- snp_psgenes_sub %>%
+  dplyr::filter(wmis_cv > 1.4) %>% #cutoff
+  ggplot(aes(x = wmis_cv, y = reorder(name, wmis_cv)), size = wmis_cv) +
   geom_segment(aes(x = 0, xend = wmis_cv, y = name), color = "gray") +
-  geom_point(alpha = 0.7, color = "#E31A1C") +
-  geom_text(aes(label = signif, x = -0.15), size = 3, color = "#3c4d76") +
-  geom_text(aes(label = path, x = -0.15), size = 2, color = "#3c4d76") +
-  geom_vline(xintercept = 1, linetype = 1, linewidth = 0.2) +
-  geom_vline(xintercept = max(snp_psgenes_sub$wmis_cv), linetype = 1, linewidth = 0.2) +
-  labs(x = "PS strength", y = "Gene CDS") +
-  theme_classic()
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 11, width = 7)
+  geom_point(alpha = 0.7, size = 2, color = "#e31a1c") +
+  geom_text(aes(label = signif, x = -0.15), size = 6.5, color = "#3c4d76") +
+  geom_text(aes(label = path, x = -0.15), size = 5.5, color = "#3c4d76") +
+  geom_vline(xintercept = 1, linetype = 2, linewidth = 0.4) +
+  geom_vline(xintercept = max(snp_psgenes_sub$wmis_cv), linetype = 1, linewidth = 0.4) +
+  labs(x = "PS strength", y = "Gene name") +
+  theme_classic() +
+  theme(axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+#ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 12, width = 7)
 
 # Plot PS genes mutations per type, then merge
-snp_plot_muts <- ggplot(snp_psgenes_sub, aes(x = n_mut, y = reorder(name, wmis_cv), fill = mut)) +
+snp_plot_muts <- snp_psgenes_sub %>%
+  dplyr::filter(wmis_cv > 1.4) %>% #cutoff
+  ggplot(aes(x = n_mut, y = reorder(name, wmis_cv), fill = mut)) +
   geom_col(position = position_stack(reverse = TRUE)) +
@@
   theme_classic() +
-  theme(legend.position = c(0.85, 0.15),
-        strip.placement = "outside",
-        axis.text.y = element_blank(),
-        axis.line.y = element_blank(),
-        axis.ticks.y = element_blank())
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psmuts.pdf", snp_plot_muts, height = 11, width = 7)
-
-snp_plot_psgenes <- (snp_plot_psgenes | snp_plot_muts) +
-  plot_layout(nrow = 1, ncol = 2, widths = c(0.55, 0.5))
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_psgenmuts.pdf", snp_plot_psgenes, height = 11, width = 11)
+  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),
+        legend.position = c(0.88, 0.15), strip.placement = "outside",
+        axis.ticks.y = element_blank(), legend.text = element_text(size = 16),
+        axis.text = element_text(size = 16), axis.title = element_text(size = 18))
+#ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psmuts.pdf", snp_plot_muts, height = 12, width = 7)
+#merge
+snp_plot_psgenes <- wrap_plots(snp_plot_psgenes, snp_plot_muts, ncol = 2, widths = c(0.55, 0.5))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 12, width = 11)
 rm(snp_plot_muts) #clean memory
@@
 # Load negative selected (NS) genes from selected pathways
-snp_nsgenes <- read.csv("mags-all/5-SNPs/dndscv/results/selected_cds_metals.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
+snp_nsgenes <- read.csv("article_profiling/5-SNPs/dndscv/results/selected_cds_metals.txt", sep = "\t", header = TRUE, stringsAsFactors = FALSE)
 
@@
 # Save NS gene ids to file
-if (file.exists("mags-all/5-SNPs/dndscv/results/kofam/ns_genes_ids.txt") == FALSE) {
-  write.table(snp_nsgenes$gene_id, "mags-all/5-SNPs/dndscv/results/kofam/ns_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+if (file.exists("article_profiling/5-SNPs/dndscv/results/kofam/ns_genes_ids.txt") == FALSE) {
+  write.table(snp_nsgenes$gene_id, "article_profiling/5-SNPs/dndscv/results/kofam/ns_genes_ids.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
 }
@@
                              shape = "pathway", xlab = "Pathway", ylab = "dN/dS",
-                             order = snp_sort_ns$pathway, add = "jitter", legend = "none",
-                             palette = snp_sort_ns$color) +
-  scale_y_continuous(limits = c(-0.2, 1.3)) +
-  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
-  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
-  stat_compare_means(method = "anova", label.y = 1.3) +
+                             order = snp_sort_ns$pathway, palette = snp_sort_ns$color,
+                             size = 0.75, add = "jitter", legend = "none") +
+  scale_y_continuous(limits = c(-0.2, 1.5)) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.4) +
+  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.4) +
+  stat_compare_means(method = "anova", label.y = 1.5, size = 6) +
   stat_compare_means(comparisons = list(c("Copper", "Iron"), c("Sulfur", "Nickel"),
-                                        c("Copper", "Nickel")), label = "p.format", method = "t.test")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 9, width = 7)
+                                        c("Copper", "Nickel")), label = "p.format",
+                     method = "t.test", size = 5) +
+  theme(axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 10, width = 7)
 
@@
 snp_plot_nsgensub <- ggdotchart(snp_nsgenes_sub, x = "product", y = "dnds", color = "pathway",
-                                xlab = "NS gene product", ylab = "Mean dN/dS", y.text.col = TRUE,
-                                rotate = TRUE, palette = snp_nsgenes_sub_col, dot.size = 3,
-                                sorting = "descending", ggtheme = theme_pubr()) +
+                                xlab = "Product of genes under NS", ylab = "Mean dN/dS",
+                                y.text.col = TRUE, rotate = TRUE, palette = snp_nsgenes_sub_col,
+                                dot.size = 3, sorting = "descending", ggtheme = theme_pubr()) +
   scale_y_continuous(limits = c(0, 1)) +
-  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.2) +
-  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2)
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_nsgenesprod.pdf", snp_plot_nsgensub, height = 9, width = 10)
+  geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.4) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.4) +
+  font("legend.title", size = 14) +
+  theme(legend.text = element_text(size = 14),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenesprod.pdf", snp_plot_nsgensub, height = 11, width = 13)
 
@@
 }
-snp_estim['contig'] <- snp_gene2contig
-snp_estim['bin'] <- snp_contig2bin
-snp_estim['name'] <- snp_genenames
+snp_estim["contig"] <- snp_gene2contig
+snp_estim["bin"] <- snp_contig2bin
+snp_estim["name"] <- snp_genenames
 rm(snp_refcds, snp_bins, snp_genids, snp_gene2contig, snp_contig2bin, snp_genenames) #clean memory
@@
 
-# Sort phylum by median and set color
+# Sort phylum by median and set colors
 snp_estim_sort_phy <- snp_estim_phy %>%
   dplyr::group_by(phylum) %>%
-  dplyr::summarise(n = median(dnds_glob)) %>% #by mean?
+  dplyr::summarise(n = median(dnds_glob)) %>%
   dplyr::arrange(desc(n))
-snp_estim_sort_phy$color <- rep("#4c6194", nrow(snp_estim_sort_phy))
+snp_estim_sort_phy$color <- c("#b15928", "#98d277", "#a6cee3")
 
@@
                                 order = snp_estim_sort_phy$phylum, palette = snp_estim_sort_phy$color,
-                                add = "jitter", legend = "none") +
-  geom_hline(yintercept = mean(snp_estim_phy$dnds_glob), linetype = 2, linewidth = 0.2) +
-  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
+                                add = "jitter", size = 0.75, legend = "none") +
+  geom_hline(yintercept = mean(snp_estim_phy$dnds_glob), linetype = 2, linewidth = 0.4) +
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.4) +
   labs(x = "Phylum", y = "Global normalized dN/dS ratio") +
-  stat_compare_means(method = "anova", label.y = 1.6) +
+  stat_compare_means(method = "anova", label.y = 1.6, size = 6) +
   stat_compare_means(comparisons = list(c("Proteobacteria", "Actinobacteriota"),
@@
                                         c("Actinobacteriota", "Acidobacteriota")),
-                     label = "p.format", method = "t.test")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_estim_phy.pdf", snp_plot_estim_phy, height = 9, width = 7)
+                     label = "p.format", method = "t.test", size = 5) +
+  theme(axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_phy.pdf", snp_plot_estim_phy, height = 10, width = 7)
 
@@
   ungroup() %>%
-  dplyr::mutate(dnds_glob = dnds_sum / snps) #divide by total genes
+  dplyr::mutate(dnds_glob = dnds_sum / snps) %>% #divide by total genes
+  dplyr::mutate(bin = str_replace_all(bin, "SPAdes-MaxBin2Refined-", "")) %>% #trim bin names
+  dplyr::mutate(bin = str_replace_all(bin, "SPAdes-MetaBAT2Refined-", ""))
 
@@
                                 fill = "phylum", rotate = TRUE) +
-  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.2) +
-  labs(x = "MAG bin", y = "Global normalized dN/dS ratio")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_estim_bin.pdf", snp_plot_estim_bin, height = 9, width = 8)
+  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.4) +
+  labs(x = "MAG bin", y = "Global normalized dN/dS ratio") +
+  font("legend.title", size = 14) +
+  theme(legend.text = element_text(size = 14),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_bin.pdf", snp_plot_estim_bin, height = 11, width = 8)
 
 # Load lists of PS and NS genes (externally* matched against KEGG KOs from KOfamScan hits)
-snp_kops <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
-snp_kons <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ns_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
+snp_kops <- read.csv("article_profiling/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
+snp_kons <- read.csv("article_profiling/5-SNPs/dndscv/results/kofam/ns_genes_kos.txt", sep = "\t", header = FALSE, stringsAsFactors = FALSE)
 
@@
                           color = "white", palette = "aaas", sort.val = "asc",
-                          rotate = TRUE, xlab = "KO definition", ylab = "Aggregated KOs")
-ggsave("mags-all/5-SNPs/dndscv/results/plot/plot_dnds_kos.pdf", snp_plot_kos, height = 9, width = 12)
+                          rotate = TRUE, xlab = "KO definition", ylab = "Aggregated KOs") +
+  font("legend.title", size = 14) +
+  theme(legend.text = element_text(size = 14),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_kos.pdf", snp_plot_kos, height = 10, width = 15)
 
@@
 }
-snp_cirmag['node_id'] <- snp_nodeid
+snp_cirmag["node_id"] <- snp_nodeid
 snp_cirmag$node_id <- as.integer(snp_cirmag$node_id)
@@
 # Export gene names & contigs
-if (file.exists("mags-all/5-SNPs/dndscv/results/circos/S15.49_genes.txt") == FALSE) {
-  write.table(snp_cirmag$gene_name, "mags-all/5-SNPs/dndscv/results/circos/S15.49_genes.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
-}
-if (file.exists("mags-all/5-SNPs/dndscv/results/circos/contigs/circos_S15.49.txt") == FALSE) {
-  write.table(snp_cirmag$contig, "mags-all/5-SNPs/dndscv/results/circos/contigs/circos_S15.49.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+if (file.exists("article_profiling/5-SNPs/dndscv/results/circos/S15.49_genes.txt") == FALSE) {
+  write.table(snp_cirmag$gene_name, "article_profiling/5-SNPs/dndscv/results/circos/S15.49_genes.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
+}
+if (file.exists("article_profiling/5-SNPs/dndscv/results/circos/contigs/circos_S15.49.txt") == FALSE) {
+  write.table(snp_cirmag$contig, "article_profiling/5-SNPs/dndscv/results/circos/contigs/circos_S15.49.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
 }
@@
 # Read parsed gene annotation (externally* processed)
-snp_cirgenes <- read.csv("mags-all/5-SNPs/dndscv/results/circos/S15.49_genes_pos.txt", sep = "\t", stringsAsFactors = FALSE)
+snp_cirgenes <- read.csv("article_profiling/5-SNPs/dndscv/results/circos/S15.49_genes_pos.txt", sep = "\t", stringsAsFactors = FALSE)
 
@@
 # Use global positions for genes (externally* processed)
-snp_cirgenpos <- read.csv("mags-all/5-SNPs/dndscv/results/circos/contigs/circos_S15.49_relpos.txt", sep = "\t", stringsAsFactors = FALSE)
+snp_cirgenpos <- read.csv("article_profiling/5-SNPs/dndscv/results/circos/contigs/circos_S15.49_relpos.txt", sep = "\t", stringsAsFactors = FALSE)
 snp_cirgenes <- left_join(snp_cirgenes, snp_cirgenpos, by = c("contig"))
@@
 circos.clear()
-pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_circos.pdf", height = 4, width = 4)
+pdf("article_profiling/5-SNPs/dndscv/results/plot/snp_dnds_circos.pdf", height = 4, width = 4.5)
 circos.par(cell.padding = c(0, 0))
@@
   #            snp_cirmagname, facing = "bending.inside", niceFacing = TRUE)
-  text(0, 0, snp_cirmagname, cex = 0.65)
-  text(0, -0.14, paste("Genome size 2703949 bp", "phylum Actinobacteriota | genus Acidithrix",
-                       "168 contigs | N50 26686 | GC 47.7", sep = "\n"), cex = 0.55)
+  text(0, 0, "S15.49", cex = 0.7)
+  text(0, -0.15, paste("Genome size 2703949 bp", "phylum Actinobacteriota | genus Acidithrix",
+                       "168 contigs | N50 26686 | GC 47.7", sep = "\n"), cex = 0.6)
   circos.axis(h = "top", major.at = snp_cirran, labels = paste0(round(snp_cirran/10^6, 1)), #, " Mb"
-              labels.cex = 0.35, lwd = 0.7, labels.facing = "clockwise")
+              labels.cex = 0.5, lwd = 0.7, labels.facing = "clockwise") #0.35
 }, bg.col = "#6a9b85", bg.border = FALSE, track.height = 0.06)
-circos.genomicLabels(snp_cirgenes, labels.column = 7, cex = 0.35, line_lwd = 0.5,
+circos.genomicLabels(snp_cirgenes, labels.column = 7, cex = 0.45, line_lwd = 0.5,
                      side = "inside", line_col = "black", col = "black",
-                     connection_height = 0.05, labels_height = 0.06)
+                     connection_height = 0.06, labels_height = 0.06)
 dev.off()
@@
 ## PS genes for enrichment (we need KEGG KOs from KOfamScan)
-#snp_psenrich <- read.csv("mags-all/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = #FALSE, stringsAsFactors = FALSE)
+#snp_psenrich <- read.csv("article_profiling/5-SNPs/dndscv/results/kofam/ps_genes_kos.txt", sep = "\t", header = #FALSE, stringsAsFactors = FALSE)
 #
@@
 ## Dotplot
-#pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichdot.pdf", height = 9, width = 7)
+#pdf("article_profiling/5-SNPs/dndscv/results/plot/snp_dnds_enrichdot.pdf", height = 9, width = 7)
 #dotplot(snp_keggenrich, showCategory = 30)
@@
 ## Barplot
-#pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichbar.pdf", height = 9, width = 7)
+#pdf("article_profiling/5-SNPs/dndscv/results/plot/snp_dnds_enrichbar.pdf", height = 9, width = 7)
 #barplot(snp_keggenrich, showCategory = 30)
@@
 ## Category netplot
-#pdf("mags-all/5-SNPs/dndscv/results/plot/snp_dnds_enrichnet.pdf", height = 9, width = 13)
+#pdf("article_profiling/5-SNPs/dndscv/results/plot/snp_dnds_enrichnet.pdf", height = 9, width = 13)
 #cnetplot(snp_keggenrich, showCategory = 30, categorySize = "pvalue", foldChange = snp_psenrich$KO)
@@
 # Merge plots using patchwork
-fig_3a <- wrap_plots(plot_spacer(), snp_plot_psgenes, snp_plot_estim_phy,
-                     snp_plot_estim_bin, plot_spacer(),
-                     nrow = 1, ncol = 5, widths = c(0.05, 0.75, 0.55, 0.65, 0.15))
-fig_3b <- wrap_plots(snp_plot_nsgenes, snp_plot_nsgensub, snp_plot_kos,
-                     nrow = 1, ncol = 3, widths = c(0.85, 0.85, 1))
-fig_3 <- wrap_plots(fig_3a, fig_3b, nrow = 2, heights = c(11.5, 10.5)) +
-  plot_annotation(tag_levels = list(c("a", "", "b", "c", "d", "e", "f"))) &
-  theme(plot.tag =  element_text(face = "bold"))
-ggsave("mags-all/fig_3.pdf", fig_3, height = 22, width = 28)
-#ggsave("mags-all/fig_3.png", fig_3, height = 22, width = 28)
-rm(fig_3a, fig_3b) #clean memory
+fig_3a <- wrap_plots(snp_plot_estim_bin, snp_plot_estim_phy, snp_plot_nsgenes,
+                     design = "AABBCC", heights = c(1, 0.9, 0.9)) +
+  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+ggsave("article_profiling/fig_3abc.pdf", fig_3a, height = 12, width = 21)
+#ggsave("article_profiling/fig_3abc.png", fig_3a, height = 12, width = 21)
+fig_3b <- wrap_plots(snp_plot_nsgensub, snp_plot_psgenes,
+                     design = "DDEE", heights = c(0.9, 1)) +
+  plot_annotation(tag_levels = list(c("d", "e"))) &
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+ggsave("article_profiling/fig_3de.pdf", fig_3b, height = 12, width = 20)
+#ggsave("article_profiling/fig_3de.png", fig_3b, height = 12, width = 20)
+fig_3c <- wrap_plots(snp_plot_kos, design = "FFF") +
+  plot_annotation(tag_levels = list(c("f"))) &
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+ggsave("article_profiling/fig_3f.pdf", fig_3c, height = 10, width = 15)
+#ggsave("article_profiling/fig_3f.png", fig_3c, height = 10, width = 15)
 
@@
   plot_annotation(tag_levels = list(c("a", "", ""))) &
-  theme(plot.tag =  element_text(face = "bold"))
-ggsave("mags-all/fig_s3.pdf", fig_s3, height = 9, width = 13)
-#ggsave("mags-all/fig_s3.png", fig_s3, height = 9, width = 13)
+  theme(plot.tag = element_text(size = 22, face = "bold"))
+#ggsave("article_profiling/fig_s3.pdf", fig_s3, height = 11, width = 17)
+#ggsave("article_profiling/fig_s3.png", fig_s3, height = 11, width = 17)
 ```

### 2024-07-26 R/paper.Rmd
revised manuscript & figures

@@
 
-# Inspect completeness and contamination per sample, then select one
-filter(mags_qc, grepl("\\-S15\\.", bin)) %>% dplyr::select(2:3) %>% summary()
-mag_samples <- filter(gtdb_taxonomy, grepl("\\-S15\\.", genome))
+# Inspect and select sample S15
+#filter(mags_qc, grepl("\\-S15\\.", bin)) %>% dplyr::select(2:3) %>% summary()
+mag_samples <- dplyr::filter(gtdb_taxonomy, grepl("\\-S15\\.", genome))
 
@@
 mags_metadata <- merge(mag_samples, smag_data, all = TRUE)
-if (file.exists("4-mags-smag/mags_metadata.tsv") == FALSE) {
 write.table(mags_metadata, "4-mags-smag/mags_metadata.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
-}
 
 # Save parsed SMAG metadata
-if (file.exists("4-mags-smag/smag_data.tsv") == FALSE) {
 write.table(smag_data, "4-mags-smag/smag_data.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
-}
-
-# Save biomining metadata (S15)
-if (file.exists("4-mags-smag/mags_biomining.tsv") == FALSE) {
-  write.table(mag_samples, "4-mags-smag/mags_biomining.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
-}
+
+# Save metadata for sample S15
+write.table(mag_samples, "4-mags-smag/mags_mining_S15.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
 
@@
 
+# Change ecosystem from "biomining" to "mining"
+mag_samples <- mag_samples %>% dplyr::mutate(ecosystem = "Mining")
+mags_metadata_bac <- mags_metadata_bac %>%
+  dplyr::mutate(ecosystem = case_when(ecosystem == "Biomining" ~ "Mining", TRUE ~ ecosystem))
+
 # Select top phyla and plot a histogram
@@
   sort(TRUE) %>%
-  head(n = 40) %>%
+  head(n = 40) %>% #top 40
   as.data.frame() %>%
@@
   geom_text(aes(label = freq), vjust = 0.5, hjust = -0.25, size = 3) +
-  labs(x = "Number", y = "Phylum") +
+  labs(x = "Number of MAGs", y = "Phylum") +
   scale_x_continuous(limits = c(0, 10000)) +
   theme_classic() +
-  theme(axis.text = element_text(size = 14),
-        axis.title = element_text(size = 16))
-ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)
+  theme(axis.text = element_text(size = 15),
+        axis.title = element_text(size = 17))
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 8, width = 6)
 
@@
 #mags_colors_eco <- sample(mags_colors_eco)
-mags_colors_eco <- c("#1F78B4", "#FF7F00", "#B2DF8A", "#E31A1C", "#A6CEE3",
-                     "#6A3D9A", "#33A02C", "#CAB2D6", "#FDBF6F", "#FB9A99")
-
-# Which bacteria from biomining sites are annotated as known?
+mags_colors_eco <- c("#1F78B4", "#FF7F00", "#B2DF8A", "#A6CEE3", "#6A3D9A",
+                     "#33A02C", "#E31A1C", "#CAB2D6", "#FDBF6F", "#FB9A99")
+
+# Which bacteria from mining sites are annotated as known?
 mags_depth_spec_biom <- mags_metadata_bac %>%
   dplyr::select(mag_name, family, genus, species, ecosystem) %>%
-  dplyr::filter(species != "" & ecosystem == "Biomining")
+  dplyr::filter(species != "" & ecosystem == "Mining")
 
@@
   geom_bar(aes(x = species, fill = phylum)) +
-  labs(x = "", y = "Number of kSGBs from biomining") +
+  labs(x = "", y = "Number of kSGBs from mining") +
   scale_fill_manual(values = "#78acc2") +
@@
   theme(legend.title = element_blank(),
-        legend.position = c(0.88, 0.12),
+        legend.position = "top", #c(0.88, 0.12),
         strip.background = element_blank(),
-        axis.text = element_text(size = 12),
-        axis.title = element_text(size = 14),
-        strip.text.x = element_text(size = 10)) +
+        axis.text = element_text(size = 13),
+        axis.title = element_text(size = 15),
+        strip.text.x = element_text(size = 11)) +
   facet_grid(.~ecosystem)
@@
 mags_depth_spec_eco <- mags_depth_spec_eco %>%
-  dplyr::mutate(mags = replace(mags, ecosystem == "Biomining", 1)) %>%
+  dplyr::mutate(mags = replace(mags, ecosystem == "Mining", 1)) %>%
   dplyr::mutate(ratio = count_eco / mags) #max 6.52
@@
   geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.25, size = 3) +
-  labs(x = "Ecosystem", y = "Number of kSGBs/MAGs") +
+  labs(x = "Ecosystem", y = "Number of kSGBs/samples") +
   scale_y_continuous(limits = c(0, 10)) +
@@
         strip.placement = "outside",
+        legend.text = element_text(size = 10),
         axis.text = element_text(size = 12),
@@
         axis.title = element_text(size = 14))
-
 #merge
@@
 
-# And unknown/known SGBs for biomining?
+# And unknown/known SGBs for mining?
 mags_depth_sgbs_biom <- mags_metadata_bac %>%
-  dplyr::filter(ecosystem == "Biomining") %>%
+  dplyr::filter(ecosystem == "Mining") %>%
   dplyr::group_by(phylum, species) %>%
@@
   scale_y_continuous(limits = c(0, 25)) +
-  labs(x = "Phylum", y = "Biomining species") +
+  labs(x = "Phylum", y = "Mining species") +
   coord_flip() +
@@
         panel.border = element_blank(),
-        axis.text = element_text(size = 14),
-        axis.title = element_text(size = 16),
-        strip.text.x = element_text(size = 10)) +
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 18),
+        strip.text.x = element_text(size = 12)) +
   facet_grid(.~ecosystem, scales = "free")
-ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 9, width = 14)
-
-# What about the annotated bins at genus level (kGGBs) from biomining?
+ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 11, width = 17)
+
+# What about the annotated bins at genus level (kGGBs) from mining?
 mags_depth_kggbs_biom <- mags_metadata_bac %>%
   dplyr::select(mag_name, mag_id, phylum, family, genus, ecosystem) %>%
-  dplyr::filter(genus != "" & ecosystem == "Biomining") #nrow=21
+  dplyr::filter(genus != "" & ecosystem == "Mining") #nrow=21
 
@@
   scale_fill_manual(values = mags_colors_eco) +
-  labs(x = "% kGGBs", y = "Known genus from biomining") +
+  labs(x = "% kGGBs", y = "kGGBs from mining") +
   theme_classic() +
@@
 #  ungroup()
-#unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Biomining", "class"])
+#unique(mags_md_unk_gen2[mags_md_unk_gen2[, "ecosystem"] == "Mining", "class"])
 #
-##unique genus from biomining: Acidithrix, Ferrithrix, SCQM01
+##unique genus from mining: Acidithrix, Ferrithrix, SCQM01
 #unique(mag_samples[mag_samples[, "family"] != "", "genus"])
@@
 
-# Select S15 sample
+# Select sample S15 and export to file
 mags_qc <- mags_qc %>% filter(grepl("Refined-S15", bin))
+write.table(mags_qc, "article_profiling/2-statistics/mags_S15.txt", sep = "\t", quote = FALSE, row.names = FALSE)
 
@@
   scale_fill_manual(values = c("#db7971", "#5ca650", "#859ac7")) +
-  labs(x = "Number of MAGs", y = "") +
+  labs(x = "Number of MAGs", y = "MIMAG") +
   theme_classic() +
   theme(aspect.ratio = 0.875,
+        legend.position = "none",
         axis.text = element_text(size = 12),
-        axis.title.x = element_text(size = 14),
-        legend.position = "none")
+        axis.title = element_text(size = 14))
 ggsave("article_profiling/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 4, width = 6)
@@
   mutate_all(function(x) gsub("d__|p__|c__|o__|f__|g__|s__", "", x))
+
+##export to file parsed gtdb_taxonomy 1-taxonomy/gtdbtk_mags_S15.tsv
+##checkm_parsed_bins.tsv 2-statistics/checkm_mags_S15.tsv
 
@@
 
-## Figure 1 & supplementary 1
+## Figures 1-2 & supplementary 1
 
@@
 # Merge plots using patchwork
-fig_1a <- wrap_plots(plot_tree_s15_md, plot_mags_qc, plot_depth_sgbs_biom,
-                     design = "\nAAAB\nAAAC", heights = c(1, 0.85))
-#ggsave("article_profiling/fig_1a.pdf", fig_1a, height = 8, width = 15)
-fig_1b <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco, plot_depth_kggbs_biom,
-                     design = "\nAAABBCC\n")
-#ggsave("article_profiling/fig_1b.pdf", fig_1b, height = 8, width = 15)
-
-fig_1 <- wrap_plots(fig_1a, fig_1b, nrow = 2, heights = c(1, 0.85)) +
-  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "", "e", "f"))) &
-  theme(plot.tag = element_text(face = "bold"))
-ggsave("article_profiling/fig_1.pdf", fig_1, height = 16, width = 16)
-#ggsave("article_profiling/fig_1.png", fig_1, height = 16, width = 16)
-rm(fig_1a, fig_1b) #clean
+fig_1 <- wrap_plots(plot_mags_qc, plot_depth_sgbs_biom, plot_tree_s15_md,
+                    design = "\nACCC\nBCCC", heights = c(0.85, 1)) +
+  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
+  theme(plot.tag = element_text(size = 18, face = "bold"))
+ggsave("article_profiling/fig_1.pdf", fig_1, height = 8, width = 15)
+ggsave("article_profiling/fig_1.png", fig_1, height = 8, width = 15)
+
+fig_2 <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco,
+                    plot_depth_kggbs_biom, design = "\nAAABBCC\n") +
+  plot_annotation(tag_levels = list(c("a", "", "b", "c"))) &
+  theme(plot.tag = element_text(size = 18, face = "bold"))
+ggsave("article_profiling/fig_2.pdf", fig_2, height = 8, width = 16)
+ggsave("article_profiling/fig_2.png", fig_2, height = 8, width = 16)
 
@@
 fig_s1 <-  { plot_depth_top_phyla |
-    (plot_depth_usgbs / plot_depth_spec + plot_layout(heights = c(0.85, 0.15))) } +
+    (plot_depth_spec / plot_depth_usgbs + plot_layout(heights = c(0.15, 0.85))) } +
   plot_layout(ncol = 2, widths = c(0.25, 1)) +
   plot_annotation(tag_levels = list(c("a", "b", "c"))) &
-  theme(plot.tag = element_text(size = 20, face = "bold"))
-fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.title.y = element_text(vjust = 0))
-fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.line.y.left = element_blank(),
+  theme(plot.tag = element_text(size = 18, face = "bold"))
+fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.line.y.left = element_blank(),
                                              axis.ticks.y.left = element_blank())#,
                                              #axis.text.y = element_text(hjust = 2.2))
-ggsave("article_profiling/fig_s1.pdf", fig_s1, height = 13, width = 22)
-#ggsave("article_profiling/fig_s1.png", fig_s1, height = 13, width = 22)
+fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.title.y = element_text(vjust = 0))
+ggsave("article_profiling/fig_s1.pdf", fig_s1, height = 15, width = 24)
+ggsave("article_profiling/fig_s1.png", fig_s1, height = 15, width = 24)
 ```
@@
 # Read output matrix from KOfamScan
-kegg_annot <- read.csv("/mnt/beegfs/home/mrojas/DiGenomaLab/Systemix/analysis/kegg/merge_hits/kofamscan_hits.tsv", sep = "\t", check.names = FALSE)
+kegg_annot <- read.csv("article_profiling/4-mags-smag/kofamscan_hits.tsv", sep = "\t", check.names = FALSE)
 colnames(kegg_annot)[2] <- "KO.definition"
@@
   #soxr: SoxR 2Fe-2S cluster
-  #nitrogen: a negative marker (ubiquous; ex: MerR family)
-  dplyr::filter(grepl("iron|sulf|soxr|copper|nitrogen",
-                      Definition, ignore.case = TRUE) == TRUE) %>%
+  dplyr::filter(grepl("iron|sulf|soxr|copper", Definition, ignore.case = TRUE) == TRUE) %>%
   dplyr::select(KO, Definition)
@@
     return("Sulfur")
-  if (grepl("nitrogen", def, ignore.case = TRUE) == TRUE)
-      return("Nitrogen")
   else
@@
   if (grepl("spades", mag, ignore.case = TRUE) == TRUE) {
-    return("Biomining")
+    return("Mining")
   }
@@
 colnames(kegg_mags) <- c("mag")
-
 kegg_mags_eco <- vector()
@@
 # Explore ecosystems..
-table(kegg_mags$ecosystem)
-kegg_mags %>% dplyr::filter(ecosystem == "Biomining") %>% dplyr::summarise(total = sum(length))
-kegg_mags %>% dplyr::group_by(ecosystem) %>% dplyr::summarise(total = sum(length))
+#table(kegg_mags$ecosystem)
+#kegg_mags %>% dplyr::filter(ecosystem == "Mining") %>% dplyr::summarise(total = sum(length))
+#kegg_mags %>% dplyr::group_by(ecosystem) %>% dplyr::summarise(total = sum(length))
 ```
@@
 # Plot using ComplexHeatmap
-kegg_hm_colors <- colorRamp2(c(-2, 0, 2), c("#5073AD", "white", "#D33D35"))
+kegg_hm_colors <- colorRamp2(c(-2, 0, 2), c("#5073ad", "white", "#d33d35"))
 
@@
 # Annotate rows: KO pathways
-kegg_hm_rows_path <- c("Iron" = "#666666", "Sulfur" = "#DDAD3B", "Copper" = "#CA6627")
+kegg_hm_rows_path <- c("Iron" = "#666666", "Sulfur" = "#ddad3b", "Copper" = "#ca6627")
 kegg_hm_rows_path <- list(Def = kegg_hm_paths$Pathway, Pathways = kegg_hm_rows_path)
@@
 # Plot heatmap with annotations
-pdf("plot_annot_heatmap.pdf", height = 9, width = 9)
-kegg_hm_plot <- Heatmap(kegg_filt_norm, name = "Abundance",
+pdf("article_profiling/4-mags-smag/2-plots-annotation/heatmap/plot_annot_heatmap.pdf", height = 9, width = 9)
+kegg_hm_plot <- Heatmap(kegg_hm_data, name = "Abundance",
                         row_title = "KEGG KOs", column_title = "Ecosystem MAGs",
@@
 
-### PCA
+### PCA (**discarded**)
 
@@
 # Count total KOs and length in all across ecosystems for selected pathways
-# (time-consuming loop: runtime ~12 mins for rows in kegg_paths*)
+# (loop runtime ~2 mins for rows in kegg_paths)
 kegg_bubble_counts <- tibble(ecosystem = character(), count = numeric(), length = numeric())
@@
 #explore
-kegg_bubble %>% dplyr::group_by(ecosystem) %>% dplyr::summarise(n = sum(count_norm))
+#kegg_bubble %>% dplyr::group_by(ecosystem) %>% dplyr::summarise(n = sum(count_norm)) %>% dplyr::arrange(desc(n))
 
@@
 kegg_bubble_keep <- kegg_bubble_cu %>%
-  dplyr::filter(ecosystem == "Biomining" & count_norm > 0) %>%
+  dplyr::filter(ecosystem == "Mining" & count_norm > 0) %>%
   dplyr::select(KO)
@@
 kegg_bubble_keep <- kegg_bubble_fe %>%
-  dplyr::filter(ecosystem == "Biomining" & count_norm > 0.01) %>%
+  dplyr::filter(ecosystem == "Mining" & count_norm > 0.01) %>%
   dplyr::select(KO)
@@
 kegg_bubble_keep <- kegg_bubble_s %>%
-  dplyr::filter(ecosystem == "Biomining" & count_norm > 0.05) %>%
+  dplyr::filter(ecosystem == "Mining" & count_norm > 0.05) %>%
   dplyr::select(KO)
@@
                            color = "ecosystem", palette = anova_sort_cu$colors,
-                           order = anova_sort_cu$ecosystem, add = "jitter", legend = "none") +
+                           order = anova_sort_cu$ecosystem, add = "jitter", legend = "none",
+                           add.params = list(size = 2, alpha = 0.75)) +
   geom_hline(yintercept = mean(kegg_bubble_cu$count_norm), linetype = 2, linewidth = 0.5) +
-  labs(x = "Ecosystem", y = "Copper pathway (KOs/Mb)") +
+  labs(x = "Ecosystem", y = "Copper proteins (KOs/Mb)") +
   rotate_x_text(angle = 45) + ylim(c(0, 0.3)) +
   #stat_summary(fun = mean, geom = "point", shape = 18, size = 2.5, color = "#FC4E07") +
-  stat_compare_means(method = "anova", label.y = 0.3, size = 5) + #add global annova p-value
+  stat_compare_means(method = "anova", label.y = 0.3, size = 6) + #add global annova p-value
   stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE) +
-  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))
+  theme(axis.text = element_text(size = 15),
+        axis.title = element_text(size = 17))
 ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_boxplot_copper.pdf", anova_boxplot_cu, height = 7, width = 7)
@@
                            color = "ecosystem", palette = anova_sort_fe$colors,
-                           order = anova_sort_fe$ecosystem, add = "jitter", legend = "none") +
+                           order = anova_sort_fe$ecosystem, add = "jitter", legend = "none",
+                           add.params = list(size = 2, alpha = 0.75)) +
   geom_hline(yintercept = mean(kegg_bubble_fe$count_norm), linetype = 2, linewidth = 0.5) +
-  labs(x = "Ecosystem", y = "Iron pathway (KOs/Mb)") +
+  labs(x = "Ecosystem", y = "Iron proteins (KOs/Mb)") +
   rotate_x_text(angle = 45) + ylim(c(0, 0.4)) +
-  stat_compare_means(method = "anova", label.y = 0.4, size = 5) + #add global annova p-value
+  stat_compare_means(method = "anova", label.y = 0.4, size = 6) + #add global annova p-value
   stat_compare_means(label = "p.signif", method = "t.test",
                      ref.group = ".all.", hide.ns = TRUE, label.y = 0.35) +
-  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))
+  theme(axis.text = element_text(size = 15),
+        axis.title = element_text(size = 17))
 ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_boxplot_iron.pdf", anova_boxplot_fe, height = 7, width = 7)
@@
                              color = "ecosystem", palette = anova_sort_s$colors,
-                             order = anova_sort_s$ecosystem, add = "jitter", legend = "none") +
+                             order = anova_sort_s$ecosystem, add = "jitter", legend = "none",
+                             add.params = list(size = 2, alpha = 0.75)) +
   geom_hline(yintercept = mean(kegg_bubble_s$count_norm), linetype = 2, linewidth = 0.5) +
-  labs(x = "Ecosystem", y = "Sulfur pathway (KOs/Mb)") +
+  labs(x = "Ecosystem", y = "Sulfur proteins (KOs/Mb)") +
   rotate_x_text(angle = 45) + ylim(c(0, 0.8)) +
-  stat_compare_means(method = "anova", label.y = 0.8, size = 5) + #add global annova p-value
+  stat_compare_means(method = "anova", label.y = 0.8, size = 6) + #add global annova p-value
   stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.",
-                     hide.ns = TRUE, label.y = 0.75, size = 6) +
-  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 16))
+                     hide.ns = TRUE, label.y = 0.75, size = 8) +
+  theme(axis.text = element_text(size = 15),
+        axis.title = element_text(size = 17))
 ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_boxplot_sulfur.pdf", anova_boxplot_s, height = 7, width = 7)
@@
              y = Definition, size = count_norm, color = ecosystem)) +
-  geom_point(alpha = 0.5) +
+  geom_point(alpha = 0.65) +
   scale_size(range = c(0, 10)) +
@@
   theme_classic() +
-  theme(legend.title = element_text(size = 12), legend.position = "top",
-        axis.text.x = element_text(size = 14, angle = 90, vjust = 1, hjust = 1),
-        axis.text.y = element_text(size = 14), axis.title.x = element_text(size = 16),
-        axis.line.y.right = element_blank(), axis.ticks.y.right = element_blank())
-ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_bubbles.pdf", kegg_bubbles_plot, height = 7, width = 12)
+  theme(legend.position = "top",
+        axis.line.y.right = element_blank(),
+        axis.ticks.y.right = element_blank(),
+        legend.title = element_text(size = 12),
+        axis.text.x = element_text(size = 16, angle = 60, vjust = 1, hjust = 1),
+        axis.text.y = element_text(size = 16),
+        axis.title.x = element_text(size = 18))
+ggsave("article_profiling/4-mags-smag/2-plots-annotation/bubble/plot_annot_bubbles.pdf", kegg_bubbles_plot, height = 8, width = 13)
 
@@
 
-## Figure 2 & supplementary 2
-
-```{r fig2}
+## Figure 3
+
+```{r fig3}
 # Merge plots using patchwork
-fig_2 <- wrap_plots(anova_boxplot_cu, anova_boxplot_fe, anova_boxplot_s, design = "AABBCC") +
+fig_3 <- wrap_plots(anova_boxplot_cu, anova_boxplot_fe, anova_boxplot_s, design = "AABBCC") +
   plot_annotation(tag_levels = list(c("a", "b", "c"))) &
   theme(plot.tag = element_text(size = 22, face = "bold"))
-ggsave("article_profiling/fig_2.pdf", fig_2, height = 7, width = 19)
-ggsave("article_profiling/fig_2.png", fig_2, height = 7, width = 19)
-
-#sup 2
-fig_s2 <- wrap_plots(kegg_bubbles_plot, design = "AAA") +
-  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
-  theme(plot.tag = element_text(size = 20, face = "bold"))
-ggsave("article_profiling/fig_s2a.pdf", fig_s2, height = 7, width = 12)
-#ggsave("article_profiling/fig_s2a.png", fig_s2, height = 7, width = 12)
+ggsave("article_profiling/fig_3.pdf", fig_3, height = 7, width = 20)
+ggsave("article_profiling/fig_3.png", fig_3, height = 7, width = 20)
 ```
@@
   dplyr::mutate(signif = ifelse(pmis_cv < 0.05, "*", "")) %>%
-  dplyr::mutate(path = case_when(gene_name == "KJCAEPPB_04437" ~ "Ni",
-                                 gene_name == "NEBJFEFK_01910" ~ "S", TRUE ~ "")) %>%
+  dplyr::mutate(path = case_when(gene_name == "NEBJFEFK_01910" ~ "S", TRUE ~ "")) %>%
   distinct(name, .keep_all = TRUE) %>% #remove repeated names
@@
   geom_text(aes(label = num), vjust = 0.5, hjust = -0.25, size = 4) +
-  labs(x = "Number of genes under PS", y = "MAG Bin") +
+  labs(x = "Number of genes under PS", y = "MAG") +
   scale_x_continuous(limits = c(0, 255)) +
   theme_classic() +
-  theme(strip.background = element_blank(), axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18), strip.text.x = element_text(size = 14)) +
+  theme(strip.background = element_blank(),
+        strip.text.x = element_text(size = 14),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 20)) +
   facet_grid(.~species)
@@
         axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 snp_plot_psbins3 <- snp_dnds %>%
@@
         axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 #merge
@@
   geom_segment(aes(x = 0, xend = wmis_cv, y = name), color = "gray") +
-  geom_point(alpha = 0.7, size = 2, color = "#e31a1c") +
-  geom_text(aes(label = signif, x = -0.15), size = 6.5, color = "#3c4d76") +
-  geom_text(aes(label = path, x = -0.15), size = 5.5, color = "#3c4d76") +
-  geom_vline(xintercept = 1, linetype = 2, linewidth = 0.4) +
-  geom_vline(xintercept = max(snp_psgenes_sub$wmis_cv), linetype = 1, linewidth = 0.4) +
+  geom_point(alpha = 0.7, size = 3, color = "#e31a1c") +
+  geom_text(aes(label = signif, x = -0.15), size = 7, color = "#3c4d76") +
+  geom_text(aes(label = path, x = -0.15), size = 6, color = "#3c4d76") +
+  geom_vline(xintercept = 1, linetype = 2, linewidth = 0.5) +
+  geom_vline(xintercept = max(snp_psgenes_sub$wmis_cv), linetype = 1, linewidth = 0.5) +
   labs(x = "PS strength", y = "Gene name") +
@@
   theme(axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 #ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psgenes.pdf", snp_plot_psgenes, height = 12, width = 7)
@@
                     values = c("#5db4b2", "#b687b6", "#2374a4")) +
-  labs(x = "Number of mutations", y = "") +
+  labs(x = "# mutations", y = "") +
   theme_classic() +
-  theme(axis.text.y = element_blank(), axis.line.y = element_blank(),
-        legend.position = c(0.88, 0.15), strip.placement = "outside",
-        axis.ticks.y = element_blank(), legend.text = element_text(size = 16),
-        axis.text = element_text(size = 16), axis.title = element_text(size = 18))
+  theme(legend.position = c(0.88, 0.15),
+        strip.placement = "outside",
+        axis.line.y = element_blank(),
+        axis.text.y = element_blank(),
+        axis.ticks.y = element_blank(),
+        legend.text = element_text(size = 14),
+        axis.text = element_text(size = 16),
+        axis.title = element_text(size = 20))
 #ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_psmuts.pdf", snp_plot_muts, height = 12, width = 7)
@@
 
+# Remove nickel due to low counts
+snp_nsgenes <- snp_nsgenes %>% dplyr::filter(pathway != "Nickel")
+
 # Sort pathways and set colors to plot
@@
   as.data.frame()
-snp_sort_ns$color <- c("#ddad3b", "#8c2421", "#666666", "#4f4fad") #"#6a9b85"
+snp_sort_ns$color <- c("#ddad3b", "#8c2421", "#666666") #"#4f4fad", "#6a9b85"
 
@@
                              order = snp_sort_ns$pathway, palette = snp_sort_ns$color,
-                             size = 0.75, add = "jitter", legend = "none") +
+                             size = 0.75, add = "jitter", legend = "none",
+                             add.params = list(size = 2, alpha = 0.75)) +
   scale_y_continuous(limits = c(-0.2, 1.5)) +
@@
   stat_compare_means(method = "anova", label.y = 1.5, size = 6) +
-  stat_compare_means(comparisons = list(c("Copper", "Iron"), c("Sulfur", "Nickel"),
-                                        c("Copper", "Nickel")), label = "p.format",
+  stat_compare_means(comparisons = list(c("Sulfur", "Copper"), c("Sulfur", "Iron"),
+                                        c("Copper", "Iron")), label = "p.format",
                      method = "t.test", size = 5) +
   theme(axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 10, width = 7)
@@
   dplyr::group_by(product, pathway) %>%
-  dplyr::summarise(dnds = mean(wmis_cv))
-snp_nsgenes_sub_col <- c("#8c2421", "#666666", "#4f4fad", "#ddad3b")
+  dplyr::summarise(dnds = mean(wmis_cv), .groups = "drop")
+snp_nsgenes_sub_col <- c("#8c2421", "#666666", "#ddad3b")
 
@@
         axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenesprod.pdf", snp_plot_nsgensub, height = 11, width = 13)
@@
                                 order = snp_estim_sort_phy$phylum, palette = snp_estim_sort_phy$color,
-                                add = "jitter", size = 0.75, legend = "none") +
+                                add = "jitter", size = 0.75, legend = "none",
+                                add.params = list(size = 3, alpha = 0.75)) +
   geom_hline(yintercept = mean(snp_estim_phy$dnds_glob), linetype = 2, linewidth = 0.4) +
@@
   theme(axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_phy.pdf", snp_plot_estim_phy, height = 10, width = 7)
@@
         axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_bin.pdf", snp_plot_estim_bin, height = 11, width = 8)
@@
         axis.text = element_text(size = 16),
-        axis.title = element_text(size = 18))
+        axis.title = element_text(size = 20))
 ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_kos.pdf", snp_plot_kos, height = 10, width = 15)
@@
 
-## Figure 3 & supplementary 3
-
-```{r fig3}
+## Figures 4-5 & supplementary 3
+
+```{r fig4}
 # Merge plots using patchwork
-fig_3a <- wrap_plots(snp_plot_estim_bin, snp_plot_estim_phy, snp_plot_nsgenes,
-                     design = "AABBCC", heights = c(1, 0.9, 0.9)) +
-  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
-  theme(plot.tag = element_text(size = 20, face = "bold"))
-ggsave("article_profiling/fig_3abc.pdf", fig_3a, height = 12, width = 21)
-#ggsave("article_profiling/fig_3abc.png", fig_3a, height = 12, width = 21)
-fig_3b <- wrap_plots(snp_plot_nsgensub, snp_plot_psgenes,
-                     design = "DDEE", heights = c(0.9, 1)) +
-  plot_annotation(tag_levels = list(c("d", "e"))) &
-  theme(plot.tag = element_text(size = 20, face = "bold"))
-ggsave("article_profiling/fig_3de.pdf", fig_3b, height = 12, width = 20)
-#ggsave("article_profiling/fig_3de.png", fig_3b, height = 12, width = 20)
-fig_3c <- wrap_plots(snp_plot_kos, design = "FFF") +
-  plot_annotation(tag_levels = list(c("f"))) &
-  theme(plot.tag = element_text(size = 20, face = "bold"))
-ggsave("article_profiling/fig_3f.pdf", fig_3c, height = 10, width = 15)
-#ggsave("article_profiling/fig_3f.png", fig_3c, height = 10, width = 15)
+fig_4 <- wrap_plots(snp_plot_estim_bin, snp_plot_estim_phy, design = "AABB", heights = c(1, 0.9)) +
+  plot_annotation(tag_levels = list(c("a", "b"))) &
+  theme(plot.tag = element_text(size = 22, face = "bold"))
+ggsave("article_profiling/fig_4.pdf", fig_4, height = 12, width = 15)
+ggsave("article_profiling/fig_4.png", fig_4, height = 12, width = 15)
+
+fig_5a <- wrap_plots(snp_plot_nsgenes, snp_plot_nsgensub, design = "AABB", heights = c(0.9, 1)) +
+  plot_annotation(tag_levels = list(c("a", "b"))) &
+  theme(plot.tag = element_text(size = 22, face = "bold"))
+ggsave("article_profiling/fig_5a.pdf", fig_5a, height = 11, width = 19)
+ggsave("article_profiling/fig_5a.png", fig_5a, height = 11, width = 19)
+
+fig_5b <- wrap_plots(snp_plot_psgenes, snp_plot_kos, design = "AABBB", heights = c(0.9, 1)) +
+  plot_annotation(tag_levels = list(c("c", "d"))) &
+  theme(plot.tag = element_text(size = 22, face = "bold"))
+ggsave("article_profiling/fig_5b.pdf", fig_5b, height = 12, width = 23)
+ggsave("article_profiling/fig_5b.png", fig_5b, height = 12, width = 23)
 

### 2024-08-22 R/paper.Rmd
improved figures & added abundances

@@
 suppressWarnings(library(circlize)) #circular viz
+suppressWarnings(library(ggbeeswarm)) #scatter plots
 ```
 
-## Prepare and merge MAG-SMAG metadata
+## Prepare and merge MAG-SMAG
 
@@
 
-### In-depth biomining MAGs vs SMAG comparison
+### 1.1: In-depth biomining MAGs vs SMAG
 
@@
 
-### QC
+### 1.2: MAGs QC
 
@@
 
-### Phylogenetic tree
+### 1.3: Phylogenetic tree
 
@@
 
+## Section 2: Abundances
+
+```{r abund}
+# Read matrix of raw abundances
+abu_depths <- read.csv("article_profiling/3-abundances/raw/bin_refined_depths_summary.tsv", sep = "\t")
+
+# Filter list to S15 sample
+abu_depths <- abu_depths %>%
+  dplyr::filter(!grepl("Unbinned", bin)) %>%
+  dplyr::filter(grepl("S15", bin)) %>%
+  dplyr::mutate(bin = str_replace_all(bin, ".fa", "")) %>%
+  dplyr::select(bin, S15)
+
+# Add phylum
+abu_depths <- left_join(abu_depths, dplyr::select(gtdb_taxonomy, label, phylum),
+                        by = join_by(bin == label))
+
+# Dotplot of global abundances by MAG size
+abu_plot_depth <- abu_depths %>%
+  dplyr::mutate(bin = str_replace_all(bin, "SPAdes-MaxBin2Refined-", "")) %>% #trim bin names
+  dplyr::mutate(bin = str_replace_all(bin, "SPAdes-MetaBAT2Refined-", "")) %>%
+  ggplot(aes(x = S15, y = reorder(bin, S15), color = phylum)) +
+  geom_point(aes(size = S15), alpha = 0.7) +
+  scale_size(range = c(0, 10)) +
+  scale_color_manual(values = tree_colors) +
+  scale_x_log10(expand = c(0.1, 0)) + #add x-axis space
+  scale_y_discrete(expand = c(0.025, 0)) + #add y-axis space
+  guides(color = guide_legend(title = "Phylum", ncol = 1), size = "none") +
+  labs(x = "Abundance (mean coverage)", y = "MAG") +
+  theme_classic() +
+  theme(legend.text = element_text(size = 11),
+        legend.title = element_text(size = 13),
+        axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
+ggsave("article_profiling/3-abundances/plots/plot_abu_rank.pdf", abu_plot_depth, height = 8, width = 6)
+
+# Define top phyla
+abu_top_phy <- c("Acidobacteriota", "Actinobacteriota", "Proteobacteria")
+
+# Boxplot of abundances by top phyla
+abu_plot_depth.2 <- abu_depths %>%
+  dplyr::filter(phylum %in% abu_top_phy) %>%
+  dplyr::mutate(phylum = factor(phylum, levels = c("Actinobacteriota", "Acidobacteriota", "Proteobacteria"))) %>% #sort
+  ggplot(aes(x = phylum, y = S15, color = phylum)) +
+  geom_quasirandom(aes(size = S15), alpha = 0.7) +
+  scale_size(range = c(0, 10)) +
+  scale_y_log10() +
+  geom_boxplot(fill = NA, outlier.size = 0, outliers = FALSE) +
+  scale_color_manual(values = c("#98d277", "#a6cee3", "#b15928")) +
+  guides(size = "none", color = "none") +
+  stat_compare_means(method = "anova", label.y = 3.8, size = 4.5) + #pos by log10
+  stat_compare_means(method = "t.test", label = "p.format", ref.group = ".all.", vjust = 0.25,
+                     comparisons = list(c("Actinobacteriota", "Acidobacteriota"),
+                                        c("Actinobacteriota", "Proteobacteria"),
+                                        c("Acidobacteriota", "Proteobacteria")),
+                     label.y = c(3.3, 3.45, 3.64), bracket.size = 0.25, size = 4) +
+  labs(x = "Phylum", y = "Abundance (mean coverage)") +
+  theme_classic() +
+  theme(axis.text = element_text(size = 12),
+        axis.title = element_text(size = 14))
+ggsave("article_profiling/3-abundances/plots/plot_abu_phy.pdf", abu_plot_depth.2, height = 7, width = 5)
+```
+
 ## Figures 1-2 & supplementary 1
@@
 # Merge plots using patchwork
-fig_1 <- wrap_plots(plot_mags_qc, plot_depth_sgbs_biom, plot_tree_s15_md,
-                    design = "\nACCC\nBCCC", heights = c(0.85, 1)) +
+fig_1a <- wrap_plots(plot_mags_qc, plot_depth_sgbs_biom, plot_tree_s15_md,
+                     design = "ACCC\nBCCC", widths = c(0.6, 1)) +
   plot_annotation(tag_levels = list(c("a", "b", "c"))) &
-  theme(plot.tag = element_text(size = 18, face = "bold"))
-ggsave("article_profiling/fig_1.pdf", fig_1, height = 8, width = 15)
-ggsave("article_profiling/fig_1.png", fig_1, height = 8, width = 15)
+  theme(plot.tag = element_text(size = 16, face = "bold"))
+ggsave("article_profiling/fig_1a.pdf", fig_1, height = 8.5, width = 13)
+#ggsave("article_profiling/fig_1a.png", fig_1, height = 8.5, width = 13)
+fig_1b <- wrap_plots(abu_plot_depth, abu_plot_depth.2, widths = c(0.65, 1)) +
+  plot_annotation(tag_levels = list(c("d", "e"))) &
+  theme(plot.tag = element_text(size = 16, face = "bold"))
+ggsave("article_profiling/fig_1b.pdf", fig_1b, height = 8, width = 11)
+#ggsave("article_profiling/fig_1b.png", fig_1b, height = 8, width = 11)
+#compose
+fig_1 <- wrap_plots(fig_1a / fig_1b) +
+  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "e"))) &
+  theme(plot.tag = element_text(size = 16, face = "bold"))
+ggsave("article_profiling/fig_1.pdf", fig_1, height = 16, width = 13)
+ggsave("article_profiling/fig_1.png", fig_1, height = 16, width = 13)
+rm(fig_1a, fig_1b) #clean memory
 
 fig_2 <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco,
-                    plot_depth_kggbs_biom, design = "\nAAABBCC\n") +
+                    plot_depth_kggbs_biom, design = "AAABBCC") +
   plot_annotation(tag_levels = list(c("a", "", "b", "c"))) &
@@
 
-## Section 2: Functional Annotation
-
-### Pre-process matrix of KEGG KOs from KOfamScan
+## Section 3: Functional Annotation
+
+### 3.1: Process KEGG KOs
 
@@
 
-### Heatmap
+### 3.2: Heatmap
 
@@
 
-### PCA (**discarded**)
+### 3.3: PCA (**discarded**)
 
@@
 
-### Bubble and boxplots
+### 3.4: Bubble/boxplots
 
@@
 
-## Section 3: Evolutionary analysis
-
-### dN/dS
+## Section 4: Evolutionary analysis
 
@@
 # Violin plot for NS genes with global mutations
-snp_plot_nsgenes <- ggviolin(snp_nsgenes, x = "pathway", y = "wmis_cv", color = "pathway",
-                             shape = "pathway", xlab = "Pathway", ylab = "dN/dS",
-                             order = snp_sort_ns$pathway, palette = snp_sort_ns$color,
-                             size = 0.75, add = "jitter", legend = "none",
-                             add.params = list(size = 2, alpha = 0.75)) +
+snp_plot_nsgenes <- snp_nsgenes %>%
+  ggviolin(x = "pathway", y = "wmis_cv", color = "pathway", add = "jitter", shape = "pathway",
+           order = snp_sort_ns$pathway, palette = snp_sort_ns$color, legend = "none",
+           size = 0.55, add.params = list(size = 3, alpha = 0.7)) +
   scale_y_continuous(limits = c(-0.2, 1.5)) +
@@
   geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.4) +
-  stat_compare_means(method = "anova", label.y = 1.5, size = 6) +
-  stat_compare_means(comparisons = list(c("Sulfur", "Copper"), c("Sulfur", "Iron"),
-                                        c("Copper", "Iron")), label = "p.format",
-                     method = "t.test", size = 5) +
+  labs(x = "Pathway", y = "dN/dS") +
+  stat_compare_means(method = "anova", label.y = 1.475, size = 6) +
+  stat_compare_means(method = "t.test", label = "p.format", vjust = 0.25,
+                     comparisons = list(c("Sulfur", "Copper"), c("Sulfur", "Iron"), c("Copper", "Iron")),
+                     label.y = c(1.2, 1.27, 1.36), bracket.size = 0.25, size = 5) +
   theme(axis.text = element_text(size = 16),
-        axis.title = element_text(size = 20))
-ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 10, width = 7)
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenes.pdf", snp_plot_nsgenes, height = 9, width = 6)
 
@@
 
-# Lollipop plot of grouped NS gene products
-snp_plot_nsgensub <- ggdotchart(snp_nsgenes_sub, x = "product", y = "dnds", color = "pathway",
-                                xlab = "Product of genes under NS", ylab = "Mean dN/dS",
-                                y.text.col = TRUE, rotate = TRUE, palette = snp_nsgenes_sub_col,
-                                dot.size = 3, sorting = "descending", ggtheme = theme_pubr()) +
+# Dotplot of grouped products of NS genes
+snp_plot_nsgensub <- snp_nsgenes_sub %>%
+  dplyr::rename(Pathway = pathway) %>% #rename col.
+  ggdotchart(x = "product", y = "dnds", color = "Pathway", y.text.col = TRUE, rotate = TRUE,
+             palette = snp_nsgenes_sub_col, dot.size = 3, sorting = "descending") + #add = "segments"
   scale_y_continuous(limits = c(0, 1)) +
   geom_hline(yintercept = mean(snp_nsgenes$wmis_cv), linetype = 2, linewidth = 0.4) +
-  geom_hline(yintercept = 1, linetype = 1, linewidth = 0.4) +
-  font("legend.title", size = 14) +
+  labs(x = "NS gene products", y = "Mean dN/dS") +
+  font("legend.title", size = 15) +
   theme(legend.text = element_text(size = 14),
-        axis.text = element_text(size = 16),
-        axis.title = element_text(size = 20))
-ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenesprod.pdf", snp_plot_nsgensub, height = 11, width = 13)
+        axis.text = element_text(size = 15),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_nsgenesprod.pdf", snp_plot_nsgensub, height = 10, width = 11)
 
@@
 # Boxplot global dN/dS ratios by phylum
-snp_plot_estim_phy <- ggboxplot(snp_estim_phy, x = "phylum", y = "dnds_glob", color = "phylum",
+snp_plot_estim_phy <- snp_estim_phy %>%
+  ggboxplot(x = "phylum", y = "dnds_glob", color = "phylum", add = "jitter",
             order = snp_estim_sort_phy$phylum, palette = snp_estim_sort_phy$color,
-                                add = "jitter", size = 0.75, legend = "none",
-                                add.params = list(size = 3, alpha = 0.75)) +
+             size = 0.55, legend = "none", add.params = list(size = 3.25, alpha = 0.7)) +
   geom_hline(yintercept = mean(snp_estim_phy$dnds_glob), linetype = 2, linewidth = 0.4) +
@@
   labs(x = "Phylum", y = "Global normalized dN/dS ratio") +
-  stat_compare_means(method = "anova", label.y = 1.6, size = 6) +
-  stat_compare_means(comparisons = list(c("Proteobacteria", "Actinobacteriota"),
+  stat_compare_means(method = "anova", label.y = 1.475, size = 6) +
+  stat_compare_means(method = "t.test", label = "p.format", vjust = 0.25,
+                     comparisons = list(c("Proteobacteria", "Actinobacteriota"),
                                         c("Proteobacteria", "Acidobacteriota"),
                                         c("Actinobacteriota", "Acidobacteriota")),
-                     label = "p.format", method = "t.test", size = 5) +
-  theme(axis.text = element_text(size = 16),
-        axis.title = element_text(size = 20))
-ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_phy.pdf", snp_plot_estim_phy, height = 10, width = 7)
+                     label.y = c(1.25, 1.325, 1.4), bracket.size = 0.25, size = 5) +
+  theme(axis.text = element_text(size = 15),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_phy.pdf", snp_plot_estim_phy, height = 9, width = 6)
 
@@
 # Barplot global dN/dS ratios by MAG bins
-snp_plot_estim_bin <- ggbarplot(snp_estim_bin, x = "bin", y = "dnds_glob", color = "phylum",
-                                order = snp_estim_sort_bin$bin, palette = tree_colors,
-                                fill = "phylum", rotate = TRUE) +
+snp_plot_estim_bin <- snp_estim_bin %>%
+  dplyr::rename(Phylum = phylum) %>% #rename col.
+  ggbarplot(x = "bin", y = "dnds_glob", color = "Phylum", fill = "Phylum", legend = "right",
+            order = snp_estim_sort_bin$bin, palette = tree_colors, rotate = TRUE) +
   geom_hline(yintercept = 1, linetype = 1, linewidth = 0.4) +
-  labs(x = "MAG bin", y = "Global normalized dN/dS ratio") +
-  font("legend.title", size = 14) +
+  labs(x = "MAG", y = "Global normalized dN/dS ratio") +
+  font("legend.title", size = 15) +
   theme(legend.text = element_text(size = 14),
-        axis.text = element_text(size = 16),
-        axis.title = element_text(size = 20))
-ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_bin.pdf", snp_plot_estim_bin, height = 11, width = 8)
+        axis.text = element_text(size = 15),
+        axis.title = element_text(size = 18))
+ggsave("article_profiling/5-SNPs/dndscv/results/plot/plot_dnds_estim_bin.pdf", snp_plot_estim_bin, height = 9, width = 7)
 
@@
 # Merge plots using patchwork
-fig_4 <- wrap_plots(snp_plot_estim_bin, snp_plot_estim_phy, design = "AABB", heights = c(1, 0.9)) +
+fig_4 <- wrap_plots(snp_plot_estim_bin, snp_plot_estim_phy, widths = c(0.6, 1)) +
   plot_annotation(tag_levels = list(c("a", "b"))) &
-  theme(plot.tag = element_text(size = 22, face = "bold"))
-ggsave("article_profiling/fig_4.pdf", fig_4, height = 12, width = 15)
-ggsave("article_profiling/fig_4.png", fig_4, height = 12, width = 15)
-
-fig_5a <- wrap_plots(snp_plot_nsgenes, snp_plot_nsgensub, design = "AABB", heights = c(0.9, 1)) +
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+ggsave("article_profiling/fig_4.pdf", fig_4, height = 10, width = 13)
+ggsave("article_profiling/fig_4.png", fig_4, height = 10, width = 13)
+
+fig_5a <- wrap_plots(snp_plot_nsgenes, snp_plot_nsgensub, widths = c(1, 1)) +
   plot_annotation(tag_levels = list(c("a", "b"))) &
-  theme(plot.tag = element_text(size = 22, face = "bold"))
-ggsave("article_profiling/fig_5a.pdf", fig_5a, height = 11, width = 19)
-ggsave("article_profiling/fig_5a.png", fig_5a, height = 11, width = 19)
-
+  theme(plot.tag = element_text(size = 20, face = "bold"))
+ggsave("article_profiling/fig_5a.pdf", fig_5a, height = 10, width = 15)
+#ggsave("article_profiling/fig_5a.png", fig_5a, height = 10, width = 15)
 fig_5b <- wrap_plots(snp_plot_psgenes, snp_plot_kos, design = "AABBB", heights = c(0.9, 1)) +
@@
   theme(plot.tag = element_text(size = 22, face = "bold"))
-ggsave("article_profiling/fig_5b.pdf", fig_5b, height = 12, width = 23)
-ggsave("article_profiling/fig_5b.png", fig_5b, height = 12, width = 23)
+ggsave("article_profiling/fig_5b.pdf", fig_5b, height = 12, width = 22)
+#ggsave("article_profiling/fig_5b.png", fig_5b, height = 12, width = 22)
+#compose
+fig_5 <- wrap_plots(wrap_plots(fig_5a, plot_spacer(), widths = c(1, 0.35)) / fig_5b) +
+  plot_annotation(tag_levels = list(c("a", "b", "c", "d"))) &
+  theme(plot.tag = element_text(size = 22, face = "bold"))
+ggsave("article_profiling/fig_5.pdf", fig_5, height = 22, width = 22)
+ggsave("article_profiling/fig_5.png", fig_5, height = 22, width = 22)
+rm(fig_5a, fig_5b) #clean memory
 
 #sup 3
-fig_s3 <- wrap_plots(snp_plot_psbins, widths = c(1)) +
-  plot_annotation(tag_levels = list(c("a", "", ""))) &
-  theme(plot.tag = element_text(size = 22, face = "bold"))
+#fig_s3 <- wrap_plots(snp_plot_psbins, widths = c(1)) +
+#  plot_annotation(tag_levels = list(c("a", "", ""))) &
+#  theme(plot.tag = element_text(size = 22, face = "bold"))
 #ggsave("article_profiling/fig_s3.pdf", fig_s3, height = 11, width = 17)
